### 2018-05-17 index.rmd
initial commit

---
title: "smooth-sea"
author: "Chris Hess"
date: "May 17, 2018"
output: 
  flexdashboard::flex_dashboard:
    navbar:
      - { title: "hesscl.com", href: "http://hesscl.com", align: right }
    orientation: columns
    theme: bootstrap
    logo: logo.svg
    source_code: "https://github.com/hesscl/smooth-sea"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 2018-05-17 index.rmd
progress on modeling, setting up format and basic data for site

@@
 title: "smooth-sea"
-author: "Chris Hess"
-date: "May 17, 2018"
 output: 
@@
 knitr::opts_chunk$set(echo = TRUE)
+
+#source("./smoothSpatiotemporal.R", echo = FALSE)
 ```
+
+Home {data-icon="fa-home"}
+=====================================
+
+Column
+-------------------------------------
+
+### Welcome!
+
+This website provides information for the paper _Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_
+
+You can download the R source code and data for this project from Github [here](https://github.com/hesscl/smooth-sea)
+
+Contact Chris Hess at [hesscl@uw.edu](mailto://hesscl@uw.edu) for more information about this report.
+
+_This page was last updated: `r Sys.Date()`_

### 2018-05-17 smoothSpatiotemporal.R

@@ -1,0 +1,492 @@
+#### INLA Rent Maps for CL ----------------------------------------------------
+#### CSSS 512
+
+#### Preamble -----------------------------------------------------------------
+
+#packages
+library(tidyverse)
+library(forcats)
+library(haven)
+library(lubridate)
+library(haven)
+library(zoo)
+library(ggthemes)
+library(RColorBrewer)
+library(sp)
+library(spdplyr)
+library(geosphere)
+library(rgdal)
+library(rgeos)
+library(sqldf)
+library(spdep)
+library(viridis)
+library(latex2exp)
+
+#setwd to location of file (REQUIRES RSTUDIO)
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+
+#load ACS data extract (codebook in folder)
+census <- read_csv(file = "./input/acsExtract.csv")
+
+#read in tract shapefiles for king county
+kc_shp <- readOGR(dsn = "R:/Project/seattle_rental_market/data/geo/KCTract2010/KCTract2010.shp",
+                  layer = "KCTract2010",
+                  GDAL1_integer64_policy = TRUE,
+                  stringsAsFactors = F)
+
+#determine config (dev or github)
+if(file.exists("../data/cl/craigslistDB.sqlite")){
+  #connect to database if file.exists
+  DB <- dbConnect(SQLite(), dbname="../data/cl/craigslistDB.sqlite")
+  cl <- tbl(DB, "clean") #create dbi for clean listing table
+  
+  #compute tract aggregates for CL listing count
+  tract <- cl %>%
+    collect %>% #bring db query into memory
+    filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft)) %>% #only listings with valid Bed/Rent
+    filter(GISJOIN %in% kc_shp@data$GISJOIN) %>% #filter to KC only (db has metro area)
+    distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
+    dplyr::select(listingDate, GEOID10, GISJOIN, seattle, matchAddress, matchType, cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
[…]

### 2018-05-19 index.rmd
some progress on index

@@
 
-#source("./smoothSpatiotemporal.R", echo = FALSE)
+#packages
+library(tidyverse)
+library(forcats)
+library(haven)
+library(lubridate)
+library(haven)
+library(zoo)
+library(ggthemes)
+library(RColorBrewer)
+library(sp)
+library(spdplyr)
+library(geosphere)
+library(rgdal)
+library(rgeos)
+library(sqldf)
+library(spdep)
+library(viridis)
+library(latex2exp)
+library(plotly)
+
+#load ACS data extract (codebook in folder)
+census <- read_csv(file = "./input/acsExtract.csv")
+
+#read in tract shapefiles for seattle
+sea_shp <- readOGR(dsn = "R:/Project/seattle_rental_market/data/geo/sea_tract_2010/sea_tract_2010.shp",
+                  layer = "sea_tract_2010",
+                  GDAL1_integer64_policy = TRUE,
+                  stringsAsFactors = F)
+
+#determine config (dev or github)
+if(file.exists("../data/cl/craigslistDB.sqlite")){
+  #connect to database if file.exists
+  DB <- dbConnect(SQLite(), dbname="../data/cl/craigslistDB.sqlite")
+  cl <- tbl(DB, "clean") #create dbi for clean listing table
+
+  #compute tract aggregates for CL listing count
+  tract <- cl %>%
+    collect %>% #bring db query into memory
+    filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
+           GISJOIN %in% sea_shp@data$GISJOIN) %>% #only listings with valid Bed/Rent, seattle tracts
+    distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
+    dplyr::select(listingDate, GEOID10, GISJOIN, seattle, matchAddress, matchType, cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
+    mutate(listingDate = as.Date(listingDate),
+           listingQtr = as.yearqtr(listingDate)) %>%
+    filter(cleanBeds %in% c(0, 1, 2, 3)) %>%
+    group_by(GEOID10, GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
+    summarize(nListings = n(),
+              n1B = sum(cleanBeds == 1),
+              nGT2B = sum(cleanBeds > 2),
+              medRent = median(cleanRent),
+              med0B = median(cleanRent[cleanBeds==0]),
+              med1B = median(cleanRent[cleanBeds==1]),
+              med2B = median(cleanRent[cleanBeds==2]),
+              med3B = median(cleanRent[cleanBeds==3]),
+              lagRent = lag(cleanRent)) %>% #create tract aggregates
+    mutate(pGT2B = nGT2B/nListings) %>%
+    ungroup %>%
+    arrange(GISJOIN, listingQtr)
+  dbDisconnect(DB)
+  write_csv(tract, "./input/tractCl.csv")
+
+} else{
+  #if github config, read in extract from cl db
+  tract <- read_csv("./input/tractCl.csv")
+}
+
+#### A. construct panel (some missingness) ------------------------------------
+
+census <- census[match(sea_shp@data$GISJOIN, census$GISJOIN),]
+
+census <- census %>%
+  mutate(nHU = AF7PE001,
+         pOwnoccHU = AF7PE002/AF7PE001,
+         medHUVal = AF9LE001)
+
+### Training periods
+
+#studios
+Q2_2017 <- tract %>%
+  filter(listingQtr == "2017 Q2") %>%
+  right_join(census) %>%
+  ungroup %>%
+  mutate(listingQtr = "2017 Q2",
+         Qtr = "Q2",
+         actualRent = med1B)
+
+Q3_2017 <- tract %>%
+  filter(listingQtr == "2017 Q3") %>%
+  right_join(census) %>%
+  ungroup %>%
+  mutate(listingQtr = "2017 Q3",
+         Qtr = "Q3",
+         actualRent = med1B)
+
+Q4_2017 <- tract %>%
+  filter(listingQtr == "2017 Q4") %>%
+  right_join(census) %>%
+  ungroup %>%
+  mutate(listingQtr = "2017 Q4",
+         Qtr = "Q4",
+         actualRent = med1B)
+
+Q1_2018 <- tract %>%
+  filter(listingQtr == "2018 Q1") %>%
+  right_join(census) %>%
+  ungroup %>%
+  mutate(listingQtr = "2018 Q1",
+         Qtr = "Q1",
+         actualRent = med1B)
+
+### Forecast/test period
+
+Q2_2018 <- tract %>%
+  filter(listingQtr == "2018 Q2") %>%
+  right_join(census) %>%
+  ungroup %>%
+  mutate(listingQtr = "2018 Q2",
+         Qtr = "Q2",
+         actualRent = med1B,
+         medRent = NA)
+Q2_2018 <- Q2_2018[match(sea_shp@data$GISJOIN, Q2_2018$GISJOIN),]
+
+sea_df <- bind_rows(Q2_2017, Q3_2017, Q4_2017, Q1_2018, Q2_2018)
+sea_df$listingQtr <- as.yearqtr(sea_df$listingQtr)
+
+sea_df <- sea_df %>% 
+  dplyr::select(-GEOID10) %>%
+  dplyr::select(GISJOIN, listingQtr, Qtr, medRent, med1B, lagRent, actualRent, nHU, 
+                pOwnoccHU, medHUVal, pGT2B)
+
+#### ACF for panel ------------------------------------------------------------
+
+acf_df <- data.frame(sea_df = unique(sea_df$GISJOIN), acf = rep(NA, length(unique(sea_df$GISJOIN))))
+for(i in unique(sea_df$GISJOIN)){
+  obs <- sea_df[sea_df$GISJOIN == i,]
+  ts <- ts(obs$medRent)
+  if(sum(is.na(ts)) > 0){
+    next
+  } else{
+    #print(ts)
+    acf <- acf(ts)
+    #print(acf)
+    acf_df$acf[acf_df$sea_df == tract] = list(acf)
+  }
+}
+
+#ACFs are weak with only 5 periods to assess with. Highest L1 value is around .5 but many are
+#weakened by a noisey quarter
+
+#### F. INLA Models of CL Rent ------------------------------------------------
+
+library(INLA)
+
+#geo information for INLA
+sea_shp <- readOGR(dsn = "R:/Project/seattle_rental_market/data/geo/sea_tract_2010/sea_tract_2010.shp",
+                   layer = "sea_tract_2010",
+                   GDAL1_integer64_policy = TRUE,
+                   stringsAsFactors = F)
+
+cens <- gCentroid(sea_shp, byid = T)
+cens <- data.frame(long = cens@coords[,1],
+                   lat = cens@coords[,2])
+cens <- cens %>% mutate(row = row_number())
+sea_ff <- fortify(sea_shp)
+
+ggplot(sea_ff, aes(x = long, y = lat, group = group)) + 
+  geom_polygon(fill = "grey90", color = "white") +
+  geom_text(data = cens, aes(x  = long, y = lat, label = row, group = row))
+
+sea_adj <- poly2nb(sea_shp)
+nb2INLA("R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph", sea_adj)
+sea_df$idtract <- rep(1:134, 5)
+sea_df$idtract1 <- rep(1:134, 5)
+sea_df$idqtr <- as.numeric(sea_df$listingQtr)
+sea_df$idqtr1 <- sea_df$idqtr
+sea_df$idtractqtr <- paste(sea_df$idtract, sea_df$idqtr)
+
+#Seattle median rent models -----------------------------------------------
+
+#fixed effect for qtr only
+form.int <- med1B ~ 1 + Qtr
+
+m.int <- inla(form.int, 
+             family = "lognormal", 
+             data = sea_df,
+             control.predictor = list(compute = TRUE),
+             control.compute = list(dic = TRUE, waic = TRUE))
+
+summary(m.int)
+sea_df$int_Med <- m.int$summary.linear.predictor[, "0.5quant"]
+sea_df$int_SD <- m.int$summary.linear.predictor[, "sd"]
+sea_df$int_postWidth <- m.int$summary.linear.predictor[, "0.975quant"] - m.int$summary.linear.predictor[,"0.025quant"]
+sea_df$int_Eff <- m.int$summary.random$idtract[sea_df$idtract, "0.5quant"]
+
+#fixed qtr + iid tract random effect
+form.ns <- med1B ~ 1 + Qtr + 
+   f(idtract, model = "iid")
+
+m.ns <- inla(form.ns, 
+              family = "lognormal", 
+              data = sea_df,
+              control.predictor = list(compute = TRUE),
+              control.compute = list(dic = TRUE, waic = TRUE))
+
+summary(m.ns)
+sea_df$ns_Med <- m.ns$summary.linear.predictor[, "0.5quant"]
+sea_df$ns_SD <- m.ns$summary.linear.predictor[, "sd"]
+sea_df$ns_postWidth <- m.ns$summary.linear.predictor[, "0.975quant"] - m.ns$summary.linear.predictor[,"0.025quant"]
+sea_df$ns_Eff <- m.ns$summary.random$idtract[sea_df$idtract, "0.5quant"]
+
+#IID tract random effect + AR(1) process
+form.nsar1 <- med1B ~ 1 +
+  f(idtract, model = "iid") +
+  f(idqtr, model = "ar1") + f(idqtr1, model = "iid")
+
+m.nsar1 <- inla(form.nsar1, 
+               family = "lognormal", 
+               data = sea_df,
+               control.predictor = list(compute = TRUE),
+               control.compute = list(dic = TRUE, waic = TRUE))
+
+summary(m.nsar1) #NB: does not improve fit to use random effect time specification, using fixed effects
+sea_df$nsar1_Med <- m.nsar1$summary.linear.predictor[, "0.5quant"]
+sea_df$nsar1_SD <- m.nsar1$summary.linear.predictor[, "sd"]
+sea_df$nsar1_postWidth <- m.nsar1$summary.linear.predictor[, "0.975quant"] - m.nsar1$summary.linear.predictor[,"0.025quant"]
+sea_df$nsar1_Eff <- m.nsar1$summary.random$idtract[sea_df$idtract, "0.5quant"]
+
+#spatial random effect model (BYM 1991) with seasonal dummies
+form.bym <- med1B ~ 1 + Qtr +
+  f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
+    scale.model = T,
+    graph = "R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph")
+
+m.bym <- inla(form.bym, 
+             family = "lognormal", 
+             data = sea_df,
+             control.predictor = list(compute = TRUE),
+             control.compute = list(dic = TRUE, waic = TRUE))
+
+summary(m.bym)
+sea_df$bym_Med <- m.bym$summary.linear.predictor[, "0.5quant"]
+sea_df$bym_SD <- m.bym$summary.linear.predictor[, "sd"]
+sea_df$bym_postWidth <- m.bym$summary.linear.predictor[, "0.975quant"] - m.bym$summary.linear.predictor[,"0.025quant"]
+sea_df$bym_Eff <- m.bym$summary.random$idtract[sea_df$idtract, "0.5quant"]
+
+#spatio-temporal smoothing model with linear space-time interaction (Bernadelli 1995)
+form.spt <- med1B ~ 1 + Qtr +
+  f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
+    scale.model = T,
+    graph = "R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph") +
+  f(idtractqtr, model="iid") + #exchangeable RE for period
+  f(idqtr, model = "iid")
+
+m.spt <- inla(form.spt, 
+              family = "lognormal", 
+              data = sea_df,
+              control.predictor = list(compute = TRUE),
+              control.compute = list(dic = TRUE, waic = TRUE))
+
+summary(m.spt)
+sea_df$spt_Med <- m.spt$summary.linear.predictor[, "0.5quant"]
+sea_df$spt_SD <- m.spt$summary.linear.predictor[, "sd"]
+sea_df$spt_postWidth <- m.spt$summary.linear.predictor[, "0.975quant"] - m.spt$summary.linear.predictor[,"0.025quant"]
+sea_df$spt_Eff <- m.spt$summary.random$idtract[sea_df$idtract, "0.5quant"]
+
+#overfitting? too much bumpiness qtr-to-qtr
+
+#### Maps of Model Output ------------------------------------------------------------------
+
+sea_df <- sea_df %>%
+  mutate_at(.vars = vars(matches("_Med")),
+            .funs = exp) %>%
+  mutate_at(.vars = vars(matches("_SD")),
+            .funs = exp) %>%
+  group_by(GISJOIN) %>% 
+  arrange(GISJOIN, listingQtr) %>%
+  mutate(spt_traj = spt_Med - lag(spt_Med, 4)) %>%
+  ungroup %>%
+  group_by(GISJOIN) %>%
+  mutate(spt_traj = sum(spt_traj, na.rm=T))
+
+sea_df
+
+sea_shp@data$id <- rownames(sea_shp@data)
+sea_shp@data <- left_join(sea_shp@data, sea_df)
+sea_f <- fortify(sea_shp)
+sea_f <- inner_join(sea_f, sea_shp@data,"id")
+
+rmse <- function(error){sqrt(mean(error^2))}
+mae <- function(error){mean(abs(error))}
+
+#table of fit statistics
+fit_stats <- sea_df %>%
+  mutate(train_test = ifelse(listingQtr == "2018 Q2", "Test", "Training")) %>%
+  mutate(int_err = actualRent - int_Med,
+         ns_err = actualRent - ns_Med,
+         nsar1_err = actualRent - nsar1_Med,
+         bym_err = actualRent - bym_Med,
+         spt_err = actualRent - spt_Med) %>% 
+  group_by(train_test) %>%
+  summarize(int_rmse = rmse(int_err[!is.na(int_err)]),
+            int_mae = mae(int_err[!is.na(int_err)]),
+            ns_rmse = rmse(ns_err[!is.na(ns_err)]),
+            ns_mae = mae(ns_err[!is.na(ns_err)]),
+            nsar1_rmse = rmse(nsar1_err[!is.na(nsar1_err)]),
+            nsar1_mae = mae(nsar1_err[!is.na(nsar1_err)]),
+            bym_rmse = rmse(bym_err[!is.na(bym_err)]),
+            bym_mae = mae(bym_err[!is.na(bym_err)]),
+            spt_rmse = rmse(spt_err[!is.na(spt_err)]),
+            spt_mae = mae(spt_err[!is.na(spt_err)])) %>%
+  dplyr::select(train_test, ends_with("_rmse"), ends_with("_mae"))
+
 ```
@@
 _This page was last updated: `r Sys.Date()`_
+
+Column
+-------------------------------------
+
+### Observed vs. Smoothed Rent Estimates
+
+```{r index.maps, echo=F}
+smoothed <- ggplot(sea_f %>% filter(listingQtr == "2018 Q2"), 
+                   aes(x = long, y = lat, group = group, fill = bym_Med, label = GISJOIN)) +
+  #facet_grid(~ listingQtr) +
+  geom_polygon(color = "grey85", lwd = .05) +
+  scale_fill_viridis_c(limits = c(750, 2500), labels = scales::dollar) +
+  coord_quickmap() +
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        panel.background = element_blank()) +
+  labs(title = "Spatially Smoothed 1B Median Rent",
+       fill = "Smoothed")
+ggplotly(smoothed, tooltip = c("label", "fill"))
+```

### 2018-05-19 smoothSpatiotemporal.R

@@ -16,7 +16,6 @@ library(sp)
 library(spdplyr)
 library(geosphere)
 library(rgdal)
-library(rgeos)
 library(sqldf)
 library(spdep)
 library(viridis)
@@ -28,9 +27,9 @@ setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 #load ACS data extract (codebook in folder)
 census <- read_csv(file = "./input/acsExtract.csv")
 
-#read in tract shapefiles for king county
-kc_shp <- readOGR(dsn = "R:/Project/seattle_rental_market/data/geo/KCTract2010/KCTract2010.shp",
-                  layer = "KCTract2010",
+#read in tract shapefiles for seattle
+sea_shp <- readOGR(dsn = "R:/Project/seattle_rental_market/data/geo/sea_tract_2010/sea_tract_2010.shp",
+                  layer = "sea_tract_2010",
                   GDAL1_integer64_policy = TRUE,
                   stringsAsFactors = F)
 
@@ -43,23 +42,22 @@ if(file.exists("../data/cl/craigslistDB.sqlite")){
   #compute tract aggregates for CL listing count
   tract <- cl %>%
     collect %>% #bring db query into memory
-    filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft)) %>% #only listings with valid Bed/Rent
-    filter(GISJOIN %in% kc_shp@data$GISJOIN) %>% #filter to KC only (db has metro area)
+    filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), GISJOIN %in% sea_shp@data$GISJOIN) %>% #only listings with valid Bed/Rent
     distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
     dplyr::select(listingDate, GEOID10, GISJOIN, seattle, matchAddress, matchType, cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
     mutate(listingDate = as.Date(listingDate),
            listingQtr = as.yearqtr(listingDate)) %>%
-    filter(listingDate >= "2017-04-01", cleanBeds %in% c(0, 1, 2, 3)) %>%
[…]

### 2018-05-20 index.rmd
slider for index map

@@
 ```{r index.maps, echo=F}
-smoothed <- ggplot(sea_f %>% filter(listingQtr == "2018 Q2"), 
-                   aes(x = long, y = lat, group = group, fill = bym_Med, label = GISJOIN)) +
+
+smoothed <- ggplot(sea_f, 
+                   aes(x = long, y = lat, group = group, fill = bym_Med, label = GISJOIN, frame = listingQtr)) +
   #facet_grid(~ listingQtr) +

### 2018-05-20 index.rmd
tidier panel computation, fixing section labels, avp forecast map as index right column

@@
            GISJOIN %in% sea_shp@data$GISJOIN) %>% #only listings with valid Bed/Rent, seattle tracts
-    distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
-    dplyr::select(listingDate, GEOID10, GISJOIN, seattle, matchAddress, matchType, cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
+    distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
+             matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
+    dplyr::select(listingDate, GISJOIN, seattle, matchAddress, matchType, 
+                  cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
     mutate(listingDate = as.Date(listingDate),
            listingQtr = as.yearqtr(listingDate)) %>%
-    filter(cleanBeds %in% c(0, 1, 2, 3)) %>%
-    group_by(GEOID10, GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
+    filter(cleanBeds %in% c(0, 1, 2, 3), listingQtr >= "2017 Q2") %>%
+    group_by(GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
     summarize(nListings = n(),
@@
 
+#ensure order of census df matches shapefile
 census <- census[match(sea_shp@data$GISJOIN, census$GISJOIN),]
 
+#clean a few fields we might use
 census <- census %>%
@@
 
-### Training periods
+#store NULL object to bind iterated dfs into
+panel <- NULL
 
-#studios
-Q2_2017 <- tract %>%
-  filter(listingQtr == "2017 Q2") %>%
-  right_join(census) %>%
-  ungroup %>%
-  mutate(listingQtr = "2017 Q2",
-         Qtr = "Q2",
-         actualRent = med1B)
+#for each unique quarter in the CL data
+for(i in unique(as.character(tract$listingQtr))){
 
-Q3_2017 <- tract %>%
-  filter(listingQtr == "2017 Q3") %>%
-  right_join(census) %>%
-  ungroup %>%
-  mutate(listingQtr = "2017 Q3",
-         Qtr = "Q3",
-         actualRent = med1B)
-
-Q4_2017 <- tract %>%
-  filter(listingQtr == "2017 Q4") %>%
-  right_join(census) %>%
-  ungroup %>%
-  mutate(listingQtr = "2017 Q4",
-         Qtr = "Q4",
-         actualRent = med1B)
-
-Q1_2018 <- tract %>%
-  filter(listingQtr == "2018 Q1") %>%
-  right_join(census) %>%
-  ungroup %>%
-  mutate(listingQtr = "2018 Q1",
-         Qtr = "Q1",
-         actualRent = med1B)
-
-### Forecast/test period
-
-Q2_2018 <- tract %>%
-  filter(listingQtr == "2018 Q2") %>%
-  right_join(census) %>%
-  ungroup %>%
-  mutate(listingQtr = "2018 Q2",
-         Qtr = "Q2",
-         actualRent = med1B,
-         medRent = NA)
-Q2_2018 <- Q2_2018[match(sea_shp@data$GISJOIN, Q2_2018$GISJOIN),]
-
-sea_df <- bind_rows(Q2_2017, Q3_2017, Q4_2017, Q1_2018, Q2_2018)
-sea_df$listingQtr <- as.yearqtr(sea_df$listingQtr)
-
-sea_df <- sea_df %>% 
-  dplyr::select(-GEOID10) %>%
-  dplyr::select(GISJOIN, listingQtr, Qtr, medRent, med1B, lagRent, actualRent, nHU, 
-                pOwnoccHU, medHUVal, pGT2B)
-
-#### ACF for panel ------------------------------------------------------------
-
-acf_df <- data.frame(sea_df = unique(sea_df$GISJOIN), acf = rep(NA, length(unique(sea_df$GISJOIN))))
-for(i in unique(sea_df$GISJOIN)){
-  obs <- sea_df[sea_df$GISJOIN == i,]
-  ts <- ts(obs$medRent)
-  if(sum(is.na(ts)) > 0){
-    next
-  } else{
-    #print(ts)
-    acf <- acf(ts)
-    #print(acf)
-    acf_df$acf[acf_df$sea_df == tract] = list(acf)
-  }
+  period <- tract %>% #start pipe with tract, end with period
+    mutate(listingQtr = factor(listingQtr, ordered = T)) %>% #make factor
+    filter(listingQtr == i) %>% #filter to qtr i
+    right_join(census) %>% #right join to census, allows missingness
+    ungroup() %>% #remove grouping by tract
+    mutate(listingQtr = i, #listingQtr == current iteration
+           Qtr = gsub(pattern = "^\\d{4} ", replacement = "", x = listingQtr), #subset to QX
+           actualRent = med1B) #actual for actual versus predicted
+  panel <- bind_rows(panel, period) #append to panel object
 }
 
-#ACFs are weak with only 5 periods to assess with. Highest L1 value is around .5 but many are
-#weakened by a noisey quarter
+sea_df <- panel %>% #set to NA for INLA to forecast
+  mutate(med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B))%>%
+  dplyr::select(GISJOIN, listingQtr, Qtr, medRent, med1B, lagRent, actualRent, nHU, n1B,
+                pOwnoccHU, medHUVal, pGT2B) %>%
+  arrange(listingQtr, GISJOIN)
 
-#### F. INLA Models of CL Rent ------------------------------------------------
+#### B. INLA Models of CL Rent ------------------------------------------------
 
@@
 
+#compute centroid for map of tract numbers
 cens <- gCentroid(sea_shp, byid = T)
@@
 
+#plot seattle tracts with tract number labels
 ggplot(sea_ff, aes(x = long, y = lat, group = group)) + 
@@
 
+#create adjacency matrix from shapefile
 sea_adj <- poly2nb(sea_shp)
+
+#create neighbor file that INLA takes for hyperparameter args
 nb2INLA("R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph", sea_adj)
-sea_df$idtract <- rep(1:134, 5)
+sea_df$idtract <- rep(1:134, 5) #a tract ID (nb: panel must be ordered tract)
 sea_df$idtract1 <- rep(1:134, 5)
-sea_df$idqtr <- as.numeric(sea_df$listingQtr)
+sea_df$idqtr <- factor(sea_df$listingQtr, ordered = T)
 sea_df$idqtr1 <- sea_df$idqtr
 sea_df$idtractqtr <- paste(sea_df$idtract, sea_df$idqtr)
-
-#Seattle median rent models -----------------------------------------------
 
@@
   group_by(GISJOIN) %>%
-  mutate(spt_traj = sum(spt_traj, na.rm=T))
+  mutate(spt_traj = sum(spt_traj, na.rm=T)) %>%
+  ungroup() %>%
+  mutate(bym_cuts = cut(bym_Med, 6),
+         spt_cuts = cut(spt_Med, 6))
 
@@
 ```{r index.maps, echo=F}
+obs <- sea_f %>%
+  dplyr::select(long, lat, group, GISJOIN, listingQtr, n1B, rent = actualRent) %>%
+  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
+  mutate(var = "Observed")
 
-smoothed <- ggplot(sea_f, 
-                   aes(x = long, y = lat, group = group, fill = bym_Med, label = GISJOIN, frame = listingQtr)) +
-  #facet_grid(~ listingQtr) +
+pred <- sea_f %>%
+  dplyr::select(long, lat, group, GISJOIN, listingQtr, n1B, rent = nsar1_Med) %>%
+  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
+  mutate(var = "Forecasted")
+
+comb <- bind_rows(obs, pred)
+
+avp.gg <- ggplot(comb, aes(x = long, y = lat, group = group, fill = rent, label = GISJOIN, text = n1B)) +
+  facet_wrap(~ var) +
   geom_polygon(color = "grey85", lwd = .05) +
@@
         panel.background = element_blank()) +
-  labs(title = "Spatially Smoothed 1B Median Rent",
-       fill = "Smoothed")
-ggplotly(smoothed, tooltip = c("label", "fill"))
+  labs(fill = "Median 1B Rent")
+
+ggplotly(avp.gg, tooltip = c("facet", "label", "fill", "text"), hovermode="compare") %>%
+  animation_opts(frame = 1000, redraw = FALSE) 
 ```
+
+Home {data-icon="fa-home"}
+=====================================
+
+Home {data-icon="fa-home"}
+=====================================

### 2018-05-20 smoothSpatiotemporal.R

@@ -47,7 +47,7 @@ if(file.exists("../data/cl/craigslistDB.sqlite")){
     dplyr::select(listingDate, GEOID10, GISJOIN, seattle, matchAddress, matchType, cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
     mutate(listingDate = as.Date(listingDate),
            listingQtr = as.yearqtr(listingDate)) %>%
-    filter(cleanBeds %in% c(0, 1, 2, 3)) %>%
+    filter(cleanBeds %in% c(0, 1, 2, 3), listingQtr >= "2017 Q2") %>%
     group_by(GEOID10, GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
     summarize(nListings = n(),
               n1B = sum(cleanBeds == 1),
@@ -72,72 +72,36 @@ if(file.exists("../data/cl/craigslistDB.sqlite")){
 
 #### A. construct panel (some missingness) ------------------------------------
 
+#ensure order of census df matches shapefile
 census <- census[match(sea_shp@data$GISJOIN, census$GISJOIN),]
 
+#clean a few fields we might use
 census <- census %>%
   mutate(nHU = AF7PE001,
          pOwnoccHU = AF7PE002/AF7PE001,
          medHUVal = AF9LE001)
 
-### Training periods
+#store NULL object to bind iterated dfs into
+panel <- NULL
 
-#studios
-Q2_2017 <- tract %>%
-  filter(listingQtr == "2017 Q2") %>%
-  right_join(census) %>%
-  ungroup %>%
-  mutate(listingQtr = "2017 Q2",
-         Qtr = "Q2",
-         actualRent = med1B)
+#for each unique quarter in the CL data
+for(i in unique(as.character(tract$listingQtr))){
+  
+  period <- tract %>% #start pipe with tract, end with period
+    mutate(listingQtr = factor(listingQtr, ordered = T)) %>% #make factor
+    filter(listingQtr == i) %>% #filter to qtr i
+    right_join(census) %>% #right join to census, allows missingness
+    ungroup() %>% #remove grouping by tract
+    mutate(listingQtr = i, #listingQtr == current iteration
+           Qtr = gsub(pattern = "Q\\d", replacement = "", x = listingQtr), #subset to QX
[…]

### 2018-05-21 index.rmd
updated index.html to full set of pages, still under construction

@@
 #packages
-library(tidyverse)
+library(tidyverse) #need dev version of ggplot
 library(forcats)
@@
 library(plotly)
+library(DT)
+library(crosstalk)
 
@@
               n1B = sum(cleanBeds == 1),
-              nGT2B = sum(cleanBeds > 2),
               medRent = median(cleanRent),
@@
               lagRent = lag(cleanRent)) %>% #create tract aggregates
-    mutate(pGT2B = nGT2B/nListings) %>%
     ungroup %>%
@@
 
+full_panel <- tract %>%
+  group_by(GISJOIN) %>% 
+  summarize(n = n()) %>%
+  pull(n) %>%
+  max
+
 #### A. construct panel (some missingness) ------------------------------------
@@
   mutate(nHU = AF7PE001,
+         idtract = row_number(),
          pOwnoccHU = AF7PE002/AF7PE001,
          medHUVal = AF9LE001)
+
+#periods <- tract %>%
+#  filter(listingQtr != as.yearqtr(Sys.Date())) %>%
+#  summarize(n = length(unique(listingQtr))) %>% 
+#  pull(n)
 
@@
     mutate(listingQtr = i, #listingQtr == current iteration
-           Qtr = gsub(pattern = "^\\d{4} ", replacement = "", x = listingQtr), #subset to QX
-           actualRent = med1B) #actual for actual versus predicted
+           Qtr = gsub(pattern = "^\\d{4} ", replacement = "", x = listingQtr)) #subset str to Q1,2,3,4
   panel <- bind_rows(panel, period) #append to panel object
@@
 sea_df <- panel %>% #set to NA for INLA to forecast
-  mutate(med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B))%>%
-  dplyr::select(GISJOIN, listingQtr, Qtr, medRent, med1B, lagRent, actualRent, nHU, n1B,
-                pOwnoccHU, medHUVal, pGT2B) %>%
+  mutate(actualRent = med1B,
+         med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B)) %>%
+  dplyr::select(GISJOIN, listingQtr, Qtr, medRent, med1B, lagRent, actualRent, nHU,
+                n1B, nListings, pOwnoccHU, medHUVal, idtract) %>%
+  mutate(p1B = n1B/nListings) %>%
   arrange(listingQtr, GISJOIN)
@@
 
-#geo information for INLA
-sea_shp <- readOGR(dsn = "R:/Project/seattle_rental_market/data/geo/sea_tract_2010/sea_tract_2010.shp",
-                   layer = "sea_tract_2010",
-                   GDAL1_integer64_policy = TRUE,
-                   stringsAsFactors = F)
-
-#compute centroid for map of tract numbers
-cens <- gCentroid(sea_shp, byid = T)
-cens <- data.frame(long = cens@coords[,1],
-                   lat = cens@coords[,2])
-cens <- cens %>% mutate(row = row_number())
-sea_ff <- fortify(sea_shp)
-
-#plot seattle tracts with tract number labels
-ggplot(sea_ff, aes(x = long, y = lat, group = group)) + 
-  geom_polygon(fill = "grey90", color = "white") +
-  geom_text(data = cens, aes(x  = long, y = lat, label = row, group = row))
-
 #create adjacency matrix from shapefile
@@
 nb2INLA("R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph", sea_adj)
-sea_df$idtract <- rep(1:134, 5) #a tract ID (nb: panel must be ordered tract)
-sea_df$idtract1 <- rep(1:134, 5)
 sea_df$idqtr <- factor(sea_df$listingQtr, ordered = T)
@@
 sea_df$idtractqtr <- paste(sea_df$idtract, sea_df$idqtr)
+sea_df$idtract1 <- sea_df$idtract
 
 #fixed effect for qtr only
-form.int <- med1B ~ 1 + Qtr
+form.int <- log(med1B) ~ 1 + Qtr
 
 m.int <- inla(form.int, 
-             family = "lognormal", 
+             family = "normal", 
              data = sea_df,
@@
 #fixed qtr + iid tract random effect
-form.ns <- med1B ~ 1 + Qtr + 
+form.ns <- log(med1B) ~ 1 + Qtr +
    f(idtract, model = "iid")
@@
 m.ns <- inla(form.ns, 
-              family = "lognormal", 
+              family = "normal", 
               data = sea_df,
@@
 #IID tract random effect + AR(1) process
-form.nsar1 <- med1B ~ 1 +
+form.nsar1 <- log(med1B) ~ 1 +
   f(idtract, model = "iid") +
@@
 m.nsar1 <- inla(form.nsar1, 
-               family = "lognormal", 
+               family = "normal", 
                data = sea_df,
@@
 #spatial random effect model (BYM 1991) with seasonal dummies
-form.bym <- med1B ~ 1 + Qtr +
+form.bym <- log(med1B) ~ 1 + Qtr +
   f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
@@
 m.bym <- inla(form.bym, 
-             family = "lognormal", 
+             family = "normal", 
              data = sea_df,
@@
 #spatio-temporal smoothing model with linear space-time interaction (Bernadelli 1995)
-form.spt <- med1B ~ 1 + Qtr +
+form.spt <- log(med1B) ~ 1 + Qtr +
   f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
@@
   f(idtractqtr, model="iid") + #exchangeable RE for period
-  f(idqtr, model = "iid")
+  f(idqtr, model = "rw1")
 
 m.spt <- inla(form.spt, 
-              family = "lognormal", 
+              family = "normal", 
               data = sea_df,
@@
 fit_stats <- sea_df %>%
-  mutate(train_test = ifelse(listingQtr == "2018 Q2", "Test", "Training")) %>%
+  mutate(train_test = ifelse(listingQtr == as.yearqtr(Sys.Date()), "Test", "Training")) %>%
   mutate(int_err = actualRent - int_Med,
@@
   dplyr::select(train_test, ends_with("_rmse"), ends_with("_mae"))
-
 ```
@@
 obs <- sea_f %>%
-  dplyr::select(long, lat, group, GISJOIN, listingQtr, n1B, rent = actualRent) %>%
+  mutate(rent = actualRent) %>%
   filter(listingQtr == as.yearqtr(Sys.Date())) %>%
-  mutate(var = "Observed")
+  mutate(var = paste(as.yearqtr(Sys.Date()), "to-Date"))
 
 pred <- sea_f %>%
-  dplyr::select(long, lat, group, GISJOIN, listingQtr, n1B, rent = nsar1_Med) %>%
+  mutate(rent = nsar1_Med) %>%
   filter(listingQtr == as.yearqtr(Sys.Date())) %>%
-  mutate(var = "Forecasted")
+  mutate(var = paste(as.yearqtr(Sys.Date()), "Forecast"))
 
-comb <- bind_rows(obs, pred)
+comb <- bind_rows(obs, pred) %>%
+  mutate(tooltip = paste0(NAMELSAD10, "<br>", "Forecast: $", round(nsar1_Med), "<br>", "Qtr-to-Date: $", round(actualRent), "<br>", "N Qtr-to-Date: ", n1B))
+comb <- SharedData$new(comb, ~GISJOIN)
 
-avp.gg <- ggplot(comb, aes(x = long, y = lat, group = group, fill = rent, label = GISJOIN, text = n1B)) +
+avp_map <- ggplot(comb, aes(x = long, y = lat, group = group, fill = rent, label = tooltip)) +
   facet_wrap(~ var) +
@@
         panel.background = element_blank()) +
-  labs(fill = "Median 1B Rent")
+  labs(fill = "Median<br>1B Rent")
 
-ggplotly(avp.gg, tooltip = c("facet", "label", "fill", "text"), hovermode="compare") %>%
-  animation_opts(frame = 1000, redraw = FALSE) 
+ggplotly(avp_map, tooltip = c("label")) %>%
+  layout(autosize = T) %>%
+  highlight(on = "plotly_hover", off = NULL)
 ```
 
-Home {data-icon="fa-home"}
+Distribution
 =====================================
 
-Home {data-icon="fa-home"}
+```{r, echo=F}
+gg_hist <- ggplot(sea_df, aes(x = med1B, group = listingQtr, color = listingQtr)) +
+  geom_density() +
+  theme_minimal() +
+  scale_color_viridis_d(end = .8)
+
+ggplotly(gg_hist) %>%
+  layout(autosize = T)
+```
+
+Panel Time-Series
 =====================================
+
+```{r temp.lineplot, echo=F}
+obs_df <- sea_df %>%
+  group_by(listingQtr) %>%
+  mutate(value = actualRent,
+         qtrDev = value-mean(value, na.rm=T),
+         model = "A. Observed")
+
+int_df <- sea_df %>%
+  group_by(listingQtr) %>%
+  mutate(value = int_Med,
+         qtrDev = value-mean(value, na.rm=T),
+         model = "B. Quarter FE")
+
+ns_df <- sea_df %>%
+  group_by(listingQtr) %>%
+  mutate(value = ns_Med,
+         qtrDev = value-mean(value, na.rm=T),
+         model = "C. Non-Spatial RE")
+
+nsar1_df <- sea_df %>%
+  group_by(listingQtr) %>%
+  mutate(value = nsar1_Med,
+         qtrDev = value-mean(value, na.rm=T),
+         model = "D. Non-Spatial AR(1) RE")
+
+bym_df <- sea_df %>%
+  group_by(listingQtr) %>%
+  mutate(value = bym_Med,
+         qtrDev = value-mean(value, na.rm=T),
+         model = "E. Spatial RE")
+
+spt_df <- sea_df %>%
+  group_by(listingQtr) %>%
+  mutate(value =  spt_Med,
+         qtrDev = value-mean(value, na.rm=T),
+         model = "F. Spatiotemporal")
+
+panel_df <- bind_rows(obs_df, int_df, ns_df, nsar1_df, bym_df, spt_df)
+
+panel_df <- SharedData$new(panel_df, ~GISJOIN)
+
+obs_lineplot <- ggplot(panel_df, aes(x = listingQtr, y = value, group = GISJOIN, label = GISJOIN, color = qtrDev, frame = model)) +
+  geom_line(lwd = .25, color = "black", alpha = .5) +
+  geom_point() +
+  scale_color_gradient2_tableau()
+
+ggplotly(obs_lineplot, tooltip = c("label", "y")) %>%
+  layout(autosize = T) %>%
+  animation_opts(frame = 1000, transition = 0, redraw = FALSE) %>%
+  animation_button(hide = T) %>%
+  highlight(on = "plotly_hover", off = NULL)
+```
+
+Spatial Time-Series
+=====================================
+
+```{r spat.series, echo=F}
+map_panel <- SharedData$new(sea_f, ~GISJOIN)
+
+map_panel_gg <- ggplot(map_panel, aes(x = long, y = lat, group = group, fill = actualRent, label = GISJOIN, frame = listingQtr)) +
+  geom_polygon(data=sea_f, aes(fill = NULL, label = NULL), fill = "grey80") +
+  geom_polygon(color = "grey85", lwd = .05) +
+  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
+  coord_quickmap() +
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        panel.background = element_blank()) +
+  labs(fill = "Median<br>1B Rent")
+
+ggplotly(map_panel_gg, tooltip = c("label", "fill")) %>%
+  layout(autosize = T) %>%
+  animation_opts(frame = 1000, transition = 0, redraw = FALSE)
+```
+
+Model Fit
+=====================================
+
+Column {data-width=400}
+-------------------------------------
+
+### RMSE
+
+```{r, echo=F}
+knitr::kable(fit_stats %>% select(train_test, ends_with("rmse")))
+```
+
+### MAE
+
+```{r, echo=F}
+knitr::kable(fit_stats %>% select(train_test, ends_with("mae")))
+```
+
+Column {data-width=600}
+-------------------------------------
+
+### Actual vs. Predicted Scatterplot
+
+```{r, echo=F}
+avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = bym_Med, size = n1B, frame = listingQtr)) +
+  geom_point() +
+  geom_abline(intercept = 0, slope = 1) +
+  scale_size_area() +
+  coord_equal() +
+  theme_minimal()
+
+ggplotly(avp_scatter) %>%
+  layout(autosize = T)
+```
+
+Forecast
+=====================================

### 2018-05-21 smoothSpatiotemporal.R

@@ -42,23 +42,24 @@ if(file.exists("../data/cl/craigslistDB.sqlite")){
   #compute tract aggregates for CL listing count
   tract <- cl %>%
     collect %>% #bring db query into memory
-    filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), GISJOIN %in% sea_shp@data$GISJOIN) %>% #only listings with valid Bed/Rent
-    distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
-    dplyr::select(listingDate, GEOID10, GISJOIN, seattle, matchAddress, matchType, cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
+    filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
+           GISJOIN %in% sea_shp@data$GISJOIN) %>% #only listings with valid Bed/Rent, seattle tracts
+    distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
+             matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
+    dplyr::select(listingDate, GISJOIN, seattle, matchAddress, matchType, 
+                  cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
     mutate(listingDate = as.Date(listingDate),
            listingQtr = as.yearqtr(listingDate)) %>%
     filter(cleanBeds %in% c(0, 1, 2, 3), listingQtr >= "2017 Q2") %>%
-    group_by(GEOID10, GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
+    group_by(GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
     summarize(nListings = n(),
               n1B = sum(cleanBeds == 1),
-              nGT2B = sum(cleanBeds > 2),
               medRent = median(cleanRent),
               med0B = median(cleanRent[cleanBeds==0]),
               med1B = median(cleanRent[cleanBeds==1]),
               med2B = median(cleanRent[cleanBeds==2]),
               med3B = median(cleanRent[cleanBeds==3]),
               lagRent = lag(cleanRent)) %>% #create tract aggregates
[…]

### 2018-05-22 index.rmd
first draft complete

@@
 
-full_panel <- tract %>%
-  group_by(GISJOIN) %>% 
-  summarize(n = n()) %>%
-  pull(n) %>%
-  max
-
 #### A. construct panel (some missingness) ------------------------------------
@@
 
-#periods <- tract %>%
-#  filter(listingQtr != as.yearqtr(Sys.Date())) %>%
-#  summarize(n = length(unique(listingQtr))) %>% 
-#  pull(n)
-
 #store NULL object to bind iterated dfs into
@@
 for(i in unique(as.character(tract$listingQtr))){
-
   period <- tract %>% #start pipe with tract, end with period
@@
 sea_df$nsar1_postWidth <- m.nsar1$summary.linear.predictor[, "0.975quant"] - m.nsar1$summary.linear.predictor[,"0.025quant"]
-sea_df$nsar1_Eff <- m.nsar1$summary.random$idtract[sea_df$idtract, "0.5quant"]
+sea_df$nsar1_Eff <- m.nsar1$summary.random$idqtr[sea_df$idqtr, "0.5quant"]
 
@@
 
-#spatio-temporal smoothing model with linear space-time interaction (Bernadelli 1995)
-form.spt <- log(med1B) ~ 1 + Qtr +
+#spatio-temporal smoothing model, AR(1) process with linear space-time interaction (Bernadelli 1995)
+form.spt <- log(med1B) ~ 1 +
   f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
@@
     graph = "R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph") +
-  f(idtractqtr, model="iid") + #exchangeable RE for period
-  f(idqtr, model = "rw1")
+  f(idqtr, model = "ar1") +
+  f(idqtr1, model = "iid") +
+  f(idtractqtr, model = "iid")
 
@@
 
-This website provides information for the paper _Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_
+This website provides information for the paper _Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_.
+The goal of this research is to a.) understand the temporal dynamics of rent estimates in Seattle and b.) forecast the current quarter's rent levels based off of the prior periods.
+Throughout the following analyses, the training data are 2017 Q2 up to the prior quarter (i.e.
+`r as.yearqtr(Sys.Date())-.25`).
+The test period is a forecast for the current period and includes comparison to the appropriate median rent estimates for data observed so far in this period.
 
-You can download the R source code and data for this project from Github [here](https://github.com/hesscl/smooth-sea)
+Most graphics include some level of interactivity, usually either hover-over tooltip information or a slider to control various views of the graphic.
+You can download the source code and data for this project from Github [here](https://github.com/hesscl/smooth-sea)
 
@@
 _This page was last updated: `r Sys.Date()`_
+
+<br>
+<hr>
+
+**Table of Contents**
+
+*  Distribution: density graphic to investigate the distribution of rents among Seattle neighborhoods for each quarter
+*  Panel Time-Series: line graphic to show the observed or modeled temporal structure
+*  Spatial Time-Series: series of maps to show observed change across time
+*  Model fit: tables of model root mean square error (RMSE), mean absolute error (MAE), and deviance information criterion (DIC) across training and test data
 
@@
   filter(listingQtr == as.yearqtr(Sys.Date())) %>%
-  mutate(var = paste(as.yearqtr(Sys.Date()), "to-Date"))
+  mutate(var = paste0(as.yearqtr(Sys.Date()), "-to-Date")) %>%
+  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, spt_Med, NAMELSAD10, n1B)
 
 pred <- sea_f %>%
-  mutate(rent = nsar1_Med) %>%
+  mutate(rent = spt_Med) %>%
   filter(listingQtr == as.yearqtr(Sys.Date())) %>%
-  mutate(var = paste(as.yearqtr(Sys.Date()), "Forecast"))
+  mutate(var = paste0(as.yearqtr(Sys.Date()), " Forecast")) %>%
+  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, spt_Med, NAMELSAD10, n1B)
 
 comb <- bind_rows(obs, pred) %>%
-  mutate(tooltip = paste0(NAMELSAD10, "<br>", "Forecast: $", round(nsar1_Med), "<br>", "Qtr-to-Date: $", round(actualRent), "<br>", "N Qtr-to-Date: ", n1B))
+  mutate(tooltip = paste0(NAMELSAD10, "<br>", "Forecast: $", round(spt_Med), "<br>", "Qtr-to-Date: $", round(actualRent), "<br>", "N Qtr-to-Date: ", n1B))
 comb <- SharedData$new(comb, ~GISJOIN)
@@
 
-ggplotly(avp_map, tooltip = c("label")) %>%
-  layout(autosize = T) %>%
-  highlight(on = "plotly_hover", off = NULL)
+ggplotly(avp_map, tooltip = "label") %>%
+  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
+  highlight(on = "plotly_click")
 ```
@@
   theme_minimal() +
-  scale_color_viridis_d(end = .8)
+  scale_x_continuous(labels = scales::dollar) +
+  scale_color_viridis_d(end = .8) +
+  xlab("\nMedian 1B Rent") +
+  labs(color = "Quarter")
 
-ggplotly(gg_hist) %>%
+ggplotly(gg_hist, tooltip = c("color", "x")) %>%
   layout(autosize = T)
@@
          qtrDev = value-mean(value, na.rm=T),
-         model = "B. Quarter FE")
+         model = "B. Quarter FE") %>%
+  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
 
@@
          qtrDev = value-mean(value, na.rm=T),
-         model = "C. Non-Spatial RE")
+         model = "C. Non-Spatial RE") %>%
+  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
 
@@
          qtrDev = value-mean(value, na.rm=T),
-         model = "D. Non-Spatial AR(1) RE")
+         model = "D. Non-Spatial AR(1) RE") %>%
+  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
 
@@
          qtrDev = value-mean(value, na.rm=T),
-         model = "E. Spatial RE")
+         model = "E. Spatial RE") %>%
+  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
 
@@
          qtrDev = value-mean(value, na.rm=T),
-         model = "F. Spatiotemporal")
+         model = "F. Spatiotemporal AR(1)") %>%
+  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
 
@@
   geom_point() +
-  scale_color_gradient2_tableau()
+  scale_color_gradient2_tableau() +
+  scale_y_continuous(labels = scales::dollar) +
+  xlab("") +
+  ylab("Median 1B Rent\n") +
+  theme_minimal() +
+  labs(color = "Quarterly\ndeviation")
 
@@
   layout(autosize = T) %>%
-  animation_opts(frame = 1000, transition = 0, redraw = FALSE) %>%
+  animation_opts(frame = 1000, easing = "linear", redraw = FALSE) %>%
   animation_button(hide = T) %>%
-  highlight(on = "plotly_hover", off = NULL)
+  highlight()
 ```
@@
 
-```{r spat.series, echo=F}
-map_panel <- SharedData$new(sea_f, ~GISJOIN)
+Column {.tabset}
+-------------------------------------
 
-map_panel_gg <- ggplot(map_panel, aes(x = long, y = lat, group = group, fill = actualRent, label = GISJOIN, frame = listingQtr)) +
-  geom_polygon(data=sea_f, aes(fill = NULL, label = NULL), fill = "grey80") +
+### Observed
+
+```{r spat.series.actual, echo=F}
+map_panel_obs <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = actualRent, frame = listingQtr)) +
   geom_polygon(color = "grey85", lwd = .05) +
@@
 
-ggplotly(map_panel_gg, tooltip = c("label", "fill")) %>%
+ggplotly(map_panel_obs) %>%
   layout(autosize = T) %>%
-  animation_opts(frame = 1000, transition = 0, redraw = FALSE)
+  animation_opts(transition = 0, redraw = FALSE)
+```
+
+### Spatiotemporal AR(1) RE
+
+```{r spat.series.nsar, echo=F}
+map_panel_spt <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = spt_Med, frame = listingQtr)) +
+  geom_polygon(color = "grey85", lwd = .05) +
+  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
+  coord_quickmap() +
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        panel.background = element_blank()) +
+  labs(fill = "Median<br>1B Rent")
+
+ggplotly(map_panel_spt) %>%
+  layout(autosize = T) %>%
+  animation_opts(transition = 0, redraw = FALSE)
 ```
@@
 
-Column {data-width=400}
+Column 
 -------------------------------------
+
+### Model legend
+
+* int - Quarter fixed intercept
+* ns - Non-spatial tract random effect for each tract, quarter fixed intercept
+* nsar1 - Non-spatial tract random effect model, AR(1) random effect for quarter, i.i.d. random effect for quarter
+* bym - Spatial intrinsic conditional autoregressive (ICAR) tract random-effect with quarter fixed intercept
+* spt - Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, AR(1) random effect for quarter, i.i.d. random effect for quarter and tract-quarter
 
@@
 
-Column {data-width=600}
+### DIC
+```{r, echo=F}
+knitr::kable(data.frame("int_DIC" = m.int$dic$dic,
+                        "ns_DIC" = m.ns$dic$dic,
+                        "nsar1_DIC" = m.nsar1$dic$dic,
+                        "bym_DIC" = m.bym$dic$dic,
+                        "spt_DIC" = m.spt$dic$dic))
+```
+
+Column {.tabset}
 -------------------------------------
 
-### Actual vs. Predicted Scatterplot
+### Non-Spatial AR(1)
+
+```{r, echo=F}
+avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = nsar1_Med, size = n1B, frame = listingQtr)) +
+  geom_point() +
+  geom_abline(intercept = 0, slope = 1) +
+  scale_x_continuous(labels = scales::dollar) +
+  scale_y_continuous(labels = scales::dollar) +
+  xlab("\nActual") +
+  ylab("Predicted\n") +
+  scale_size_area() +
+  coord_equal() +
+  theme_minimal()
+
+ggplotly(avp_scatter) %>%
+  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
+  animation_opts(1000, easing = "linear", redraw = FALSE)
+```
+
+### Spatial RE
 
@@
   geom_abline(intercept = 0, slope = 1) +
+  scale_x_continuous(labels = scales::dollar) +
+  scale_y_continuous(labels = scales::dollar) +
+  xlab("\nActual") +
+  ylab("Predicted\n") +
   scale_size_area() +
@@
 ggplotly(avp_scatter) %>%
-  layout(autosize = T)
+  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
+  animation_opts(1000, easing = "linear", redraw = FALSE)
 ```
 
-Forecast
-=====================================
+### Spatiotemporal AR(1)
+
+```{r, echo=F}
+avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = spt_Med, size = n1B, frame = listingQtr)) +
+  geom_point() +
+  geom_abline(intercept = 0, slope = 1) +
+  scale_x_continuous(labels = scales::dollar) +
+  scale_y_continuous(labels = scales::dollar) +
+  xlab("\nActual") +
+  ylab("Predicted\n") +
+  scale_size_area() +
+  coord_equal() +
+  theme_minimal()
+
+ggplotly(avp_scatter) %>%
+  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
+  animation_opts(1000, easing = "linear", redraw = FALSE)
+```

### 2018-05-22 index.rmd
slight change to obs spatial time series

@@
 ```{r spat.series.actual, echo=F}
-map_panel_obs <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = actualRent, frame = listingQtr)) +
+map_panel_obs <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = actualRent)) +
+  facet_grid(~ listingQtr) +
+  geom_polygon(color = "grey85", lwd = .05, ) +
+  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
+  coord_quickmap() +
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        panel.background = element_blank()) +
+  labs(fill = "Median<br>1B Rent")
+
+ggplotly(map_panel_obs) %>%
+  layout(autosize = T)
+```
+
+### Spatial RE
+
+```{r spat.series.bym, echo=F}
+map_panel_bym <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = bym_Med, frame = listingQtr)) +
   geom_polygon(color = "grey85", lwd = .05) +
@@
 
-ggplotly(map_panel_obs) %>%
+ggplotly(map_panel_bym) %>%
   layout(autosize = T) %>%
-  animation_opts(transition = 0, redraw = FALSE)
+  animation_opts(1000, transition = 0, redraw = FALSE)
 ```
@@
 
-```{r spat.series.nsar, echo=F}
+```{r spat.series.spt, echo=F}
 map_panel_spt <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = spt_Med, frame = listingQtr)) +
@@
   layout(autosize = T) %>%
-  animation_opts(transition = 0, redraw = FALSE)
+  animation_opts(1000, transition = 0, redraw = FALSE)
 ```

### 2018-05-22 index.rmd
a few small changes

@@
 
-This website provides information for the paper _Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_.
+This website provides interactive results for the forecasting models explored in the paper _Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_.
+
 The goal of this research is to a.) understand the temporal dynamics of rent estimates in Seattle and b.) forecast the current quarter's rent levels based off of the prior periods.
@@
 Most graphics include some level of interactivity, usually either hover-over tooltip information or a slider to control various views of the graphic.
+Clicking on cases will highlight data elements in most graphics, and double-clicking will reset the graphic.
 You can download the source code and data for this project from Github [here](https://github.com/hesscl/smooth-sea)
 
-Contact Chris Hess at [hesscl@uw.edu](mailto://hesscl@uw.edu) for more information about this report.
+Contact Chris Hess at [hesscl@uw.edu](mailto://hesscl@uw.edu) for more information about this research.
 
@@
 <hr>
+<br>
 

### 2018-05-22 index.rmd
removing transition for avp scatter

@@
 The goal of this research is to a.) understand the temporal dynamics of rent estimates in Seattle and b.) forecast the current quarter's rent levels based off of the prior periods.
-Throughout the following analyses, the training data are 2017 Q2 up to the prior quarter (i.e.
+Throughout the following analyses, the training data are `r min(sea_df$listingQtr)` up to the prior quarter (i.e.
 `r as.yearqtr(Sys.Date())-.25`).
@@
   scale_y_continuous(labels = scales::dollar) +
-  xlab("\nActual") +
-  ylab("Predicted\n") +
+  xlab("Actual\n") +
+  ylab("\nPredicted") +
   scale_size_area() +
@@
   layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
-  animation_opts(1000, easing = "linear", redraw = FALSE)
+  animation_opts(1000, transition = 0, redraw = T)
 ```
@@
   layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
-  animation_opts(1000, easing = "linear", redraw = FALSE)
+  animation_opts(1000, transition = 0, redraw = FALSE)
 ```

### 2018-05-23 index.rmd
solid draft, updating .gitignore for demo data add

@@
 #read in tract shapefiles for seattle
-sea_shp <- readOGR(dsn = "R:/Project/seattle_rental_market/data/geo/sea_tract_2010/sea_tract_2010.shp",
+sea_shp <- readOGR(dsn = "./input/sea_tract_2010/sea_tract_2010.shp",
                   layer = "sea_tract_2010",
@@
 #create neighbor file that INLA takes for hyperparameter args
-nb2INLA("R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph", sea_adj)
+nb2INLA("./output/seatract.graph", sea_adj)
 sea_df$idqtr <- factor(sea_df$listingQtr, ordered = T)
@@
     scale.model = T,
-    graph = "R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph")
+    graph = "./output/seatract.graph")
 
@@
     scale.model = T,
-    graph = "R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph") +
+    graph = "./output/seatract.graph") +
   f(idqtr, model = "ar1") +
@@
 The goal of this research is to a.) understand the temporal dynamics of rent estimates in Seattle and b.) forecast the current quarter's rent levels based off of the prior periods.
+The focal series of models regress median one-bedroom rent asked values on different specifications of the panel's correlation structure (i.e.
+temporal and spatial).
+All of the candidate models are estimated with [integrated nested Laplace approximations](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) (INLA) using the default, weakly-informative priors for all model hyperparameters.
 Throughout the following analyses, the training data are `r min(sea_df$listingQtr)` up to the prior quarter (i.e.
@@
 Clicking on cases will highlight data elements in most graphics, and double-clicking will reset the graphic.
-You can download the source code and data for this project from Github [here](https://github.com/hesscl/smooth-sea)
+You can download the source code and data for this project from Github [here](https://github.com/hesscl/smooth-sea).
 
@@
 
-*  Distribution: density graphic to investigate the distribution of rents among Seattle neighborhoods for each quarter
-*  Panel Time-Series: line graphic to show the observed or modeled temporal structure
-*  Spatial Time-Series: series of maps to show observed change across time
-*  Model fit: tables of model root mean square error (RMSE), mean absolute error (MAE), and deviance information criterion (DIC) across training and test data
+Page                |  Description
+------------------- | ---------------------------------------------------------
+Distribution        | density graphic to investigate the distribution of rents among Seattle neighborhoods for each quarter
+Panel Time-Series   | line graphic to show the observed or modeled temporal structure
+Spatial Time-Series | series of maps to show observed change across time
+Model Fit           | tables of model root mean square error (RMSE), mean absolute error (MAE), and deviance information criterion (DIC) across training and test data
 
@@
   facet_grid(~ listingQtr) +
-  geom_polygon(color = "grey85", lwd = .05, ) +
+  geom_polygon(color = "grey85", lwd = .05) +
   scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
@@
   layout(autosize = T)
+```
+
+### Non-Spatial AR(1)
+
+```{r spat.series.nsar1, echo=F}
+map_panel_nsar1 <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = nsar1_Med, frame = listingQtr)) +
+  geom_polygon(color = "grey85", lwd = .05) +
+  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
+  coord_quickmap() +
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        panel.background = element_blank()) +
+  labs(fill = "Median<br>1B Rent")
+
+ggplotly(map_panel_nsar1) %>%
+  layout(autosize = T) %>%
+  animation_opts(1000, transition = 0, redraw = FALSE)
 ```
@@
 
-Column 
+Column {.tabset}
 -------------------------------------
 
-### Model legend
+### Model Legend
 
-* int - Quarter fixed intercept
-* ns - Non-spatial tract random effect for each tract, quarter fixed intercept
-* nsar1 - Non-spatial tract random effect model, AR(1) random effect for quarter, i.i.d. random effect for quarter
-* bym - Spatial intrinsic conditional autoregressive (ICAR) tract random-effect with quarter fixed intercept
-* spt - Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, AR(1) random effect for quarter, i.i.d. random effect for quarter and tract-quarter
+Model abbr.                   | Description
+----------------------------- | -----------------------------------------------
+**int**    | Quarter fixed intercept
+           | `r format(form.int, )`
+**ns**     | Non-spatial tract random effect for each tract, quarter fixed intercept
+           | `r format(form.ns)`
+**nsar1**  | Non-spatial tract random effect model, AR(1) random effect for quarter, i.i.d. random effect for quarter 
+           | `r format(form.nsar1)`
+**bym**    | Spatial intrinsic conditional autoregressive (ICAR) tract random-effect with quarter fixed intercept 
+           |  `r format(form.bym)`
+**spt**    | Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, AR(1) random effect for quarter, i.i.d. random effect for quarter and tract-quarter
+           | `r format(form.spt)`
 
-### RMSE
+### Accuracy and Information Criteria
 
 ```{r, echo=F}
-knitr::kable(fit_stats %>% select(train_test, ends_with("rmse")))
+knitr::kable(fit_stats %>% dplyr::select(train_test, ends_with("rmse")))
 ```
 
-### MAE
+<br>
+<hr>
+<br>
 
 ```{r, echo=F}
-knitr::kable(fit_stats %>% select(train_test, ends_with("mae")))
+knitr::kable(fit_stats %>% dplyr::select(train_test, ends_with("mae")))
 ```
 
-### DIC
+<br>
+<hr>
+<br>
+
 ```{r, echo=F}
-knitr::kable(data.frame("int_DIC" = m.int$dic$dic,
+knitr::kable(data.frame("train_test" = "Training",
+                        "int_DIC" = m.int$dic$dic,
                         "ns_DIC" = m.ns$dic$dic,
@@
                         "spt_DIC" = m.spt$dic$dic))
+```
+
+<br>
+<hr>
+<br>
+
+```{r, echo=F}
+knitr::kable(data.frame("train_test" = "Training",
+                        "int_WAIC" = m.int$waic$waic,
+                        "ns_WAIC" = m.ns$waic$waic,
+                        "nsar1_WAIC" = m.nsar1$waic$waic,
+                        "bym_WAIC" = m.bym$waic$waic,
+                        "spt_WAIC" = m.spt$waic$waic))
 ```
@@
   scale_y_continuous(labels = scales::dollar) +
-  xlab("Actual\n") +
-  ylab("\nPredicted") +
+  xlab("Actual") +
+  ylab("Predicted") +
   scale_size_area() +

### 2018-05-27 index.rmd
close to ready to go

@@
 census <- read_csv(file = "./input/acsExtract.csv")
+
+#load demo extract
+#demo <- read_csv(file = "./input/tractDemoSumStats.csv")
+
+#load culture extract
+#culture <- read_csv(file = "./input/tractCultureSumStats.csv")
+
+#load police report extract
+#police <- read_csv(file = "./input/tractPoliceReportSumStats.csv")
 
@@
 
+#panel <- left_join(panel, demo)
+#panel <- left_join(panel, culture)
+#panel <- left_join(panel, police)
+#panel <- left_join(panel, sea_shp@data)
+
 sea_df <- panel %>% #set to NA for INLA to forecast
@@
          med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B)) %>%
-  dplyr::select(GISJOIN, listingQtr, Qtr, medRent, med1B, lagRent, actualRent, nHU,
-                n1B, nListings, pOwnoccHU, medHUVal, idtract) %>%
   mutate(p1B = n1B/nListings) %>%
@@
 sea_df$idtract1 <- sea_df$idtract
+sea_df$Qtr1 <- sea_df$Qtr
 
@@
 form.bym <- log(med1B) ~ 1 + Qtr +
+  f(idqtr, model = "iid") +
   f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
@@
     graph = "./output/seatract.graph") +
-  f(idqtr, model = "ar1") +
-  f(idqtr1, model = "iid") +
+  f(idqtr, model = "ar", order = 1) +
   f(idtractqtr, model = "iid")
@@
          spt_cuts = cut(spt_Med, 6))
-
-sea_df
 

### 2018-05-27 smoothSpatiotemporal.R

@@ -1,487 +1,0 @@
-#### INLA Rent Maps for CL ----------------------------------------------------
-#### CSSS 512
-
-#### Preamble -----------------------------------------------------------------
-
-#packages
-library(tidyverse)
-library(forcats)
-library(haven)
-library(lubridate)
-library(haven)
-library(zoo)
-library(ggthemes)
-library(RColorBrewer)
-library(sp)
-library(spdplyr)
-library(geosphere)
-library(rgdal)
-library(sqldf)
-library(spdep)
-library(viridis)
-library(latex2exp)
-
-#setwd to location of file (REQUIRES RSTUDIO)
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
-
-#load ACS data extract (codebook in folder)
-census <- read_csv(file = "./input/acsExtract.csv")
-
-#read in tract shapefiles for seattle
-sea_shp <- readOGR(dsn = "R:/Project/seattle_rental_market/data/geo/sea_tract_2010/sea_tract_2010.shp",
-                  layer = "sea_tract_2010",
-                  GDAL1_integer64_policy = TRUE,
-                  stringsAsFactors = F)
-
-#determine config (dev or github)
-if(file.exists("../data/cl/craigslistDB.sqlite")){
-  #connect to database if file.exists
-  DB <- dbConnect(SQLite(), dbname="../data/cl/craigslistDB.sqlite")
-  cl <- tbl(DB, "clean") #create dbi for clean listing table
-  
-  #compute tract aggregates for CL listing count
-  tract <- cl %>%
-    collect %>% #bring db query into memory
-    filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
-           GISJOIN %in% sea_shp@data$GISJOIN) %>% #only listings with valid Bed/Rent, seattle tracts
-    distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
-             matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
-    dplyr::select(listingDate, GISJOIN, seattle, matchAddress, matchType, 
-                  cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
-    mutate(listingDate = as.Date(listingDate),
[…]

### 2018-05-29 index.rmd
presentation version

@@
 sea_df$idqtr1 <- sea_df$idqtr
-sea_df$idtractqtr <- paste(sea_df$idtract, sea_df$idqtr)
+sea_df$idtractqtr <- paste(sea_df$idtract, sea_df$Qtr)
 sea_df$idtract1 <- sea_df$idtract
@@
 sea_df$int_postWidth <- m.int$summary.linear.predictor[, "0.975quant"] - m.int$summary.linear.predictor[,"0.025quant"]
-sea_df$int_Eff <- m.int$summary.random$idtract[sea_df$idtract, "0.5quant"]
 
@@
   f(idtract, model = "iid") +
-  f(idqtr, model = "ar1") + f(idqtr1, model = "iid")
+  f(idqtr, model = "ar1") +
+  f(idqtr1, model = "iid")
 
@@
 sea_df$nsar1_postWidth <- m.nsar1$summary.linear.predictor[, "0.975quant"] - m.nsar1$summary.linear.predictor[,"0.025quant"]
-sea_df$nsar1_Eff <- m.nsar1$summary.random$idqtr[sea_df$idqtr, "0.5quant"]
+sea_df$nsar1_Eff <- m.nsar1$summary.random$idtract[sea_df$idtract, "0.5quant"]
 
-#spatial random effect model (BYM 1991) with seasonal dummies
-form.bym <- log(med1B) ~ 1 + Qtr +
-  f(idqtr, model = "iid") +
+#spatial random effect model (BYM 1991) with AR(1)
+form.bym <- log(med1B) ~ 1 +
   f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
     scale.model = T,
-    graph = "./output/seatract.graph")
+    graph = "./output/seatract.graph") +
+  f(idqtr, model = "ar1") +
+  f(idqtr1, model = "iid")
 
@@
 
-#spatio-temporal smoothing model, AR(1) process with linear space-time interaction (Bernadelli 1995)
+#spatio-temporal smoothing model, with AR(1) and linear space-time interaction (Bernadelli 1995)
 form.spt <- log(med1B) ~ 1 +
@@
     graph = "./output/seatract.graph") +
-  f(idqtr, model = "ar", order = 1) +
+  f(idqtr, model = "ar1") +
+  f(idqtr1, model = "iid") +
   f(idtractqtr, model = "iid")
@@
 temporal and spatial).
-All of the candidate models are estimated with [integrated nested Laplace approximations](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) (INLA) using the default, weakly-informative priors for all model hyperparameters.
+All of the candidate models' posterior distributions are estimated with [integrated nested Laplace approximations](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) (INLA) using the default, weakly-informative priors for all model hyperparameters.
 Throughout the following analyses, the training data are `r min(sea_df$listingQtr)` up to the prior quarter (i.e.
@@
          qtrDev = value-mean(value, na.rm=T),
-         model = "C. Non-Spatial RE") %>%
+         model = "C. Non-Spatial") %>%
   dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
@@
          qtrDev = value-mean(value, na.rm=T),
-         model = "D. Non-Spatial AR(1) RE") %>%
+         model = "D. Non-Spatial AR(1)") %>%
   dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
@@
          qtrDev = value-mean(value, na.rm=T),
-         model = "E. Spatial RE") %>%
+         model = "E. Spatial AR(1)") %>%
   dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
@@
 
-### Spatial RE
+### Spatial AR(1)
 
@@
 
-### Spatiotemporal AR(1) RE
+### Spatiotemporal AR(1)
 
@@
            | `r format(form.ns)`
-**nsar1**  | Non-spatial tract random effect model, AR(1) random effect for quarter, i.i.d. random effect for quarter 
+**nsar1**  | Non-spatial tract random effect, AR(1) random effect for prior quarter, i.i.d random effect for current quarter
            | `r format(form.nsar1)`
-**bym**    | Spatial intrinsic conditional autoregressive (ICAR) tract random-effect with quarter fixed intercept 
+**bym**    | Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, non-spatial i.i.d tract random effect, AR(1) random effect for prior quarter, i.i.d random effect for current quarter
            |  `r format(form.bym)`
-**spt**    | Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, AR(1) random effect for quarter, i.i.d. random effect for quarter and tract-quarter
+**spt**    | Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, non-spatial i.i.d tract random effect, AR(1) random effect for prior quarter, i.i.d random effect for current quarter, i.i.d.
+random effect for tract-quarter (space-time interaction)
            | `r format(form.spt)`
@@
 
+### Hyperparameters
+
+```{r, echo=F}
+knitr::kable(summary(m.nsar1)$hyperpar, caption = "Non-Spatial AR(1)")
+```
+
+<br>
+<hr>
+<br>
+
+```{r, echo=F}
+knitr::kable(summary(m.bym)$hyperpar, caption = "Spatial AR(1)")
+```
+
+<br>
+<hr>
+<br>
+
+```{r, echo=F}
+knitr::kable(summary(m.spt)$hyperpar, caption = "Spatiotemporal AR(1)")
+```
+
 Column {.tabset}
@@
   geom_abline(intercept = 0, slope = 1) +
-  scale_x_continuous(labels = scales::dollar) +
-  scale_y_continuous(labels = scales::dollar) +
+  scale_x_continuous(limits = c(800, 2600), labels = scales::dollar) +
+  scale_y_continuous(limits = c(800, 2600), labels = scales::dollar) +
   xlab("Actual") +

### 2018-05-29 index.rmd
updated version for presentation, fixed spt to follow INLA google group's discussion about obs-specific REs

@@
               control.predictor = list(compute = TRUE),
-              control.compute = list(dic = TRUE, waic = TRUE))
+              control.compute = list(dic = TRUE, waic = TRUE),
+              control.inla = list(correct = TRUE, correct.factor = 10))
 

### 2018-05-30 index.rmd
updating to reflect bym as top model

@@
 pred <- sea_f %>%
-  mutate(rent = spt_Med) %>%
+  mutate(rent = bym_Med) %>%
   filter(listingQtr == as.yearqtr(Sys.Date())) %>%
   mutate(var = paste0(as.yearqtr(Sys.Date()), " Forecast")) %>%
-  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, spt_Med, NAMELSAD10, n1B)
+  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, bym_Med, NAMELSAD10, n1B)
 
 comb <- bind_rows(obs, pred) %>%
-  mutate(tooltip = paste0(NAMELSAD10, "<br>", "Forecast: $", round(spt_Med), "<br>", "Qtr-to-Date: $", round(actualRent), "<br>", "N Qtr-to-Date: ", n1B))
+  mutate(tooltip = paste0(NAMELSAD10, "<br>", "Forecast: $", round(bym_Med), "<br>", "Qtr-to-Date: $", round(actualRent), "<br>", "N Qtr-to-Date: ", n1B))
 comb <- SharedData$new(comb, ~GISJOIN)
@@
 **int**    | Quarter fixed intercept
-           | `r format(form.int, )`
+           | `r format(form.int)`
 **ns**     | Non-spatial tract random effect for each tract, quarter fixed intercept

### 2018-06-02 index.rmd
working on manuscript

@@
     navbar:
-      - { title: "hesscl.com", href: "http://hesscl.com", align: right }
+      - { title: "hesscl.com", href: "http://hesscl.com", align: right ,
+          title: "Manuscript", href: "http://github.com/hesscl/smooth-sea/smooth-sea.html", align: right}
     orientation: columns
@@
 
+#source data_load()
+source("helpers/load_data.R")
+
 #load ACS data extract (codebook in folder)
 census <- read_csv(file = "./input/acsExtract.csv")
-
-#load demo extract
-#demo <- read_csv(file = "./input/tractDemoSumStats.csv")
-
-#load culture extract
-#culture <- read_csv(file = "./input/tractCultureSumStats.csv")
-
-#load police report extract
-#police <- read_csv(file = "./input/tractPoliceReportSumStats.csv")
+censusRent <- read_csv(file = "./input/acsRent.csv")
+census <- inner_join(census, censusRent)
 
@@
 
-#determine config (dev or github)
-if(file.exists("../data/cl/craigslistDB.sqlite")){
-  #connect to database if file.exists
-  DB <- dbConnect(SQLite(), dbname="../data/cl/craigslistDB.sqlite")
-  cl <- tbl(DB, "clean") #create dbi for clean listing table
-
-  #compute tract aggregates for CL listing count
-  tract <- cl %>%
-    collect %>% #bring db query into memory
-    filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
-           GISJOIN %in% sea_shp@data$GISJOIN) %>% #only listings with valid Bed/Rent, seattle tracts
-    distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
-             matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
-    dplyr::select(listingDate, GISJOIN, seattle, matchAddress, matchType, 
-                  cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
-    mutate(listingDate = as.Date(listingDate),
-           listingQtr = as.yearqtr(listingDate)) %>%
-    filter(cleanBeds %in% c(0, 1, 2, 3), listingQtr >= "2017 Q2") %>%
-    group_by(GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
-    summarize(nListings = n(),
-              n1B = sum(cleanBeds == 1),
-              medRent = median(cleanRent),
-              med0B = median(cleanRent[cleanBeds==0]),
-              med1B = median(cleanRent[cleanBeds==1]),
-              med2B = median(cleanRent[cleanBeds==2]),
-              med3B = median(cleanRent[cleanBeds==3]),
-              lagRent = lag(cleanRent)) %>% #create tract aggregates
-    ungroup %>%
-    arrange(GISJOIN, listingQtr)
-  dbDisconnect(DB)
-  write_csv(tract, "./input/tractCl.csv")
-
-} else{
-  #if github config, read in extract from cl db
-  tract <- read_csv("./input/tractCl.csv")
-}
+#use sourced function to load listing summaries
+tract <- load_data()
 
@@
 
-#panel <- left_join(panel, demo)
-#panel <- left_join(panel, culture)
-#panel <- left_join(panel, police)
-#panel <- left_join(panel, sea_shp@data)
-
-sea_df <- panel %>% #set to NA for INLA to forecast
-  mutate(actualRent = med1B,
-         med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B)) %>%
-  mutate(p1B = n1B/nListings) %>%
-  arrange(listingQtr, GISJOIN)
+#establish object for modeling based off of adjusting `panel`
+sea_df <- panel %>% 
+  mutate(actualRent = med1B, #first preserve the observed / observed-to-date
+         med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B)) %>% #set to NA for INLA to forecast current qtr
+  mutate(p1B = n1B/nListings) %>% #compute share of 1B listings
+  arrange(listingQtr, GISJOIN) #arange qtr, tract
 
@@
 
-This website provides interactive results for the forecasting models explored in the paper _Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_.
+This website provides interactive results for the forecasting models explored in the paper [_Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_](http://github.com/hesscl/smooth-sea/smooth-sea.html).
 

### 2018-06-02 smooth-sea.Rmd
working on manuscript

--- 2018-06-02 index.rmd
@@
 ---
-title: "smooth-sea"
-output: 
-  flexdashboard::flex_dashboard:
-    navbar:
-      - { title: "hesscl.com", href: "http://hesscl.com", align: right ,
-          title: "Manuscript", href: "http://github.com/hesscl/smooth-sea/smooth-sea.html", align: right}
-    orientation: columns
-    theme: bootstrap
-    logo: logo.svg
-    source_code: "https://github.com/hesscl/smooth-sea"
+title: "Estimating Neighborhood Rents with Scraped Web Data"
+author: "Chris Hess"
+date: "June 2, 2018"
+output: html_document
+abstract: "Online classified ads are a key source of information about local rental housing markets.
+This research asks what temporal and spatial dynamics explain neighborhood variations in one-bedroom rents in Seattle, WA, and contributes a novel approach to forecasting quarterly neighborhood rent levels using Bayesian models estimated with integrated nested Laplace approximations (INLA).
+All models were trained on data for 2017 Q2 through 2018 Q1, with forecasts for 2018 Q2 compared against the observed-to-date data for the current quarter (i.e.
+2018 Q2).
+The top model assumes that spatial and temporal dynamics are separable, i.e.
+independent of each other, and non-interactive.
+This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $90 for the training data and $145 for the test data."
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-
 #packages
@@
 library(latex2exp)
-library(plotly)
-library(DT)
-library(crosstalk)
-
-#source data_load()
-source("helpers/load_data.R")
 
@@
 
+#go ahead and save separate versions for graphics later
+tract2000 <- readOGR(dsn = "./input/sea_tract_2000/sea_tract_2000.shp",
+                  layer = "sea_tract_2000",
+                  GDAL1_integer64_policy = TRUE,
+                  stringsAsFactors = F) 
+tract2010 <- sea_shp
+
+#source data_load()
+source("./helpers/load_data.R")
+
 #use sourced function to load listing summaries
 tract <- load_data()
+```
 
-#### A. construct panel (some missingness) ------------------------------------
+<br>
+<br>
 
+Demand for rental housing units in Seattle, WA surged in the past decade, due to relative scarcity in available housing amidst a period of historic population growth.
+Growing shares of households are unable to buy a home short of "driving till they qualify", and among households who rent a substantial share pay 30%, if not more of their incomes on their housing costs.
+The rising cost of housing therefore has a direct role in neighborhood processes of "demographic inversion" like urban gentrification and suburbanizing poverty.
+Census data (i.e.
+decennial and American Community Survey) have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more.
+This publication schedules prohibit analyzing any market dynamics beyond long-term trends, particularly among smaller areas of aggregation like neighborhoods.
+While more up-to-date options for neighborhood rent data exist through privately-held companies, such data can be somewhat of a black box in their creation and may be biased towards the costs of corporate-held properties.
+This would-be problem for academic researchers creates an opportunity for applying web scraping methods to generate a novel data on housing availability and costs.
+
+This study aims to demonstrate why scraped rental housing listings are a valuable source of data in subject areas such as housing studies and neighborhood demography through a set of descriptive analyses and models of quarterly rent levels among census tracts in Seattle, WA.
+Summaries of the Craigslist data show agreement with existing sources in their distribution but do suggest higher median levels for most neighborhoods, likely given the listings' recent collection relative to other data.
+In a more general sense, the methods and frameworks used in this study also offer a template for efficiently collecting web data, conducting reproducible research, and publishing interactive results to GitHub.
+I start by briefly reviewing some background relevant to studying neighborhood rental housing.
+I then outline a process for automated web scraping of a rental listing source like Craigslist using a program called Helena, and compare aggregates from the collected rental listings with other data to assess rent agreement in measures.
+After describing the data and process, I present a set of models predicting neighborhood rent levels to understand different how different temporal and spatial structures fit the data.
+I conclude by discussing the promise of web data for subfields like housing studies and the value of R Markdown and GitHub for social science research.
+
+<br>
+
+## Background
+
+Existing public data sources like the American Community Survey or decennial census theoretically have two sources of error relative to the ``true'' median rent in a given overall market and to an even greater degree among neighborhood approximations like census tracts.
+Census sources of rental information cannot provide insight into any change in rental cost associated with newer developments becoming available (at least not until considerable time has already passed) given the temporal delay in data publication (i.e.
+ACS estimates are usually at least 2 years old, decennial many more).
+This feature of Census data is salient given newer housing units' generally higher rents and greater-than-average vacancy, both factors which lead to underestimating the present levels of rents for available housing units when either the ACS or decennial census are used as a current indicator.
+For already-leased rental housing units that are sampled, the temporal lag generates measurement error such that data on a given building's current asking rents may be underestimated relative to current tenants' payment levels.
+
+For-profit data sources play a role in filling the information gap about rental housing costs, though they tend to possess their own unique shortcomings.
+These data sources tend to rely on survey methodologies for data collection, and ask landlords in a given market to outline the average cost for units of a given bedroom size or other specification.
+In exchange for providing detailed information for their database, the rental data provider may give participating landlords and property management companies access to a slice of information generated via accumulated data.
+These data tend to be rather opaque in their collection methodologies, with larger or corporate-owner properties most likely to appear in these data relative to independently-owned (i.e.
+mom and pop) rental units.
+
+<br>
+
+## Data
+
+```{r map_point_density, echo=FALSE, message=FALSE, warning=FALSE}
+#determine config (dev or github)
+if(file.exists("../data/cl/craigslistDB.sqlite")){
+  source("./helpers/map_point_density.R") #if config, refresh plot pdf
+}
+```
+
+![Figure 1. Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](./output/map_point_density.png)
+<br>
+
+### Web scraping with Helena
+
+<br>
+
+### Data Processing
+
+<br>
+
+### Data Description
+
+```{r rent_lineplot, echo=FALSE, message=FALSE, warning=FALSE}
+#determine config (dev or github)
+if(file.exists("../data/cl/craigslistDB.sqlite")){
+  source("./helpers/rent_lineplot.R") #if config, refresh plot pdf
+}
+```
+
+![Figure 2.
+Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018.
+Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual and ACS are annual (i.e.
+respective 1-Year Estimates for each year period).](./output/rent_lineplot.png)
+
+```{r rent_map_across, echo=FALSE, message=FALSE, warning=FALSE}
+#determine config (dev or github)
+if(file.exists("../data/cl/craigslistDB.sqlite")){
+  source("./helpers/rent_map_across.R") #if config, refresh plot pdf
+}
+```
+
+![Figure 3.
+Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates.
+Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled.](./output/rent_map_across_1B.png)
+
+```{r rent_scatter_across, echo=FALSE, message=FALSE, warning=FALSE}
+#determine config (dev or github)
+if(file.exists("../data/cl/craigslistDB.sqlite")){
+  source("./helpers/rent_scatter_across.R") #if config, refresh plot pdf
+}
+```
+
+![Figure 4.
+Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott.
+Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.](./output/rent_scatter_across_1B.png)
+<br>
+
+## Models
+
+```{r construct panel (some missingness), echo=FALSE, message=FALSE, warning=FALSE}
 #ensure order of census df matches shapefile
@@
   arrange(listingQtr, GISJOIN) #arange qtr, tract
+```
 
-#### B. INLA Models of CL Rent ------------------------------------------------
-
+```{r run models, echo=FALSE, message=FALSE, warning=FALSE}
 library(INLA)
@@
 
-Home {data-icon="fa-home"}
-=====================================
-
-Column
--------------------------------------
-
-### Welcome!
-
-This website provides interactive results for the forecasting models explored in the paper [_Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_](http://github.com/hesscl/smooth-sea/smooth-sea.html).
-
-The goal of this research is to a.) understand the temporal dynamics of rent estimates in Seattle and b.) forecast the current quarter's rent levels based off of the prior periods.
-The focal series of models regress median one-bedroom rent asked values on different specifications of the panel's correlation structure (i.e.
-temporal and spatial).
-All of the candidate models' posterior distributions are estimated with [integrated nested Laplace approximations](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) (INLA) using the default, weakly-informative priors for all model hyperparameters.
-Throughout the following analyses, the training data are `r min(sea_df$listingQtr)` up to the prior quarter (i.e.
-`r as.yearqtr(Sys.Date())-.25`).
-The test period is a forecast for the current period and includes comparison to the appropriate median rent estimates for data observed so far in this period.
-
-Most graphics include some level of interactivity, usually either hover-over tooltip information or a slider to control various views of the graphic.
-Clicking on cases will highlight data elements in most graphics, and double-clicking will reset the graphic.
-You can download the source code and data for this project from Github [here](https://github.com/hesscl/smooth-sea).
-
-Contact Chris Hess at [hesscl@uw.edu](mailto://hesscl@uw.edu) for more information about this research.
-
-_This page was last updated: `r Sys.Date()`_
+`r knitr::kable(fit_stats)`
 
 <br>
-<hr>
+
+## Results
+
 <br>
 
-**Table of Contents**
-
-Page                |  Description
-------------------- | ---------------------------------------------------------
-Distribution        | density graphic to investigate the distribution of rents among Seattle neighborhoods for each quarter
-Panel Time-Series   | line graphic to show the observed or modeled temporal structure
-Spatial Time-Series | series of maps to show observed change across time
-Model Fit           | tables of model root mean square error (RMSE), mean absolute error (MAE), and deviance information criterion (DIC) across training and test data
-
-Column
--------------------------------------
-
-### Observed vs. Smoothed Rent Estimates
-
-```{r index.maps, echo=F}
-obs <- sea_f %>%
-  mutate(rent = actualRent) %>%
-  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
-  mutate(var = paste0(as.yearqtr(Sys.Date()), "-to-Date")) %>%
-  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, spt_Med, NAMELSAD10, n1B)
-
-pred <- sea_f %>%
-  mutate(rent = bym_Med) %>%
-  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
-  mutate(var = paste0(as.yearqtr(Sys.Date()), " Forecast")) %>%
-  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, bym_Med, NAMELSAD10, n1B)
-
-comb <- bind_rows(obs, pred) %>%
-  mutate(tooltip = paste0(NAMELSAD10, "<br>", "Forecast: $", round(bym_Med), "<br>", "Qtr-to-Date: $", round(actualRent), "<br>", "N Qtr-to-Date: ", n1B))
-comb <- SharedData$new(comb, ~GISJOIN)
-
-avp_map <- ggplot(comb, aes(x = long, y = lat, group = group, fill = rent, label = tooltip)) +
-  facet_wrap(~ var) +
-  geom_polygon(color = "grey85", lwd = .05) +
-  scale_fill_viridis_c(limits = c(750, 2500), labels = scales::dollar) +
-  coord_quickmap() +
-  theme(axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank(),
-        panel.background = element_blank()) +
-  labs(fill = "Median<br>1B Rent")
-
-ggplotly(avp_map, tooltip = "label") %>%
-  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
-  highlight(on = "plotly_click")
-```
-
-Distribution
-=====================================
-
-```{r, echo=F}
-gg_hist <- ggplot(sea_df, aes(x = med1B, group = listingQtr, color = listingQtr)) +
-  geom_density() +
-  theme_minimal() +
-  scale_x_continuous(labels = scales::dollar) +
-  scale_color_viridis_d(end = .8) +
-  xlab("\nMedian 1B Rent") +
-  labs(color = "Quarter")
-
-ggplotly(gg_hist, tooltip = c("color", "x")) %>%
-  layout(autosize = T)
-```
-
-Panel Time-Series
-=====================================
-
-```{r temp.lineplot, echo=F}
-obs_df <- sea_df %>%
-  group_by(listingQtr) %>%
-  mutate(value = actualRent,
-         qtrDev = value-mean(value, na.rm=T),
-         model = "A. Observed")
-
-int_df <- sea_df %>%
-  group_by(listingQtr) %>%
-  mutate(value = int_Med,
-         qtrDev = value-mean(value, na.rm=T),
-         model = "B. Quarter FE") %>%
-  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
-
-ns_df <- sea_df %>%
-  group_by(listingQtr) %>%
-  mutate(value = ns_Med,
-         qtrDev = value-mean(value, na.rm=T),
-         model = "C. Non-Spatial") %>%
-  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
-
-nsar1_df <- sea_df %>%
-  group_by(listingQtr) %>%
-  mutate(value = nsar1_Med,
-         qtrDev = value-mean(value, na.rm=T),
-         model = "D. Non-Spatial AR(1)") %>%
-  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
-
-bym_df <- sea_df %>%
-  group_by(listingQtr) %>%
-  mutate(value = bym_Med,
-         qtrDev = value-mean(value, na.rm=T),
-         model = "E. Spatial AR(1)") %>%
-  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
-
-spt_df <- sea_df %>%
-  group_by(listingQtr) %>%
-  mutate(value =  spt_Med,
-         qtrDev = value-mean(value, na.rm=T),
-         model = "F. Spatiotemporal AR(1)") %>%
-  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
-
-panel_df <- bind_rows(obs_df, int_df, ns_df, nsar1_df, bym_df, spt_df)
-
-panel_df <- SharedData$new(panel_df, ~GISJOIN)
-
-obs_lineplot <- ggplot(panel_df, aes(x = listingQtr, y = value, group = GISJOIN, label = GISJOIN, color = qtrDev, frame = model)) +
-  geom_line(lwd = .25, color = "black", alpha = .5) +
-  geom_point() +
-  scale_color_gradient2_tableau() +
-  scale_y_continuous(labels = scales::dollar) +
-  xlab("") +
-  ylab("Median 1B Rent\n") +
-  theme_minimal() +
-  labs(color = "Quarterly\ndeviation")
-
-ggplotly(obs_lineplot, tooltip = c("label", "y")) %>%
-  layout(autosize = T) %>%
-  animation_opts(frame = 1000, easing = "linear", redraw = FALSE) %>%
-  animation_button(hide = T) %>%
-  highlight()
-```
-
-Spatial Time-Series
-=====================================
-
-Column {.tabset}
--------------------------------------
-
-### Observed
-
-```{r spat.series.actual, echo=F}
-map_panel_obs <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = actualRent)) +
-  facet_grid(~ listingQtr) +
-  geom_polygon(color = "grey85", lwd = .05) +
-  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
-  coord_quickmap() +
-  theme(axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank(),
-        panel.background = element_blank()) +
-  labs(fill = "Median<br>1B Rent")
-
-ggplotly(map_panel_obs) %>%
-  layout(autosize = T)
-```
-
-### Non-Spatial AR(1)
-
-```{r spat.series.nsar1, echo=F}
-map_panel_nsar1 <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = nsar1_Med, frame = listingQtr)) +
-  geom_polygon(color = "grey85", lwd = .05) +
-  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
-  coord_quickmap() +
-  theme(axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank(),
-        panel.background = element_blank()) +
-  labs(fill = "Median<br>1B Rent")
-
-ggplotly(map_panel_nsar1) %>%
-  layout(autosize = T) %>%
-  animation_opts(1000, transition = 0, redraw = FALSE)
-```
-
-### Spatial AR(1)
-
-```{r spat.series.bym, echo=F}
-map_panel_bym <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = bym_Med, frame = listingQtr)) +
-  geom_polygon(color = "grey85", lwd = .05) +
-  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
-  coord_quickmap() +
-  theme(axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank(),
-        panel.background = element_blank()) +
-  labs(fill = "Median<br>1B Rent")
-
-ggplotly(map_panel_bym) %>%
-  layout(autosize = T) %>%
-  animation_opts(1000, transition = 0, redraw = FALSE)
-```
-
-### Spatiotemporal AR(1)
-
-```{r spat.series.spt, echo=F}
-map_panel_spt <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = spt_Med, frame = listingQtr)) +
-  geom_polygon(color = "grey85", lwd = .05) +
-  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
-  coord_quickmap() +
-  theme(axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank(),
-        panel.background = element_blank()) +
-  labs(fill = "Median<br>1B Rent")
-
-ggplotly(map_panel_spt) %>%
-  layout(autosize = T) %>%
-  animation_opts(1000, transition = 0, redraw = FALSE)
-```
-
-Model Fit
-=====================================
-
-Column {.tabset}
--------------------------------------
-
-### Model Legend
-
-Model abbr.                   | Description
------------------------------ | -----------------------------------------------
-**int**    | Quarter fixed intercept
-           | `r format(form.int)`
-**ns**     | Non-spatial tract random effect for each tract, quarter fixed intercept
-           | `r format(form.ns)`
-**nsar1**  | Non-spatial tract random effect, AR(1) random effect for prior quarter, i.i.d random effect for current quarter
-           | `r format(form.nsar1)`
-**bym**    | Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, non-spatial i.i.d tract random effect, AR(1) random effect for prior quarter, i.i.d random effect for current quarter
-           |  `r format(form.bym)`
-**spt**    | Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, non-spatial i.i.d tract random effect, AR(1) random effect for prior quarter, i.i.d random effect for current quarter, i.i.d.
-random effect for tract-quarter (space-time interaction)
-           | `r format(form.spt)`
-
-### Accuracy and Information Criteria
-
-```{r, echo=F}
-knitr::kable(fit_stats %>% dplyr::select(train_test, ends_with("rmse")))
-```
+## Discussion
 
 <br>
-<hr>
-<br>
 
-```{r, echo=F}
-knitr::kable(fit_stats %>% dplyr::select(train_test, ends_with("mae")))
-```
-
-<br>
-<hr>
-<br>
-
-```{r, echo=F}
-knitr::kable(data.frame("train_test" = "Training",
-                        "int_DIC" = m.int$dic$dic,
-                        "ns_DIC" = m.ns$dic$dic,
-                        "nsar1_DIC" = m.nsar1$dic$dic,
-                        "bym_DIC" = m.bym$dic$dic,
-                        "spt_DIC" = m.spt$dic$dic))
-```
-
-<br>
-<hr>
-<br>
-
-```{r, echo=F}
-knitr::kable(data.frame("train_test" = "Training",
-                        "int_WAIC" = m.int$waic$waic,
-                        "ns_WAIC" = m.ns$waic$waic,
-                        "nsar1_WAIC" = m.nsar1$waic$waic,
-                        "bym_WAIC" = m.bym$waic$waic,
-                        "spt_WAIC" = m.spt$waic$waic))
-```
-
-### Hyperparameters
-
-```{r, echo=F}
-knitr::kable(summary(m.nsar1)$hyperpar, caption = "Non-Spatial AR(1)")
-```
-
-<br>
-<hr>
-<br>
-
-```{r, echo=F}
-knitr::kable(summary(m.bym)$hyperpar, caption = "Spatial AR(1)")
-```
-
-<br>
-<hr>
-<br>
-
-```{r, echo=F}
-knitr::kable(summary(m.spt)$hyperpar, caption = "Spatiotemporal AR(1)")
-```
-
-Column {.tabset}
--------------------------------------
-
-### Non-Spatial AR(1)
-
-```{r, echo=F}
-avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = nsar1_Med, size = n1B, frame = listingQtr)) +
-  geom_point() +
-  geom_abline(intercept = 0, slope = 1) +
-  scale_x_continuous(limits = c(800, 2600), labels = scales::dollar) +
-  scale_y_continuous(limits = c(800, 2600), labels = scales::dollar) +
-  xlab("Actual") +
-  ylab("Predicted") +
-  scale_size_area() +
-  coord_equal() +
-  theme_minimal()
-
-ggplotly(avp_scatter) %>%
-  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
-  animation_opts(1000, transition = 0, redraw = T)
-```
-
-### Spatial AR(1)
-
-```{r, echo=F}
-avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = bym_Med, size = n1B, frame = listingQtr)) +
-  geom_point() +
-  geom_abline(intercept = 0, slope = 1) +
-  scale_x_continuous(limits = c(800, 2600), labels = scales::dollar) +
-  scale_y_continuous(limits = c(800, 2600), labels = scales::dollar) +
-  xlab("Actual") +
-  ylab("Predicted") +
-  scale_size_area() +
-  coord_equal() +
-  theme_minimal()
-
-ggplotly(avp_scatter) %>%
-  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
-  animation_opts(1000, transition = 0, redraw = FALSE)
-```
-
-### Spatiotemporal AR(1)
-
-```{r, echo=F}
-avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = spt_Med, size = n1B, frame = listingQtr)) +
-  geom_point() +
-  geom_abline(intercept = 0, slope = 1) +
-  scale_x_continuous(limits = c(800, 2600), labels = scales::dollar) +
-  scale_y_continuous(limits = c(800, 2600), labels = scales::dollar) +
-  xlab("Actual") +
-  ylab("Predicted") +
-  scale_size_area() +
-  coord_equal() +
-  theme_minimal()
-
-ggplotly(avp_scatter) %>%
-  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
-  animation_opts(1000, transition = 0, redraw = FALSE)
-```
+## References

### 2018-06-02 helpers/load_data.R

@@ -1,0 +1,39 @@
+load_data <- function(){
+  #determine config (dev or github)
+  if(file.exists("../data/cl/craigslistDB.sqlite")){
+    #connect to database if file.exists
+    DB <- dbConnect(SQLite(), dbname="../data/cl/craigslistDB.sqlite")
+    cl <- tbl(DB, "clean") #create dbi for clean listing table
+    
+    #compute tract aggregates for CL listing count
+    tract <- cl %>%
+      collect %>% #bring db query into memory
+      filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
+             GISJOIN %in% sea_shp@data$GISJOIN, #only listings with valid Bed/Rent, seattle tracts
+             matchType != "Google Maps Lat/Long") %>% #need to have address, not approximate
+      distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
+               matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
+      dplyr::select(listingDate, GISJOIN, seattle, matchAddress, matchType, 
+                    cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
+      mutate(listingDate = as.Date(listingDate),
+             listingQtr = as.yearqtr(listingDate)) %>%
+      filter(cleanBeds %in% c(0, 1, 2, 3), listingQtr >= "2017 Q2") %>%
+      group_by(GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
+      summarize(nListings = n(),
+                n1B = sum(cleanBeds == 1),
+                medRent = median(cleanRent),
+                med0B = median(cleanRent[cleanBeds==0]),
+                med1B = median(cleanRent[cleanBeds==1]),
+                med2B = median(cleanRent[cleanBeds==2]),
+                med3B = median(cleanRent[cleanBeds==3])) %>% #create tract aggregates
+      ungroup %>%
+      arrange(GISJOIN, listingQtr)
+    dbDisconnect(DB)
+    write_csv(tract, "./input/tractCl.csv")
+    return(tract)
+  } else{
+    #if github config, read in extract from cl db
+    tract <- read_csv("./input/tractCl.csv")
+    return(tract)
+  }
+}

### 2018-06-02 helpers/map_point_density.R

@@ -1,0 +1,45 @@
+
+#### Helper script to compute map + point density map
+
+#libraries
+library(ggmap)
+library(patchwork)
+
+#data
+DB <- dbConnect(SQLite(), dbname="R:/Project/seattle_rental_market/data/cl/craigslistDB.sqlite")
+map_clean <- tbl(DB, "clean") #clean seattle listing table
+map_raw <- tbl(DB, "raw")
+
+#load table
+map_cl <- map_clean %>% 
+  filter(seattle==1) %>%
+  collect %>% #bring db query into memory
+  filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
+         GISJOIN %in% sea_shp@data$GISJOIN, matchType != "Google Maps Lat/Long") %>% #only listings with valid Bed/Rent, seattle tracts
+  distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
+           matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
+  dplyr::select(listingDate, GISJOIN, seattle, matchAddress, matchType, 
+                cleanBeds, cleanRent, cleanSqft, lat, lng) %>% #SELECT these columns
+  mutate(listingDate = as.Date(listingDate),
+         listingQtr = as.yearqtr(listingDate)) %>%
+  filter(cleanBeds %in% c(1), listingQtr >= "2017 Q2")
+dbDisconnect(DB)
+
+blank <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", geom = "blank") 
+
+with_points <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", darken = .50, legend = "bottom") +
+  stat_density_2d(aes(fill = ..level..), geom = "polygon",
+                  alpha = .75, color = NA, size = .05) +
+  scale_fill_viridis_c("1B Listings") +
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        legend.text = element_text(angle = 45, vjust = .8),
+        panel.background = element_blank(),
+        strip.background = element_blank())
+
+blank + with_points +
+  ggsave(filename = "./output/map_point_density.png",
+         width = 6, height = 6, units = "in", dpi = 300)
+
+

### 2018-06-02 helpers/rent_lineplot.R

@@ -1,0 +1,115 @@
+
+### Helper script to produce across data lineplot of median rent for Seattle
+
+#data
+DB <- dbConnect(SQLite(), dbname="R:/Project/seattle_rental_market/data/cl/craigslistDB.sqlite")
+sea <- tbl(DB, "clean") #clean seattle listing table
+raw <- tbl(DB, "raw")
+
+#for code that has not been updated to db calls
+cl <- sea %>% collect %>% filter(seattle==1)
+
+#load Seattle market summaries from D+S
+ds <- read_csv(file = "R:/Project/seattle_rental_market/data/d+s/city-level D+S.csv")
+
+
+#### A. Prepare a table for ggplot ---------------------------------------------
+
+sumCL <- cl %>%
+  #mutate(cleanRent = ifelse(cleanRent > 5000, NA, cleanRent)) %>%
+  mutate(catSqft = fct_reorder(catSqft, cleanSqft)) %>%
+  arrange(listingDate) %>%
+  select(-ends_with("10")) %>%
+  filter(!is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), !is.na(matchAddress), !is.na(matchAddress2)) %>%
+  distinct(matchAddress, matchAddress2, catBeds, cleanSqft, cleanRent, .keep_all = T) %>%
+  mutate(listingQtr = as.yearqtr(as.Date(listingDate))) %>%
+  group_by(listingQtr) %>%
+  summarize(Rent = median(cleanRent, na.rm=T),
+            nHU = n()) %>%
+  rename(moYr = listingQtr) %>%
+  filter(nHU >= 200) %>%
+  mutate(dataSrc = "Craigslist") %>%
+  select(moYr, Rent, dataSrc) %>%
+  mutate(moYr = as.Date(moYr))
+
+sumDS <- ds %>%
+  filter(rowType == "All HUs") %>%
+  mutate(surveyYear = parse_number(Survey),
+         surveySeason = sub("[:0-9:]+", "", Survey),
+         surveyMonth = ifelse(surveySeason == "Spring ", 3,
+                              ifelse(surveySeason == "Fall ", 9, NA)),
+         surveyDate = as.yearmon(paste0(surveyYear, "-0", surveyMonth, "-01"))) %>% #spring = March, Fall = Sept
+  select(-Survey, -surveySeason) %>%
[…]

### 2018-06-02 helpers/rent_map_across.R

@@ -1,0 +1,152 @@
+#### Helper secript to compute across data chloropleth maps
+
+#load census data
+rent_map_acs <- read_csv(file = "R:/Project/seattle_rental_market/data/census/acs/nhgis0051_ds226_20165_2016_tract.csv")
+
+#connect to craigslist data
+DB <- dbConnect(SQLite(), dbname="R:/Project/seattle_rental_market/data/cl/craigslistDB.sqlite")
+rent_map_sea <- tbl(DB, "clean") #clean seattle listing table
+rent_map_raw <- tbl(DB, "raw")
+
+#load d+s data
+rent_map_ds <- read_csv("R:/Project/seattle_rental_market/data/d+s/geocoded_D+S.csv")
+
+#for background
+tract2010_bg <- fortify(tract2010)
+
+#### A. Prep CL table ---------------------------------------------------------
+
+#2017-2018 listings in seattle with non-missing lat/lng, drop 2010 geog fields
+cl2017 <- rent_map_sea %>% 
+  filter(seattle==1) %>% 
+  collect %>% 
+  filter(listingYear >= 2017, 
+         !is.na(matchAddress), !is.na(matchAddress2), !is.na(catBeds), !is.na(cleanSqft), !is.na(cleanRent),
+         catBeds %in% c("0", "1", "2"), !is.na(lat) & !is.na(lng)) %>%
+  distinct(matchAddress, matchAddress2, catBeds, cleanSqft, cleanRent, .keep_all = T) %>%
+  group_by(GISJOIN, catBeds) %>%
+  summarize(rent = median(cleanRent, na.rm = T),
+            nHU = n(),
+            data = "Craigslist 2017-2018") %>%
+  select(GISJOIN, data, catBeds, rent, nHU)
+dbDisconnect(DB)
+
+#### B. Prep D+S table --------------------------------------------------------
+
+#munge starting table of tract summaries
+ds2017 <- rent_map_ds %>%
+  select(-STATEFP00, -COUNTYFP00, -TRACTCE00, -CTIDFP00, -SHAPE_AREA, -SHAPE_LEN, -NAME00,
+         -INTPTLAT00, -INTPTLON00) %>%
+  mutate(surveyYear = parse_number(Survey),
+         surveySeason = sub("[:0-9:]+", "", Survey),
+         surveyMonth = ifelse(surveySeason == "Spring ", 3,
[…]

### 2018-06-02 helpers/rent_scatter_across.R

@@ -1,0 +1,78 @@
+#### Helper script to compute scatter plot of CL X D+S
+
+#data
+DB <- dbConnect(SQLite(), dbname="R:/Project/seattle_rental_market/data/cl/craigslistDB.sqlite")
+scatter_sea <- tbl(DB, "clean") #clean seattle listing table
+scatter_raw <- tbl(DB, "raw")
+
+#read in D+S tract summaries
+scatter_ds <- read_csv("R:/Project/seattle_rental_market/data/d+s/geocoded_D+S.csv")
+
+#### A. Prepare a table for ggplot ---------------------------------------------
+
+scatter_cl <- scatter_sea %>% 
+  filter(seattle==1) %>%
+  collect %>% #bring db query into memory
+  filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
+         GISJOIN %in% sea_shp@data$GISJOIN, matchType != "Google Maps Lat/Long") %>% #only listings with valid Bed/Rent, seattle tracts
+  distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
+           matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
+  dplyr::select(listingYear, listingDate, GISJOIN, seattle, matchAddress, matchType, 
+                cleanBeds, cleanRent, cleanSqft, lat, lng) %>% #SELECT these columns
+  mutate(listingDate = as.Date(listingDate),
+         listingQtr = as.yearqtr(listingDate)) %>%
+  filter(cleanBeds %in% c(1), listingQtr >= "2017 Q2")
+
+coordinates(scatter_cl) <- cbind(scatter_cl$lng, scatter_cl$lat)
+
+#make sure shapefile has correct projection
+WGS.latlon <- CRS("+proj=longlat +datum=WGS84")
+tract2000 <- readOGR(dsn = "./input/sea_tract_2000/sea_tract_2000.shp",
+                     layer = "sea_tract_2000",
+                     GDAL1_integer64_policy = TRUE,
+                     stringsAsFactors = F) 
+tract2000 <- spTransform(tract2000, WGS.latlon)
+proj4string(scatter_cl) <- WGS.latlon
+
+#append 2000 geography to listings
[…]

### 2018-06-03 index.rmd
trying sub-pages for html files

@@
     navbar:
-      - { title: "hesscl.com", href: "http://hesscl.com", align: right ,
-          title: "Manuscript", href: "http://github.com/hesscl/smooth-sea/smooth-sea.html", align: right}
+       - {title: "hesscl.com", href: "http://hesscl.com", align: right} 
     orientation: columns
@@
 
-This website provides interactive results for the forecasting models explored in the paper [_Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_](http://github.com/hesscl/smooth-sea/smooth-sea.html).
+This website provides interactive results for the forecasting models explored in the paper _Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_.
 

### 2018-06-03 helpers/rent_scatter_across.R

@@ -30,6 +30,7 @@ WGS.latlon <- CRS("+proj=longlat +datum=WGS84")
 tract2000 <- readOGR(dsn = "./input/sea_tract_2000/sea_tract_2000.shp",
                      layer = "sea_tract_2000",
                      GDAL1_integer64_policy = TRUE,
+                     verbose = F,
                      stringsAsFactors = F) 
 tract2000 <- spTransform(tract2000, WGS.latlon)
 proj4string(scatter_cl) <- WGS.latlon

### 2018-06-03 dashboard/index.rmd
progress on site with multi-page

--- 2018-06-03 index.rmd
@@
 ---
-title: "smooth-sea"
+title: ""
 output: 
@@
        - {title: "hesscl.com", href: "http://hesscl.com", align: right}
+       - {title: "Project Home", icon: "fa-home", href: "http://hesscl.com/smooth-sea/index.html", align: left} 
     orientation: columns
     theme: bootstrap
-    logo: logo.svg
+    logo: "../logo.svg"
     source_code: "https://github.com/hesscl/smooth-sea"
@@
 #source data_load()
-source("helpers/load_data.R")
+source("../helpers/load_data.R")
 
 #load ACS data extract (codebook in folder)
-census <- read_csv(file = "./input/acsExtract.csv")
-censusRent <- read_csv(file = "./input/acsRent.csv")
+census <- read_csv(file = "../input/acsExtract.csv")
+censusRent <- read_csv(file = "../input/acsRent.csv")
 census <- inner_join(census, censusRent)
@@
 #read in tract shapefiles for seattle
-sea_shp <- readOGR(dsn = "./input/sea_tract_2010/sea_tract_2010.shp",
+sea_shp <- readOGR(dsn = "../input/sea_tract_2010/sea_tract_2010.shp",
                   layer = "sea_tract_2010",
@@
 #create neighbor file that INLA takes for hyperparameter args
-nb2INLA("./output/seatract.graph", sea_adj)
+nb2INLA("../output/seatract.graph", sea_adj)
 sea_df$idqtr <- factor(sea_df$listingQtr, ordered = T)
@@
     scale.model = T,
-    graph = "./output/seatract.graph") +
+    graph = "../output/seatract.graph") +
   f(idqtr, model = "ar1") +
@@
 
-Home {data-icon="fa-home"}
+Introduction
 =====================================
@@
 
-This website provides interactive results for the forecasting models explored in the paper _Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_.
+This website provides interactive results for the forecasting models explored in the paper _Estimating Neighborhood Rents using Scraped Data_.
 

### 2018-06-03 index.rmd
progress on site with multi-page

@@
 ---
-title: "smooth-sea"
+title: "Estimating Neighborhood Rents with Scraped Data"
+author: "Chris Hess"
+date: "June 2, 2018"
 output: 
-  flexdashboard::flex_dashboard:
-    navbar:
-       - {title: "hesscl.com", href: "http://hesscl.com", align: right} 
-    orientation: columns
-    theme: bootstrap
-    logo: logo.svg
-    source_code: "https://github.com/hesscl/smooth-sea"
+  html_document
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+<br>
+<br>
 
-#packages
-library(tidyverse) #need dev version of ggplot
-library(forcats)
-library(haven)
-library(lubridate)
-library(haven)
-library(zoo)
-library(ggthemes)
-library(RColorBrewer)
-library(sp)
-library(spdplyr)
-library(geosphere)
-library(rgdal)
-library(rgeos)
-library(sqldf)
-library(spdep)
-library(viridis)
-library(latex2exp)
-library(plotly)
-library(DT)
-library(crosstalk)
+## Project Contents
 
-#source data_load()
-source("helpers/load_data.R")
+[Working paper on agreement between Seattle neighborhood rent indicators](./manuscript/index.html)
 
-#load ACS data extract (codebook in folder)
-census <- read_csv(file = "./input/acsExtract.csv")
-censusRent <- read_csv(file = "./input/acsRent.csv")
-census <- inner_join(census, censusRent)
+[Forecasting based on models used in working paper](./dashboard/index.html)
 
-#read in tract shapefiles for seattle
-sea_shp <- readOGR(dsn = "./input/sea_tract_2010/sea_tract_2010.shp",
-                  layer = "sea_tract_2010",
-                  GDAL1_integer64_policy = TRUE,
-                  stringsAsFactors = F)
-
-#use sourced function to load listing summaries
-tract <- load_data()
-
-#### A. construct panel (some missingness) ------------------------------------
-
-#ensure order of census df matches shapefile
-census <- census[match(sea_shp@data$GISJOIN, census$GISJOIN),]
-
-#clean a few fields we might use
-census <- census %>%
-  mutate(nHU = AF7PE001,
-         idtract = row_number(),
-         pOwnoccHU = AF7PE002/AF7PE001,
-         medHUVal = AF9LE001)
-
-#store NULL object to bind iterated dfs into
-panel <- NULL
-
-#for each unique quarter in the CL data
-for(i in unique(as.character(tract$listingQtr))){
-  period <- tract %>% #start pipe with tract, end with period
-    mutate(listingQtr = factor(listingQtr, ordered = T)) %>% #make factor
-    filter(listingQtr == i) %>% #filter to qtr i
-    right_join(census) %>% #right join to census, allows missingness
-    ungroup() %>% #remove grouping by tract
-    mutate(listingQtr = i, #listingQtr == current iteration
-           Qtr = gsub(pattern = "^\\d{4} ", replacement = "", x = listingQtr)) #subset str to Q1,2,3,4
-  panel <- bind_rows(panel, period) #append to panel object
-}
-
-#establish object for modeling based off of adjusting `panel`
-sea_df <- panel %>% 
-  mutate(actualRent = med1B, #first preserve the observed / observed-to-date
-         med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B)) %>% #set to NA for INLA to forecast current qtr
-  mutate(p1B = n1B/nListings) %>% #compute share of 1B listings
-  arrange(listingQtr, GISJOIN) #arange qtr, tract
-
-#### B. INLA Models of CL Rent ------------------------------------------------
-
-library(INLA)
-
-#create adjacency matrix from shapefile
-sea_adj <- poly2nb(sea_shp)
-
-#create neighbor file that INLA takes for hyperparameter args
-nb2INLA("./output/seatract.graph", sea_adj)
-sea_df$idqtr <- factor(sea_df$listingQtr, ordered = T)
-sea_df$idqtr1 <- sea_df$idqtr
-sea_df$idtractqtr <- paste(sea_df$idtract, sea_df$Qtr)
-sea_df$idtract1 <- sea_df$idtract
-sea_df$Qtr1 <- sea_df$Qtr
-
-#fixed effect for qtr only
-form.int <- log(med1B) ~ 1 + Qtr
-
-m.int <- inla(form.int, 
-             family = "normal", 
-             data = sea_df,
-             control.predictor = list(compute = TRUE),
-             control.compute = list(dic = TRUE, waic = TRUE))
-
-summary(m.int)
-sea_df$int_Med <- m.int$summary.linear.predictor[, "0.5quant"]
-sea_df$int_SD <- m.int$summary.linear.predictor[, "sd"]
-sea_df$int_postWidth <- m.int$summary.linear.predictor[, "0.975quant"] - m.int$summary.linear.predictor[,"0.025quant"]
-
-#fixed qtr + iid tract random effect
-form.ns <- log(med1B) ~ 1 + Qtr +
-   f(idtract, model = "iid")
-
-m.ns <- inla(form.ns, 
-              family = "normal", 
-              data = sea_df,
-              control.predictor = list(compute = TRUE),
-              control.compute = list(dic = TRUE, waic = TRUE))
-
-summary(m.ns)
-sea_df$ns_Med <- m.ns$summary.linear.predictor[, "0.5quant"]
-sea_df$ns_SD <- m.ns$summary.linear.predictor[, "sd"]
-sea_df$ns_postWidth <- m.ns$summary.linear.predictor[, "0.975quant"] - m.ns$summary.linear.predictor[,"0.025quant"]
-sea_df$ns_Eff <- m.ns$summary.random$idtract[sea_df$idtract, "0.5quant"]
-
-#IID tract random effect + AR(1) process
-form.nsar1 <- log(med1B) ~ 1 +
-  f(idtract, model = "iid") +
-  f(idqtr, model = "ar1") +
-  f(idqtr1, model = "iid")
-
-m.nsar1 <- inla(form.nsar1, 
-               family = "normal", 
-               data = sea_df,
-               control.predictor = list(compute = TRUE),
-               control.compute = list(dic = TRUE, waic = TRUE))
-
-summary(m.nsar1) #NB: does not improve fit to use random effect time specification, using fixed effects
-sea_df$nsar1_Med <- m.nsar1$summary.linear.predictor[, "0.5quant"]
-sea_df$nsar1_SD <- m.nsar1$summary.linear.predictor[, "sd"]
-sea_df$nsar1_postWidth <- m.nsar1$summary.linear.predictor[, "0.975quant"] - m.nsar1$summary.linear.predictor[,"0.025quant"]
-sea_df$nsar1_Eff <- m.nsar1$summary.random$idtract[sea_df$idtract, "0.5quant"]
-
-#spatial random effect model (BYM 1991) with AR(1)
-form.bym <- log(med1B) ~ 1 +
-  f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
-    scale.model = T,
-    graph = "./output/seatract.graph") +
-  f(idqtr, model = "ar1") +
-  f(idqtr1, model = "iid")
-
-m.bym <- inla(form.bym, 
-             family = "normal", 
-             data = sea_df,
-             control.predictor = list(compute = TRUE),
-             control.compute = list(dic = TRUE, waic = TRUE))
-
-summary(m.bym)
-sea_df$bym_Med <- m.bym$summary.linear.predictor[, "0.5quant"]
-sea_df$bym_SD <- m.bym$summary.linear.predictor[, "sd"]
-sea_df$bym_postWidth <- m.bym$summary.linear.predictor[, "0.975quant"] - m.bym$summary.linear.predictor[,"0.025quant"]
-sea_df$bym_Eff <- m.bym$summary.random$idtract[sea_df$idtract, "0.5quant"]
-
-#spatio-temporal smoothing model, with AR(1) and linear space-time interaction (Bernadelli 1995)
-form.spt <- log(med1B) ~ 1 + 
-  f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
-    scale.model = T,
-    graph = "./output/seatract.graph") +
-  f(idqtr, model = "ar1") +
-  f(idqtr1, model = "iid") +
-  f(idtractqtr, model = "iid")
-
-m.spt <- inla(form.spt, 
-              family = "normal", 
-              data = sea_df,
-              control.predictor = list(compute = TRUE),
-              control.compute = list(dic = TRUE, waic = TRUE),
-              control.inla = list(correct = TRUE, correct.factor = 10))
-
-summary(m.spt)
-sea_df$spt_Med <- m.spt$summary.linear.predictor[, "0.5quant"]
-sea_df$spt_SD <- m.spt$summary.linear.predictor[, "sd"]
-sea_df$spt_postWidth <- m.spt$summary.linear.predictor[, "0.975quant"] - m.spt$summary.linear.predictor[,"0.025quant"]
-sea_df$spt_Eff <- m.spt$summary.random$idtract[sea_df$idtract, "0.5quant"]
-
-#overfitting? too much bumpiness qtr-to-qtr
-
-#### Maps of Model Output ------------------------------------------------------------------
-
-sea_df <- sea_df %>%
-  mutate_at(.vars = vars(matches("_Med")),
-            .funs = exp) %>%
-  mutate_at(.vars = vars(matches("_SD")),
-            .funs = exp) %>%
-  group_by(GISJOIN) %>% 
-  arrange(GISJOIN, listingQtr) %>%
-  mutate(spt_traj = spt_Med - lag(spt_Med, 4)) %>%
-  ungroup %>%
-  group_by(GISJOIN) %>%
-  mutate(spt_traj = sum(spt_traj, na.rm=T)) %>%
-  ungroup() %>%
-  mutate(bym_cuts = cut(bym_Med, 6),
-         spt_cuts = cut(spt_Med, 6))
-
-sea_shp@data$id <- rownames(sea_shp@data)
-sea_shp@data <- left_join(sea_shp@data, sea_df)
-sea_f <- fortify(sea_shp)
-sea_f <- inner_join(sea_f, sea_shp@data,"id")
-
-rmse <- function(error){sqrt(mean(error^2))}
-mae <- function(error){mean(abs(error))}
-
-#table of fit statistics
-fit_stats <- sea_df %>%
-  mutate(train_test = ifelse(listingQtr == as.yearqtr(Sys.Date()), "Test", "Training")) %>%
-  mutate(int_err = actualRent - int_Med,
-         ns_err = actualRent - ns_Med,
-         nsar1_err = actualRent - nsar1_Med,
-         bym_err = actualRent - bym_Med,
-         spt_err = actualRent - spt_Med) %>% 
-  group_by(train_test) %>%
-  summarize(int_rmse = rmse(int_err[!is.na(int_err)]),
-            int_mae = mae(int_err[!is.na(int_err)]),
-            ns_rmse = rmse(ns_err[!is.na(ns_err)]),
-            ns_mae = mae(ns_err[!is.na(ns_err)]),
-            nsar1_rmse = rmse(nsar1_err[!is.na(nsar1_err)]),
-            nsar1_mae = mae(nsar1_err[!is.na(nsar1_err)]),
-            bym_rmse = rmse(bym_err[!is.na(bym_err)]),
-            bym_mae = mae(bym_err[!is.na(bym_err)]),
-            spt_rmse = rmse(spt_err[!is.na(spt_err)]),
-            spt_mae = mae(spt_err[!is.na(spt_err)])) %>%
-  dplyr::select(train_test, ends_with("_rmse"), ends_with("_mae"))
-```
-
-Home {data-icon="fa-home"}
-=====================================
-
-Column
--------------------------------------
-
-### Welcome!
-
-This website provides interactive results for the forecasting models explored in the paper _Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_.
-
-The goal of this research is to a.) understand the temporal dynamics of rent estimates in Seattle and b.) forecast the current quarter's rent levels based off of the prior periods.
-The focal series of models regress median one-bedroom rent asked values on different specifications of the panel's correlation structure (i.e.
-temporal and spatial).
-All of the candidate models' posterior distributions are estimated with [integrated nested Laplace approximations](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) (INLA) using the default, weakly-informative priors for all model hyperparameters.
-Throughout the following analyses, the training data are `r min(sea_df$listingQtr)` up to the prior quarter (i.e.
-`r as.yearqtr(Sys.Date())-.25`).
-The test period is a forecast for the current period and includes comparison to the appropriate median rent estimates for data observed so far in this period.
-
-Most graphics include some level of interactivity, usually either hover-over tooltip information or a slider to control various views of the graphic.
-Clicking on cases will highlight data elements in most graphics, and double-clicking will reset the graphic.
-You can download the source code and data for this project from Github [here](https://github.com/hesscl/smooth-sea).
-
-Contact Chris Hess at [hesscl@uw.edu](mailto://hesscl@uw.edu) for more information about this research.
-
-_This page was last updated: `r Sys.Date()`_
+[GitHub Repository](https://github.com/hesscl/smooth-sea/)
 
 <br>
-<hr>
+
+## See also
+
+[Helena](http://helena-lang.org) - Web automation for end users
+
 <br>
 
-**Table of Contents**
-
-Page                |  Description
-------------------- | ---------------------------------------------------------
-Distribution        | density graphic to investigate the distribution of rents among Seattle neighborhoods for each quarter
-Panel Time-Series   | line graphic to show the observed or modeled temporal structure
-Spatial Time-Series | series of maps to show observed change across time
-Model Fit           | tables of model root mean square error (RMSE), mean absolute error (MAE), and deviance information criterion (DIC) across training and test data
-
-Column
--------------------------------------
-
-### Observed vs. Smoothed Rent Estimates
-
-```{r index.maps, echo=F}
-obs <- sea_f %>%
-  mutate(rent = actualRent) %>%
-  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
-  mutate(var = paste0(as.yearqtr(Sys.Date()), "-to-Date")) %>%
-  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, spt_Med, NAMELSAD10, n1B)
-
-pred <- sea_f %>%
-  mutate(rent = bym_Med) %>%
-  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
-  mutate(var = paste0(as.yearqtr(Sys.Date()), " Forecast")) %>%
-  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, bym_Med, NAMELSAD10, n1B)
-
-comb <- bind_rows(obs, pred) %>%
-  mutate(tooltip = paste0(NAMELSAD10, "<br>", "Forecast: $", round(bym_Med), "<br>", "Qtr-to-Date: $", round(actualRent), "<br>", "N Qtr-to-Date: ", n1B))
-comb <- SharedData$new(comb, ~GISJOIN)
-
-avp_map <- ggplot(comb, aes(x = long, y = lat, group = group, fill = rent, label = tooltip)) +
-  facet_wrap(~ var) +
-  geom_polygon(color = "grey85", lwd = .05) +
-  scale_fill_viridis_c(limits = c(750, 2500), labels = scales::dollar) +
-  coord_quickmap() +
-  theme(axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank(),
-        panel.background = element_blank()) +
-  labs(fill = "Median<br>1B Rent")
-
-ggplotly(avp_map, tooltip = "label") %>%
-  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
-  highlight(on = "plotly_click")
-```
-
-Distribution
-=====================================
-
-```{r, echo=F}
-gg_hist <- ggplot(sea_df, aes(x = med1B, group = listingQtr, color = listingQtr)) +
-  geom_density() +
-  theme_minimal() +
-  scale_x_continuous(labels = scales::dollar) +
-  scale_color_viridis_d(end = .8) +
-  xlab("\nMedian 1B Rent") +
-  labs(color = "Quarter")
-
-ggplotly(gg_hist, tooltip = c("color", "x")) %>%
-  layout(autosize = T)
-```
-
-Panel Time-Series
-=====================================
-
-```{r temp.lineplot, echo=F}
-obs_df <- sea_df %>%
-  group_by(listingQtr) %>%
-  mutate(value = actualRent,
-         qtrDev = value-mean(value, na.rm=T),
-         model = "A. Observed")
-
-int_df <- sea_df %>%
-  group_by(listingQtr) %>%
-  mutate(value = int_Med,
-         qtrDev = value-mean(value, na.rm=T),
-         model = "B. Quarter FE") %>%
-  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
-
-ns_df <- sea_df %>%
-  group_by(listingQtr) %>%
-  mutate(value = ns_Med,
-         qtrDev = value-mean(value, na.rm=T),
-         model = "C. Non-Spatial") %>%
-  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
-
-nsar1_df <- sea_df %>%
-  group_by(listingQtr) %>%
-  mutate(value = nsar1_Med,
-         qtrDev = value-mean(value, na.rm=T),
-         model = "D. Non-Spatial AR(1)") %>%
-  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
-
-bym_df <- sea_df %>%
-  group_by(listingQtr) %>%
-  mutate(value = bym_Med,
-         qtrDev = value-mean(value, na.rm=T),
-         model = "E. Spatial AR(1)") %>%
-  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
-
-spt_df <- sea_df %>%
-  group_by(listingQtr) %>%
-  mutate(value =  spt_Med,
-         qtrDev = value-mean(value, na.rm=T),
-         model = "F. Spatiotemporal AR(1)") %>%
-  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
-
-panel_df <- bind_rows(obs_df, int_df, ns_df, nsar1_df, bym_df, spt_df)
-
-panel_df <- SharedData$new(panel_df, ~GISJOIN)
-
-obs_lineplot <- ggplot(panel_df, aes(x = listingQtr, y = value, group = GISJOIN, label = GISJOIN, color = qtrDev, frame = model)) +
-  geom_line(lwd = .25, color = "black", alpha = .5) +
-  geom_point() +
-  scale_color_gradient2_tableau() +
-  scale_y_continuous(labels = scales::dollar) +
-  xlab("") +
-  ylab("Median 1B Rent\n") +
-  theme_minimal() +
-  labs(color = "Quarterly\ndeviation")
-
-ggplotly(obs_lineplot, tooltip = c("label", "y")) %>%
-  layout(autosize = T) %>%
-  animation_opts(frame = 1000, easing = "linear", redraw = FALSE) %>%
-  animation_button(hide = T) %>%
-  highlight()
-```
-
-Spatial Time-Series
-=====================================
-
-Column {.tabset}
--------------------------------------
-
-### Observed
-
-```{r spat.series.actual, echo=F}
-map_panel_obs <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = actualRent)) +
-  facet_grid(~ listingQtr) +
-  geom_polygon(color = "grey85", lwd = .05) +
-  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
-  coord_quickmap() +
-  theme(axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank(),
-        panel.background = element_blank()) +
-  labs(fill = "Median<br>1B Rent")
-
-ggplotly(map_panel_obs) %>%
-  layout(autosize = T)
-```
-
-### Non-Spatial AR(1)
-
-```{r spat.series.nsar1, echo=F}
-map_panel_nsar1 <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = nsar1_Med, frame = listingQtr)) +
-  geom_polygon(color = "grey85", lwd = .05) +
-  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
-  coord_quickmap() +
-  theme(axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank(),
-        panel.background = element_blank()) +
-  labs(fill = "Median<br>1B Rent")
-
-ggplotly(map_panel_nsar1) %>%
-  layout(autosize = T) %>%
-  animation_opts(1000, transition = 0, redraw = FALSE)
-```
-
-### Spatial AR(1)
-
-```{r spat.series.bym, echo=F}
-map_panel_bym <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = bym_Med, frame = listingQtr)) +
-  geom_polygon(color = "grey85", lwd = .05) +
-  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
-  coord_quickmap() +
-  theme(axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank(),
-        panel.background = element_blank()) +
-  labs(fill = "Median<br>1B Rent")
-
-ggplotly(map_panel_bym) %>%
-  layout(autosize = T) %>%
-  animation_opts(1000, transition = 0, redraw = FALSE)
-```
-
-### Spatiotemporal AR(1)
-
-```{r spat.series.spt, echo=F}
-map_panel_spt <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = spt_Med, frame = listingQtr)) +
-  geom_polygon(color = "grey85", lwd = .05) +
-  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
-  coord_quickmap() +
-  theme(axis.title = element_blank(),
-        axis.text = element_blank(),
-        axis.ticks = element_blank(),
-        panel.background = element_blank()) +
-  labs(fill = "Median<br>1B Rent")
-
-ggplotly(map_panel_spt) %>%
-  layout(autosize = T) %>%
-  animation_opts(1000, transition = 0, redraw = FALSE)
-```
-
-Model Fit
-=====================================
-
-Column {.tabset}
--------------------------------------
-
-### Model Legend
-
-Model abbr.                   | Description
------------------------------ | -----------------------------------------------
-**int**    | Quarter fixed intercept
-           | `r format(form.int)`
-**ns**     | Non-spatial tract random effect for each tract, quarter fixed intercept
-           | `r format(form.ns)`
-**nsar1**  | Non-spatial tract random effect, AR(1) random effect for prior quarter, i.i.d random effect for current quarter
-           | `r format(form.nsar1)`
-**bym**    | Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, non-spatial i.i.d tract random effect, AR(1) random effect for prior quarter, i.i.d random effect for current quarter
-           |  `r format(form.bym)`
-**spt**    | Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, non-spatial i.i.d tract random effect, AR(1) random effect for prior quarter, i.i.d random effect for current quarter, i.i.d.
-random effect for tract-quarter (space-time interaction)
-           | `r format(form.spt)`
-
-### Accuracy and Information Criteria
-
-```{r, echo=F}
-knitr::kable(fit_stats %>% dplyr::select(train_test, ends_with("rmse")))
-```
-
-<br>
-<hr>
-<br>
-
-```{r, echo=F}
-knitr::kable(fit_stats %>% dplyr::select(train_test, ends_with("mae")))
-```
-
-<br>
-<hr>
-<br>
-
-```{r, echo=F}
-knitr::kable(data.frame("train_test" = "Training",
-                        "int_DIC" = m.int$dic$dic,
-                        "ns_DIC" = m.ns$dic$dic,
-                        "nsar1_DIC" = m.nsar1$dic$dic,
-                        "bym_DIC" = m.bym$dic$dic,
-                        "spt_DIC" = m.spt$dic$dic))
-```
-
-<br>
-<hr>
-<br>
-
-```{r, echo=F}
-knitr::kable(data.frame("train_test" = "Training",
-                        "int_WAIC" = m.int$waic$waic,
-                        "ns_WAIC" = m.ns$waic$waic,
-                        "nsar1_WAIC" = m.nsar1$waic$waic,
-                        "bym_WAIC" = m.bym$waic$waic,
-                        "spt_WAIC" = m.spt$waic$waic))
-```
-
-### Hyperparameters
-
-```{r, echo=F}
-knitr::kable(summary(m.nsar1)$hyperpar, caption = "Non-Spatial AR(1)")
-```
-
-<br>
-<hr>
-<br>
-
-```{r, echo=F}
-knitr::kable(summary(m.bym)$hyperpar, caption = "Spatial AR(1)")
-```
-
-<br>
-<hr>
-<br>
-
-```{r, echo=F}
-knitr::kable(summary(m.spt)$hyperpar, caption = "Spatiotemporal AR(1)")
-```
-
-Column {.tabset}
--------------------------------------
-
-### Non-Spatial AR(1)
-
-```{r, echo=F}
-avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = nsar1_Med, size = n1B, frame = listingQtr)) +
-  geom_point() +
-  geom_abline(intercept = 0, slope = 1) +
-  scale_x_continuous(limits = c(800, 2600), labels = scales::dollar) +
-  scale_y_continuous(limits = c(800, 2600), labels = scales::dollar) +
-  xlab("Actual") +
-  ylab("Predicted") +
-  scale_size_area() +
-  coord_equal() +
-  theme_minimal()
-
-ggplotly(avp_scatter) %>%
-  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
-  animation_opts(1000, transition = 0, redraw = T)
-```
-
-### Spatial AR(1)
-
-```{r, echo=F}
-avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = bym_Med, size = n1B, frame = listingQtr)) +
-  geom_point() +
-  geom_abline(intercept = 0, slope = 1) +
-  scale_x_continuous(limits = c(800, 2600), labels = scales::dollar) +
-  scale_y_continuous(limits = c(800, 2600), labels = scales::dollar) +
-  xlab("Actual") +
-  ylab("Predicted") +
-  scale_size_area() +
-  coord_equal() +
-  theme_minimal()
-
-ggplotly(avp_scatter) %>%
-  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
-  animation_opts(1000, transition = 0, redraw = FALSE)
-```
-
-### Spatiotemporal AR(1)
-
-```{r, echo=F}
-avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = spt_Med, size = n1B, frame = listingQtr)) +
-  geom_point() +
-  geom_abline(intercept = 0, slope = 1) +
-  scale_x_continuous(limits = c(800, 2600), labels = scales::dollar) +
-  scale_y_continuous(limits = c(800, 2600), labels = scales::dollar) +
-  xlab("Actual") +
-  ylab("Predicted") +
-  scale_size_area() +
-  coord_equal() +
-  theme_minimal()
-
-ggplotly(avp_scatter) %>%
-  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
-  animation_opts(1000, transition = 0, redraw = FALSE)
-```
+## Contact Informatiion

### 2018-06-03 manuscript/index.Rmd
progress on site with multi-page

--- 2018-06-02 smooth-sea.Rmd
@@
 ---
-title: "Estimating Neighborhood Rents with Scraped Web Data"
+title: "Estimating Neighborhood Rents with Scraped Data"
 author: "Chris Hess"
@@
 output: html_document
+navbar:
+  - {title: "hesscl.com", href: "http://hesscl.com", align: right}
+  - {title: "Project Home", icon: "fa-home", href: "http://hesscl.com/smooth-sea/index.html", align: left}
 abstract: "Online classified ads are a key source of information about local rental housing markets.
@@
 This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $90 for the training data and $145 for the test data."
+
 ---
@@
 #load ACS data extract (codebook in folder)
-census <- read_csv(file = "./input/acsExtract.csv")
-censusRent <- read_csv(file = "./input/acsRent.csv")
+census <- read_csv(file = "../input/acsExtract.csv")
+censusRent <- read_csv(file = "../input/acsRent.csv")
 census <- inner_join(census, censusRent)
@@
 #go ahead and save separate versions for graphics later
-tract2000 <- readOGR(dsn = "./input/sea_tract_2000/sea_tract_2000.shp",
+tract2000 <- readOGR(dsn = "../input/sea_tract_2000/sea_tract_2000.shp",
                   layer = "sea_tract_2000",
@@
 #source data_load()
-source("./helpers/load_data.R")
+source("../helpers/load_data.R")
 
@@
 #determine config (dev or github)
-if(file.exists("../data/cl/craigslistDB.sqlite")){
-  source("./helpers/map_point_density.R") #if config, refresh plot pdf
+if(file.exists("../../data/cl/craigslistDB.sqlite")){
+  source("../helpers/map_point_density.R") #if config, refresh plot pdf
 }
@@
 
-![Figure 1. Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](./output/map_point_density.png)
+![Figure 1. Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](../output/map_point_density.png)
 <br>
@@
 #determine config (dev or github)
-if(file.exists("../data/cl/craigslistDB.sqlite")){
-  source("./helpers/rent_lineplot.R") #if config, refresh plot pdf
+if(file.exists("../../data/cl/craigslistDB.sqlite")){
+  source("../helpers/rent_lineplot.R") #if config, refresh plot pdf
 }
@@
 Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual and ACS are annual (i.e.
-respective 1-Year Estimates for each year period).](./output/rent_lineplot.png)
+respective 1-Year Estimates for each year period).](../output/rent_lineplot.png)
 
@@
 #determine config (dev or github)
-if(file.exists("../data/cl/craigslistDB.sqlite")){
-  source("./helpers/rent_map_across.R") #if config, refresh plot pdf
+if(file.exists("../../data/cl/craigslistDB.sqlite")){
+  source("../helpers/rent_map_across.R") #if config, refresh plot pdf
 }
@@
 Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates.
-Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled.](./output/rent_map_across_1B.png)
+Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled.](../output/rent_map_across_1B.png)
 
@@
 #determine config (dev or github)
-if(file.exists("../data/cl/craigslistDB.sqlite")){
-  source("./helpers/rent_scatter_across.R") #if config, refresh plot pdf
+if(file.exists("../../data/cl/craigslistDB.sqlite")){
+  source("../helpers/rent_scatter_across.R") #if config, refresh plot pdf
 }
@@
 Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott.
-Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.](./output/rent_scatter_across_1B.png)
+Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.](../output/rent_scatter_across_1B.png)
 <br>

### 2018-06-03 helpers/load_data.R

@@ -1,8 +1,8 @@
 load_data <- function(){
   #determine config (dev or github)
-  if(file.exists("../data/cl/craigslistDB.sqlite")){
+  if(file.exists("../../data/cl/craigslistDB.sqlite")){
     #connect to database if file.exists
-    DB <- dbConnect(SQLite(), dbname="../data/cl/craigslistDB.sqlite")
+    DB <- dbConnect(SQLite(), dbname="../../data/cl/craigslistDB.sqlite")
     cl <- tbl(DB, "clean") #create dbi for clean listing table
     
     #compute tract aggregates for CL listing count
@@ -29,11 +29,11 @@ load_data <- function(){
       ungroup %>%
       arrange(GISJOIN, listingQtr)
     dbDisconnect(DB)
-    write_csv(tract, "./input/tractCl.csv")
+    write_csv(tract, "../input/tractCl.csv")
     return(tract)
   } else{
     #if github config, read in extract from cl db
-    tract <- read_csv("./input/tractCl.csv")
+    tract <- read_csv("../input/tractCl.csv")
     return(tract)
   }
 }

### 2018-06-03 helpers/map_point_density.R

@@ -39,7 +39,7 @@ with_points <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", darken = 
         strip.background = element_blank())
 
 blank + with_points +
-  ggsave(filename = "./output/map_point_density.png",
+  ggsave(filename = "../output/map_point_density.png",
          width = 6, height = 6, units = "in", dpi = 300)
 
 

### 2018-06-03 helpers/rent_lineplot.R

@@ -111,5 +111,5 @@ ggplot(sumRent, aes(x = moYr, y = Rent, group = dataSrc, color = dataSrc, shape 
   ylab("Median Rent\n") +
   labs(color = "Data Source",
        shape = "Data Source") +
-  ggsave(file = "./output/rent_lineplot.png",
+  ggsave(file = "../output/rent_lineplot.png",
          device = "png", width = 6, height = 4)

### 2018-06-03 helpers/rent_map_across.R

@@ -148,5 +148,5 @@ ggplot(tract_f %>% filter(catBeds == "1 Bedroom"), aes(x = long, y = lat, group 
         legend.title = element_text(size = 14),
         legend.text = element_text(size = 12)) +
   labs(fill = "Asking Rent") +
-  ggsave(filename = "./output/rent_map_across_1B.png",
+  ggsave(filename = "../output/rent_map_across_1B.png",
          width = 8, height = 4, dpi = 300)

### 2018-06-03 helpers/rent_scatter_across.R

@@ -27,10 +27,10 @@ coordinates(scatter_cl) <- cbind(scatter_cl$lng, scatter_cl$lat)
 
 #make sure shapefile has correct projection
 WGS.latlon <- CRS("+proj=longlat +datum=WGS84")
-tract2000 <- readOGR(dsn = "./input/sea_tract_2000/sea_tract_2000.shp",
+tract2000 <- readOGR(dsn = "../input/sea_tract_2000/sea_tract_2000.shp",
                      layer = "sea_tract_2000",
                      GDAL1_integer64_policy = TRUE,
-                     verbose = F,
+                     verbose = FALSE,
                      stringsAsFactors = F) 
 tract2000 <- spTransform(tract2000, WGS.latlon)
 proj4string(scatter_cl) <- WGS.latlon
@@ -75,5 +75,5 @@ ggplot(across2017, aes(x = dsRent, y = clRent, size = nHU)) +
   xlab("\nDupre+Scott September 2017 Median 1B Rent") +
   ylab("Craigslist 2017-2018 Median 1B Rent\n") +
   labs(size = "N Craigslist\nlistings") +
-  ggsave(file = "./output/rent_scatter_across_1B.png",
+  ggsave(file = "../output/rent_scatter_across_1B.png",
          dpi = 300, width = 6, height = 4)

### 2018-06-03 dashboard/index.rmd
a little more progress on the project site

@@
 ---
-title: ""
 output: 
@@
 Clicking on cases will highlight data elements in most graphics, and double-clicking will reset the graphic.
-You can download the source code and data for this project from Github [here](https://github.com/hesscl/smooth-sea).
-
-Contact Chris Hess at [hesscl@uw.edu](mailto://hesscl@uw.edu) for more information about this research.
 

### 2018-06-03 index.rmd
fixing landing page + adding under construction notice
a little more progress on the project site

@@
 author: "Chris Hess"
-date: "June 2, 2018"
 output: 
-  html_document
+  html_document:
+      include:
+        after_body: footer.html
 ---
@@
 <br>
+
+This project is under construction and subject to change at any moment!
+
 <br>
 
-## Project Contents
+### Contents
 
@@
 
-## See also
+### See also
 
@@
 
-## Contact Informatiion
+### Contact
+
+Email Chris Hess at [hesscl@uw.edu](mailto://hesscl@uw.edu) for more information about this research.

### 2018-06-03 manuscript/index.Rmd
a little more progress on the project site

@@
 The rising cost of housing therefore has a direct role in neighborhood processes of "demographic inversion" like urban gentrification and suburbanizing poverty.
-Census data (i.e.
-decennial and American Community Survey) have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more.
-This publication schedules prohibit analyzing any market dynamics beyond long-term trends, particularly among smaller areas of aggregation like neighborhoods.
+American Community Survey data have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more.
+This publication lag for ACS data prohibits analyzing any market dynamics beyond long-term trends, particularly for estimates of smaller areas of aggregation like neighborhoods where ACS uses 5-year rolling-average methods as well.
 While more up-to-date options for neighborhood rent data exist through privately-held companies, such data can be somewhat of a black box in their creation and may be biased towards the costs of corporate-held properties.
@@
 ![Figure 1. Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](../output/map_point_density.png)
+
 <br>
@@
 Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.](../output/rent_scatter_across_1B.png)
+
 <br>

### 2018-06-03 index.rmd
adding navbar

@@
   html_document:
-      include:
+      includes:
+        in_header: header.html
         after_body: footer.html
@@
 
-[Working paper on agreement between Seattle neighborhood rent indicators](./manuscript/index.html)
+[Working paper on estimating Seattle neighborhood rent indicators](./manuscript/index.html)
 
-[Forecasting based on models used in working paper](./dashboard/index.html)
+[`flexdashboard` for working paper's model diagnostics and forecasts](./dashboard/index.html)
 

### 2018-06-03 manuscript/index.Rmd
adding navbar

@@
 date: "June 2, 2018"
-output: html_document
-navbar:
-  - {title: "hesscl.com", href: "http://hesscl.com", align: right}
-  - {title: "Project Home", icon: "fa-home", href: "http://hesscl.com/smooth-sea/index.html", align: left}
+output: 
+  html_document:
+    toc: true
+    toc_float: true
+    include:
+      in_header: header.html
 abstract: "Online classified ads are a key source of information about local rental housing markets.

### 2018-06-03 helpers/map_point_density.R

@@ -40,6 +40,6 @@ with_points <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", darken = 
 
 blank + with_points +
   ggsave(filename = "../output/map_point_density.png",
-         width = 6, height = 6, units = "in", dpi = 300)
+         width = 5, height = 5, units = "in", dpi = 300)
 
 

### 2018-06-03 manuscript/index.Rmd
a few more formatting improvements

@@
     toc_float: true
-    include:
-      in_header: header.html
+    includes:
+      in_header: "../header.html"
+      after_body: "../footer.html"
 abstract: "Online classified ads are a key source of information about local rental housing markets.
@@
 This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $90 for the training data and $145 for the test data."
+---
 
----
+<style>
+.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
+    background-color: #4b2e83;
+}
+</style>
 

### 2018-06-03 dashboard/index.rmd
more tweaks to layout, relatively finalized

@@
        - {title: "hesscl.com", href: "http://hesscl.com", align: right}
-       - {title: "Project Home", icon: "fa-home", href: "http://hesscl.com/smooth-sea/index.html", align: left} 
+       - {title: "Project Home", icon: "fa-home", href: "http://hesscl.com/smooth-sea/index.html", align: right} 
     orientation: columns

### 2018-06-03 manuscript/index.Rmd
more tweaks to layout, relatively finalized

@@
     includes:
-      in_header: "../header.html"
-      after_body: "../footer.html"
+      in_header: header.html
+      after_body: footer.html
 abstract: "Online classified ads are a key source of information about local rental housing markets.

### 2018-06-03 dashboard/index.rmd
adding hesscl.com link to navbar of two pages

--- 2018-06-03 dashboard/index.rmd
@@
 ---
+title: "smooth-sea"
 output: 

### 2018-06-03 index.rmd
adding hesscl.com link to navbar of two pages

@@
 title: "Estimating Neighborhood Rents with Scraped Data"
-author: "Chris Hess"
 output: 

### 2018-06-03 dashboard/index.rmd
forgot the personal page link on the manuscript md

--- 2018-06-03 dashboard/index.rmd
@@
     logo: "../logo.svg"
-    source_code: "https://github.com/hesscl/smooth-sea"
+    source_code: "https://github.com/hesscl/smooth-sea/"
 ---

### 2018-06-03 dashboard/index.rmd
update to main index, also settled navbar for now

--- 2018-06-03 dashboard/index.rmd
@@
        - {title: "hesscl.com", href: "http://hesscl.com", align: right}
-       - {title: "Project Home", icon: "fa-home", href: "http://hesscl.com/smooth-sea/index.html", align: right} 
+       - {title: "Project Home", href: "http://hesscl.com/smooth-sea/index.html", align: left} 
     orientation: columns
@@
     logo: "../logo.svg"
-    source_code: "https://github.com/hesscl/smooth-sea"
 ---

### 2018-06-03 index.rmd
update to main index, also settled navbar for now

@@
 
+### Objective
+
+The goal of this research is to understand the spatial and temporal dynamics of rent estimates in Seattle using scraped rental housing listings from Craigslist.
+I use a series of data visualizations to describe the overall datam, after which I model a panel of quarterly census tract estimates to explain the dynamics observed so far and forecast a surface of rent values for the current quarter.
+
 ### Contents
@@
 
-[`flexdashboard` for working paper's model diagnostics and forecasts](./dashboard/index.html)
+[`flexdashboard`](./dashboard/index.html) for working paper's model diagnostics and forecasts
 
-[GitHub Repository](https://github.com/hesscl/smooth-sea/)
+[GitHub Repository](https://github.com/hesscl/smooth-sea/) with source code for manuscript and results dashboard
 

### 2018-06-03 index.rmd
adding <br> to after objective on main index
typo in index, reducing footer whitespace

@@
 The goal of this research is to understand the spatial and temporal dynamics of rent estimates in Seattle using scraped rental housing listings from Craigslist.
-I use a series of data visualizations to describe the overall datam, after which I model a panel of quarterly census tract estimates to explain the dynamics observed so far and forecast a surface of rent values for the current quarter.
+I use a series of data visualizations to describe the overall data, after which I model a panel of quarterly census tract estimates to explain the dynamics observed so far and forecast a surface of rent values for the current quarter.
+
+<br>
 
@@
 
-[Working paper on estimating Seattle neighborhood rent indicators](./manuscript/index.html)
+[Working paper](./manuscript/index.html) on estimating Seattle neighborhood rent indicators
 

### 2018-06-04 dashboard/index.rmd
progress on data section

@@
     navbar:
-       - {title: "hesscl.com", href: "http://hesscl.com", align: right}
-       - {title: "Project Home", href: "http://hesscl.com/smooth-sea/index.html", align: left} 
+       - {title: "Project Home", href: "../index.html", align: right} 
     orientation: columns
     theme: bootstrap 
-    logo: "../logo.svg"
+    source: "https://github.com/hesscl/smooth-sea/dashboard"
 ---

### 2018-06-04 index.rmd
progress on data section

--- 2018-06-03 index.rmd
@@
       includes:
-        in_header: header.html
-        after_body: footer.html
+        in_header: "header.html"
+        after_body: "footer.html"
 ---
@@
 
-### Contents
+### About
 
-[Working paper on estimating Seattle neighborhood rent indicators](./manuscript/index.html)
-
-[`flexdashboard` for working paper's model diagnostics and forecasts](./dashboard/index.html)
-
-[GitHub Repository](https://github.com/hesscl/smooth-sea/)
+The goal of this research is to understand the dynamics of Seattle neighborhoods' median rent level using scraped rental housing listings from Craigslist.
+I use a series of data visualizations to describe the accumulated data and compare the Craigslist estimates against other indicators.
+After the data description, I model a panel of quarterly census tract estimates to explain the spatial and temporal dynamics observed so far.
+Models are trained on quarterly tract estimates observed through the prior quarter and forecast the current quarter's rent surface.
 
@@
 
-### See also
+### Contents
+
+[Working paper](./manuscript/index.html) on estimating Seattle neighborhood rent indicators
+
+[Results dashboard](./dashboard/index.html) for working paper's model diagnostics and forecasts
+
+[GitHub Repository](https://github.com/hesscl/smooth-sea/) with source code for manuscript and results dashboard
+
+<br>
+
+### Powered By
 
 [Helena](http://helena-lang.org) - Web automation for end users
+
+[R Markdown](https://rmarkdown.rstudio.com) - Reproducible data analysis and manuscript creation
+
+[flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/index.html) - Easy interactive dashboards for R
+
+[plot.ly](https://plot.ly/) - translate ggplot() objects to JavaScript-based htmlwidgets
+
+[GitHub Pages](https://pages.github.com/) - Publish static web pages with repositories
 
@@
 
-Email Chris Hess at [hesscl@uw.edu](mailto://hesscl@uw.edu) for more information about this research.
+Email the project team at [searent@uw.edu](mailto://searent@uw.edu) for more information about this research.

### 2018-06-04 manuscript/index.Rmd
typo in index, reducing footer whitespace
progress on data section

@@
     includes:
-      in_header: header.html
-      after_body: footer.html
+      in_header: "header.html"
+      after_body: "../footer.html"
 abstract: "Online classified ads are a key source of information about local rental housing markets.
@@
 censusRent <- read_csv(file = "../input/acsRent.csv")
-census <- inner_join(census, censusRent)
+census <- inner_join(census, censusRent) #1:1
 
@@
 <br>
+
+## Introduction
 
@@
 
+### Web scraping using Helena
+
+[Helena](http://helena-lang.org) is a high-level programming language for end users to automate web scraping as a method of data collection.
+Users write programs in an extension for Google Chrome by demonstrating the first row of data they want to collect using clicks, form interactions and highlighting values.
+The extension translates an initial demonstration into a starting script written in Helena, which a user can further modify or set to run on a regular schedule.
+Though Helena programs collect data roughly at the speed they are recorded at, users can run multiple programs in parallel to increase the volume of data collected.
+Data are stored on an AWS server where users can later query either the full table of scraped data or a subset scraped within a specified number of hours.
+
+Helena learns where features of are through the user interface (UI) rather than the underlying document object model (DOM).
+Though the HTML code of a site may regularly change over the course of regular data scraping, websites less frequently undergo drastic overhauls to what their users interact with.
+Helena also uses skip blocks to avoid re-scraping the an already-collected item.
+These two aspects of Helena provide its programs robustness to common headaches encountered with web scraping by default.
+Overall, Helena's programming-by-demonstration method reduces time needed building a scraper from scratch, giving social scientists the capacity to quickly scrape a single table of data or build ongoing data processes.
+
+For the UW team's data collection, a Helena program runs 24-hours a day and scrapes listings from the Seattle Craigslist.
+It is important to note that web scraping does violate Craigslist's Terms of Service, and that we are using all scraped data exclusively for academic research.
+For a more thorough discussion of this issue, consult page 3 of [Boeing and Waddell (2016)](https://arxiv.org/ftp/arxiv/papers/1605/1605.05397.pdf).
+
+<br>
+
+### Data Processing
+
+The key steps of any data process can be distilled into three tasks: a.) _extract_ a table of interest from a specified origin source, b.) _transform_ it to support the desire analyses and c.) _load_ the transformed data to a storage destination for future use.
+In the present case, a program extracts scraped listings observed since the database
+<br>
+
+### Data Description
+
 ```{r map_point_density, echo=FALSE, message=FALSE, warning=FALSE}
@@
 ![Figure 1. Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](../output/map_point_density.png)
-
-<br>
-
-### Web scraping with Helena
-
-<br>
-
-### Data Processing
-
-<br>
-
-### Data Description
 

### 2018-06-04 helpers/rent_lineplot.R

@@ -20,7 +20,8 @@ sumCL <- cl %>%
   mutate(catSqft = fct_reorder(catSqft, cleanSqft)) %>%
   arrange(listingDate) %>%
   select(-ends_with("10")) %>%
-  filter(!is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), !is.na(matchAddress), !is.na(matchAddress2)) %>%
+  filter(!is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), !is.na(matchAddress), !is.na(matchAddress2),
+         matchType != "Google Maps Lat/Long") %>%
   distinct(matchAddress, matchAddress2, catBeds, cleanSqft, cleanRent, .keep_all = T) %>%
   mutate(listingQtr = as.yearqtr(as.Date(listingDate))) %>%
   group_by(listingQtr) %>%

### 2018-06-04 helpers/rent_map_across.R

@@ -21,6 +21,7 @@ cl2017 <- rent_map_sea %>%
   filter(seattle==1) %>% 
   collect %>% 
   filter(listingYear >= 2017, 
+         matchType != "Google Maps Lat/Long",
          !is.na(matchAddress), !is.na(matchAddress2), !is.na(catBeds), !is.na(cleanSqft), !is.na(cleanRent),
          catBeds %in% c("0", "1", "2"), !is.na(lat) & !is.na(lng)) %>%
   distinct(matchAddress, matchAddress2, catBeds, cleanSqft, cleanRent, .keep_all = T) %>%

### 2018-06-04 manuscript/index.Rmd
work as of 6/4 morning

@@
 abstract: "Online classified ads are a key source of information about local rental housing markets.
-This research asks what temporal and spatial dynamics explain neighborhood variations in one-bedroom rents in Seattle, WA, and contributes a novel approach to forecasting quarterly neighborhood rent levels using Bayesian models estimated with integrated nested Laplace approximations (INLA).
-All models were trained on data for 2017 Q2 through 2018 Q1, with forecasts for 2018 Q2 compared against the observed-to-date data for the current quarter (i.e.
-2018 Q2).
+This research first assesses the agreement of census tract rent estimates based on scraped rental advertisements with existing data sources.
+The second component of the paper investivates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models.
+All models' posterior distribution are estimated with integrated nested Laplace approximations (INLA).
+Training data for models cover quarterly tract estimates for 2017 Q2 through 2018 Q1, while the test data are 2018 Q2 median values as observed through June 3rd.
 The top model assumes that spatial and temporal dynamics are separable, i.e.
 independent of each other, and non-interactive.
-This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $90 for the training data and $145 for the test data."
+This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $95 for the training data and $140 for the test data."
 ---
@@
 Demand for rental housing units in Seattle, WA surged in the past decade, due to relative scarcity in available housing amidst a period of historic population growth.
-Growing shares of households are unable to buy a home short of "driving till they qualify", and among households who rent a substantial share pay 30%, if not more of their incomes on their housing costs.
-The rising cost of housing therefore has a direct role in neighborhood processes of "demographic inversion" like urban gentrification and suburbanizing poverty.
-American Community Survey data have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more.
+Substantial appreciation for almost 2 years straight raised Seattle home prices to levels where [most households are no longer able to buy a home](https://www.seattletimes.com/business/real-estate/seattle-area-homebuyers-needed-11000-pay-raise-in-past-year-to-keep-up-with-rising-prices/?utm_source=The+Seattle+Times&utm_campaign=de86e2d581-Morning_Brief_11_28_2017&utm_medium=email&utm_term=0_5beb38b61e-de86e2d581-122931741) short of "driving till they qualify".
+More and more households must rent to live in Seattle, and a [almost half](http://www.jchs.harvard.edu/ARH_2017_cost_burdens_by_metro) of these renter households pay 30% of their incomes on housing costs, if not more.
+Beyond individual households, the rising cost of housing in Seattle and in other cities has a role in neighborhood processes of "demographic inversion" like urban gentrification and [suburbanizing poverty](https://www.brookings.edu/testimonies/the-changing-geography-of-us-poverty/) too.
+
+[American Community Survey (ACS)](https://www.census.gov/programs-surveys/acs/) data have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more.
 This publication lag for ACS data prohibits analyzing any market dynamics beyond long-term trends, particularly for estimates of smaller areas of aggregation like neighborhoods where ACS uses 5-year rolling-average methods as well.
@@
 
-This study aims to demonstrate why scraped rental housing listings are a valuable source of data in subject areas such as housing studies and neighborhood demography through a set of descriptive analyses and models of quarterly rent levels among census tracts in Seattle, WA.
+This study demonstrates why scraped rental housing listings are a valuable source of data in subject areas such as housing studies and neighborhood demography.
+I draw on a set of descriptive analyses and models of quarterly rent levels among census tracts in Seattle, WA for this research objective.
 Summaries of the Craigslist data show agreement with existing sources in their distribution but do suggest higher median levels for most neighborhoods, likely given the listings' recent collection relative to other data.
+The smoothing models use the panel of quarterly neighborhood estimates' spatial and temporal structure to understand the underlying generative process, finding that a combination of local smoothing and autoregression (order 1) predict future levels with an average error of about $140 or 8%.
+
 In a more general sense, the methods and frameworks used in this study also offer a template for efficiently collecting web data, conducting reproducible research, and publishing interactive results to GitHub.
@@
 
-Existing public data sources like the American Community Survey or decennial census theoretically have two sources of error relative to the ``true'' median rent in a given overall market and to an even greater degree among neighborhood approximations like census tracts.
+Existing public data are essentially limited to the American Community Survey (ACS) following the cessation of long-form decennial censuses in 2000.
+ACS data theoretically have two sources of error relative to the ``true'' median rent in a given overall market and to an even greater degree among neighborhood approximations like census tracts.
 Census sources of rental information cannot provide insight into any change in rental cost associated with newer developments becoming available (at least not until considerable time has already passed) given the temporal delay in data publication (i.e.
@@
 Though Helena programs collect data roughly at the speed they are recorded at, users can run multiple programs in parallel to increase the volume of data collected.
-Data are stored on an AWS server where users can later query either the full table of scraped data or a subset scraped within a specified number of hours.
+Data accumulate on a remote server (soon-to-be AWS) where users can later query either the full table of scraped data or a subset scraped within a specified number of hours.
 
@@
 The key steps of any data process can be distilled into three tasks: a.) _extract_ a table of interest from a specified origin source, b.) _transform_ it to support the desire analyses and c.) _load_ the transformed data to a storage destination for future use.
-In the present case, a program extracts scraped listings observed since the database
+
+In the present case, a program extracts scraped listings observed since the data updated by sending a download request to the scraper's accumulated data in the form of a URL with a suffix indicating the number of hours that passed since the last processing.
+The processing routine then transforms the data by cleaning up some of the scraped fields and appending spatial information through a multi-step geocoding proces.
+If address information was successfully scraped from the listing metadata, the script makes a call to Google Maps Geocoding API using this listing feature and a constraint to search within Washington state.
+When this first step is not successful at geocoding to a specific address or cross-streets, or where there is no address in the listing metadata, we use SmartyStreet's Address Extraction API on the listing's full scraped text to see if an address was mentioned there.
+We are able to geocode 95\% of listings to some degree of precision, with the vast majority precise at least to Zip9 geography.
+All listings used in the following analyses are geocoded with precision at least to Zip9 geography.
+After geocoding the data and ensuring there are no duplicates among the table of new listings, the script loads the processed listings into the existing database.
+
 <br>
@@
 
+```{r n_heatmap, echo=FALSE, message=FALSE, warning=FALSE}
+#determine config (dev or github)
+if(file.exists("../../data/cl/craigslistDB.sqlite")){
+  source("../helpers/n_heatmap.R") #if config, refresh plot pdf
+}
+```
+
+![Figure 2. Heatmap of N Seattle Craigslist one-bedroom listings scraped per day time-series.](../output/n_heatmap.png)
+
 ```{r rent_lineplot, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-![Figure 2.
+![Figure 3.
 Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018.
@@
 
-![Figure 3.
+![Figure 4.
 Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates.
@@
 
-![Figure 4.
+![Figure 5.
 Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott.
@@
 ## Models
+
+- Need to build tables from output
+- Need to do write up for tables
+- Need to add graphic for predicted values
 
@@
 
-`r knitr::kable(fit_stats)`
+`r knitr::kable(fit_stats %>% select(ends_with("rmse")))`
+
+`r knitr::kable(fit_stats %>% select(ends_with("mae")))`
 

### 2018-06-04 helpers/n_heatmap.R

@@ -1,0 +1,75 @@
+
+#connect to db
+DB <- dbConnect(SQLite(), dbname="R:/Project/seattle_rental_market/data/cl/craigslistDB.sqlite")
+cl <- tbl(DB, "clean") #clean listing table
+
+#compute tract aggregates for CL listing count
+sea_cl <- cl %>%
+  filter(!is.na(GISJOIN), seattle == 1) %>% #only listings with valid Bed/Rent
+  dplyr::select(listingMoYr, listingDate, GISJOIN, seattle, matchAddress, matchAddress2, matchType, cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
+  collect %>% #bring db query into memory
+  filter(GISJOIN %in% sea_shp@data$GISJOIN) %>% #filter to KC only (db has metro area)
+  mutate(listingDate = as.Date(listingDate),
+         listingQtr = as.yearqtr(listingDate)) %>%
+  filter(listingQtr >= "2017 Q2", !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft),
+         cleanBeds == 1) %>% 
+  filter(matchType != "Google Maps Lat/Long") %>% #no Google lat/long only geocodes
+  distinct(matchAddress, matchAddress2, cleanBeds, cleanRent, cleanSqft, .keep_all = T)
+dbDisconnect(DB)
+
+computeCal <- function(x){
+  y <- x %>%
+    group_by(listingDate) %>%
+    summarize(n = n())
+  
+  fullDates <- seq(from = min(y$listingDate), to = max(y$listingDate), by = 1)
+  index <- seq(from = 1, length.out = length(fullDates), by = 1)
+  dateTable <- cbind.data.frame(fullDates, index)
+  
+  n <- NULL
+  for(i in 1:length(fullDates)){
+    if(fullDates[i] %in% y$listingDate){
+      n[i] <- y$n[y$listingDate == fullDates[i]]
+    } else{
+      n[i] <- NA
+    }
+  }
+  n <- cbind.data.frame(n, fullDates)
+}
+
+
+cal <- computeCal(sea_cl) %>%
+  mutate(weekdayf = weekdays(fullDates) %>% as.factor,
+         monthf = month(fullDates, label = T),
+         yearmon = as.yearmon(fullDates),
+         year = year(fullDates),
+         week = as.numeric(format(fullDates,"%W")))
+sum(is.na(cal$n))/nrow(cal)
+
[…]

### 2018-06-04 dashboard/index.rmd
more progress on manuscript, working on data description, tidying up other pages a little

@@
     theme: bootstrap 
-    source: "https://github.com/hesscl/smooth-sea/dashboard"
+    source: "https://github.com/hesscl/smooth-sea/tree/master/dashboard"
 ---

### 2018-06-04 index.rmd
more progress on manuscript, working on data description, tidying up other pages a little

@@
 
-[GitHub Repository](https://github.com/hesscl/smooth-sea/) with source code for manuscript and results dashboard
+[GitHub repository](https://github.com/hesscl/smooth-sea/) with source code for manuscript and results dashboard
 
@@
 
-[plot.ly](https://plot.ly/) - translate ggplot() objects to JavaScript-based htmlwidgets
+[plot.ly](https://plot.ly/) - translate ggplot() objects to JavaScript htmlwidgets
 

### 2018-06-04 manuscript/index.Rmd
more progress on manuscript, working on data description, tidying up other pages a little

@@
 Substantial appreciation for almost 2 years straight raised Seattle home prices to levels where [most households are no longer able to buy a home](https://www.seattletimes.com/business/real-estate/seattle-area-homebuyers-needed-11000-pay-raise-in-past-year-to-keep-up-with-rising-prices/?utm_source=The+Seattle+Times&utm_campaign=de86e2d581-Morning_Brief_11_28_2017&utm_medium=email&utm_term=0_5beb38b61e-de86e2d581-122931741) short of "driving till they qualify".
-More and more households must rent to live in Seattle, and a [almost half](http://www.jchs.harvard.edu/ARH_2017_cost_burdens_by_metro) of these renter households pay 30% of their incomes on housing costs, if not more.
-Beyond individual households, the rising cost of housing in Seattle and in other cities has a role in neighborhood processes of "demographic inversion" like urban gentrification and [suburbanizing poverty](https://www.brookings.edu/testimonies/the-changing-geography-of-us-poverty/) too.
+More and more households must therefore rent to live in Seattle, and [almost half](http://www.jchs.harvard.edu/ARH_2017_cost_burdens_by_metro) of these renter households pay 30% of their incomes on housing costs, if not more.
+Given the universal need for housing, the individual dynamics related to rising housing costs have a role in shaping neighborhood processes of "demographic inversion" like urban gentrification and [suburbanizing poverty](https://www.brookings.edu/testimonies/the-changing-geography-of-us-poverty/) too.
 
@@
 The smoothing models use the panel of quarterly neighborhood estimates' spatial and temporal structure to understand the underlying generative process, finding that a combination of local smoothing and autoregression (order 1) predict future levels with an average error of about $140 or 8%.
+In a more general sense, the methods and frameworks used in this study also offer a template for efficiently collecting web data, conducting reproducible research, and publishing interactive results to GitHub.
 
-In a more general sense, the methods and frameworks used in this study also offer a template for efficiently collecting web data, conducting reproducible research, and publishing interactive results to GitHub.
 I start by briefly reviewing some background relevant to studying neighborhood rental housing.
@@
 
-Existing public data are essentially limited to the American Community Survey (ACS) following the cessation of long-form decennial censuses in 2000.
-ACS data theoretically have two sources of error relative to the ``true'' median rent in a given overall market and to an even greater degree among neighborhood approximations like census tracts.
+Existing public data are essentially limited to the American Community Survey (ACS) and [American Housing Survey (AHS)](https://www.census.gov/programs-surveys/ahs.html) following the cessation of long-form decennial censuses in 2000.
+These data theoretically have two sources of error relative to the ``true'' current median rent for a market and  neighborhood approximations like census tracts.
 Census sources of rental information cannot provide insight into any change in rental cost associated with newer developments becoming available (at least not until considerable time has already passed) given the temporal delay in data publication (i.e.
@@
 
-For-profit data sources play a role in filling the information gap about rental housing costs, though they tend to possess their own unique shortcomings.
+For-profit data sources play a role in filling the information gap about rental housing costs, but can still have their own limitations.
 These data sources tend to rely on survey methodologies for data collection, and ask landlords in a given market to outline the average cost for units of a given bedroom size or other specification.
+The quality of these data therefore depend on their capacity to adequately sample housing units across all property sizes, building qualities and locations, and the "mom and pop" end of the rental market spectrum is difficult to estimate in size, let alone survey sufficiently.
 In exchange for providing detailed information for their database, the rental data provider may give participating landlords and property management companies access to a slice of information generated via accumulated data.
@@
 
-The key steps of any data process can be distilled into three tasks: a.) _extract_ a table of interest from a specified origin source, b.) _transform_ it to support the desire analyses and c.) _load_ the transformed data to a storage destination for future use.
+The key steps of any data process can be distilled into three tasks: a.) _extract_ a table of interest from a specified origin source, b.) _transform_ it to support the desired analyses and c.) _load_ the transformed data to a storage destination for future use.
 
@@
 
+We have collected `r sum(tract$nListings)` Seattle rental housing listings from Craigslist over the course of our data collection (`r min(tract$listingQtr)` to `r max(tract$listingQtr)`).
+Of the overall corpus of listings we've collected, `r sum(tract$n1B)` (about `r round(sum(tract$n1B)/sum(tract$nListings), 2)*100`%) were one-bedroom listings.
+
+The following series of graphics summarize the listing data collected to date and compare rent estimates from these data to ACS data and Dupre+Scott.
+Wherever possible, visualizations focus on one-bedroom units to make comparisons between data for the most common unit type in Seattle.
+However, for Figure 2 data limitations require that the comparison is all housing units.
+
+<br>
+<hr>
+<br>
+
+Figure 1 shows a reference map of Seattle.
+In the right pane, each point indicates a one-bedroom listing collected at some point over our study's span.
+The polygons overlaid on higher-density neighborhoods around Downtown, South Lake Union and Ballard correspond to particularly strong Craigslist listing activity.
+
 ```{r map_point_density, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-![Figure 1. Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](../output/map_point_density.png)
+![**Figure 1.** Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](../output/map_point_density.png)
+
+<br>
+<hr>
+<br>
+
+Figure 2 shows a heatmap for the time-series of one-bedroom listings per day.
+The grey cells indicate days where our scraper was offline.
+Outages were more common in the earliest phases of work, but now are relatively rare.
+
+At the end of February we implemented a procedure to have the Helena program immediately restart scraping (i.e.
+re-query the available listings as a new index to work through) in order to ensure any and all listings are scraped each day.
+The net impact of this change is fairly evident in the time-series heatmap.
+We now see more temporal structure to the listing count for each day---relatively fewer listings are posted on Sundays, though this could be related to greater listing on Saturdays (i.e.
+everything is posted at the beginning of the weekend in order to maximize weekend housing searchers' likelihoods of seeing it).
+Perhaps unsurprisingly, there were dips in listing activity on prominent holiday's like New Year's Day, Easter and Memorial Day.
 
@@
 
-![Figure 2. Heatmap of N Seattle Craigslist one-bedroom listings scraped per day time-series.](../output/n_heatmap.png)
+![**Figure 2.** Heatmap of N Seattle Craigslist one-bedroom listings scraped per day time-series.](../output/n_heatmap.png)
+
+<br>
+<hr>
+<br>
+
+Figure 3 is a line graphic indicating the Seattle-wide median rent for all sizes of rental units, with the longest set of data (Dupre+Scott) reaching back to 1997 to show long-term dynamics of Seattle rental housing costs.
+The first takeaway from the graphic is the large spike in rent levels following 2010, with a similar trajectory followed in the estimates from Dupre+Scott and the ACS.
 
@@
 
-![Figure 3.
-Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018.
+![**Figure 3.** Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018.
 Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual and ACS are annual (i.e.
 respective 1-Year Estimates for each year period).](../output/rent_lineplot.png)
+
+<br>
+<hr>
+<br>
 
@@
 
-![Figure 4.
-Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates.
+![**Figure 4.** Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates.
 Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled.](../output/rent_map_across_1B.png)
+
+<br>
+<hr>
+<br>
 
@@
 
-![Figure 5.
-Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott.
+![**Figure 5.** Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott.
 Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.](../output/rent_scatter_across_1B.png)
@@
 ## Models
+
+<br>
+
+## Results
 
@@
 
-`r knitr::kable(fit_stats %>% select(ends_with("rmse")))`
+`r knitr::kable(fit_stats %>% select(train_test, ends_with("rmse")))`
 
-`r knitr::kable(fit_stats %>% select(ends_with("mae")))`
-
-<br>
-
-## Results
+`r knitr::kable(fit_stats %>% select(train_test, ends_with("mae")))`
 

### 2018-06-04 helpers/load_data.R

@@ -17,7 +17,7 @@ load_data <- function(){
                     cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
       mutate(listingDate = as.Date(listingDate),
              listingQtr = as.yearqtr(listingDate)) %>%
-      filter(cleanBeds %in% c(0, 1, 2, 3), listingQtr >= "2017 Q2") %>%
+      filter(listingQtr >= "2017 Q2") %>%
       group_by(GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
       summarize(nListings = n(),
                 n1B = sum(cleanBeds == 1),

### 2018-06-04 helpers/n_heatmap.R

@@ -59,9 +59,7 @@ ggplot(cal, aes(x = monthweek, y = weekdayf, fill = n)) +
   geom_tile(color = "white") +
   facet_grid(year~monthf) +
   theme_minimal() +
-  theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
-        panel.spacing = unit(.1, "lines"),
-        axis.text = element_text(size = 6),
+  theme(axis.text = element_text(size = 6),
         legend.text = element_text(size = 8),
         legend.title = element_text(size = 10),
         axis.title = element_text(size = 8),
@@ -71,5 +69,5 @@ ggplot(cal, aes(x = monthweek, y = weekdayf, fill = n)) +
   ylab("Day of Week\n") +
   labs(fill = "N Listings") +
   ggsave(filename = "../output/n_heatmap.png",
-         width = 6, height = 3)
+         width = 6, height = 4, dpi = 300)
 

### 2018-06-04 manuscript/index.Rmd
more progress on manuscript

@@
 <br>
+<hr>
+<br>
 
@@
 Helena learns where features of are through the user interface (UI) rather than the underlying document object model (DOM).
-Though the HTML code of a site may regularly change over the course of regular data scraping, websites less frequently undergo drastic overhauls to what their users interact with.
+Though the HTML code of a site may regularly change over the course of regular data scraping, websites less frequently overhaul what their users actually interact with.
 Helena also uses skip blocks to avoid re-scraping the an already-collected item.
@@
 
+<br>
+<hr>
 <br>
@@
 
-Figure 3 is a line graphic indicating the Seattle-wide median rent for all sizes of rental units, with the longest set of data (Dupre+Scott) reaching back to 1997 to show long-term dynamics of Seattle rental housing costs.
+Figure 3 is a line graphic indicating agreement across data for Seattle-wide median rent (all sizes of rental units).
+The longest set of data (Dupre+Scott) cover 20 years and show long-term dynamics of Seattle rental housing costs in terms of contract rent.
+Dupre+Scott released semi-annual summaries for their 20+ unit data, so there are two data points for each year (Spring and Fall).
+ACS 1-Year Estimates show the median renter-occupied housing unit asking rent.
+Craigslist shows quarterly estimates for all listings in Seattle.
+
 The first takeaway from the graphic is the large spike in rent levels following 2010, with a similar trajectory followed in the estimates from Dupre+Scott and the ACS.
+Throughout the span of Dupre+Scott and ACS data, these two indicators differ by about $100-150 at any given point, though the disagreement is at its largest at the ACS data's current endpoint of 2016.
+Given that the Dupre+Scott data and ACS data both sample rents from already-leased-up units (exclusively in the case of ACS), a plausible explanation for this systematic difference is that Dupre+Scott rent estimates reflect only the population of rental units in larger structures (whereas ACS is theoretically sampling rental units of all property sizes and characteristics.
+
+Only the Craigslist and Dupre+Scott are particularly useful for understanding more recent median levels for the Seattle rental housing market.
 

### 2018-06-04 helpers/rent_map_across.R

@@ -98,7 +98,7 @@ acs1B <- rent_map_acs %>%
   select(GISJOIN, data, catBeds, rent, nHU)
 
 acs2B <- rent_map_acs %>%
-  mutate(rent = AGPZE003,
+  mutate(rent = AGPZE004,
          catBeds = "2",
          data = "ACS 2012-2016",
          nHU = NA) %>%

### 2018-06-04 manuscript/index.Rmd
changing T to start in 2017 Q1 now

@@
 
-### Web scraping using Helena
+### Web Scraping using Helena
 
@@
 We have collected `r sum(tract$nListings)` Seattle rental housing listings from Craigslist over the course of our data collection (`r min(tract$listingQtr)` to `r max(tract$listingQtr)`).
-Of the overall corpus of listings we've collected, `r sum(tract$n1B)` (about `r round(sum(tract$n1B)/sum(tract$nListings), 2)*100`%) were one-bedroom listings.
+Among the overall listings we've collected, `r sum(tract$n1B)` (about `r round(sum(tract$n1B)/sum(tract$nListings), 2)*100`%) were one-bedroom listings.
 
 The following series of graphics summarize the listing data collected to date and compare rent estimates from these data to ACS data and Dupre+Scott.
-Wherever possible, visualizations focus on one-bedroom units to make comparisons between data for the most common unit type in Seattle.
-However, for Figure 2 data limitations require that the comparison is all housing units.
+In an attempt to compare summaries about like rental units as much as possible, visualizations focus on one-bedroom units in all but one instance.
+One-bedroom units are tied with the combined share of two and three bedroom units for the [most common unit bedroom size](https://factfinder.census.gov/bkmk/table/1.0/en/ACS/16_1YR/S2504/1600000US5363000) in Seattle according to 2016 ACS data.
+For Figure 2, data limitations require that the comparison is all housing units since ACS only recently (i.e.
+2015-2016) started publishing bedroom-specific rent levels.
 
@@
 
-Figure 1 shows a reference map of Seattle.
-In the right pane, each point indicates a one-bedroom listing collected at some point over our study's span.
+Figure 1 shows a reference map of Seattle on the left.
+Opposite this is a visualization of the scraped Seattle rental listing data.
+Each point indicates a one-bedroom listing collected at some point over our study's span.
 The polygons overlaid on higher-density neighborhoods around Downtown, South Lake Union and Ballard correspond to particularly strong Craigslist listing activity.
@@
 
-At the end of February we implemented a procedure to have the Helena program immediately restart scraping (i.e.
+At the end of February 2018 we implemented a procedure to have the Helena program immediately restart scraping (i.e.
 re-query the available listings as a new index to work through) in order to ensure any and all listings are scraped each day.
@@
 
-Only the Craigslist and Dupre+Scott are particularly useful for understanding more recent median levels for the Seattle rental housing market.
-
 ```{r rent_lineplot, echo=FALSE, message=FALSE, warning=FALSE}
@@
 ![**Figure 3.** Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018.
-Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual and ACS are annual (i.e.
+Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual (March and September), and ACS are annual (i.e.
 respective 1-Year Estimates for each year period).](../output/rent_lineplot.png)
+
+Only the Craigslist and Dupre+Scott can offer insight into more recent median levels for the Seattle rental housing market.
+Both data suggest continued increases in the marketwide median rent into Fall 2017, though the Craigslist slope is about twice in its size.
+Some of the spike observed between the second and third data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate upward in favor of regularly posted commerical listings.
+Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability.
+This volatility in Craigslist asking rents could also be brought on by recently-developed buildings in "lease-up" adding a glut of expensive listings to the quarters' samples.
+
+However, by the final point of Dupre+Scott's series (Fall 2017, corresponding to September), the difference compared to the Craigslist median was about $75.
+Mirroring other accounts about a decline in rents during Winter 2018, the Craigslist data showed a fairly precipitous decline between
 
@@
 <br>
+
+The focus of the data description now turns towards estimating median rents for specific neighborhoods.
+While zip codes, unofficial City neighborhoods and even arbitrary grid geographies can all be used to aggregate (or in a graphical sense, rasterize) micro-data about rental housing,  .
+Figure 4
+
+[Reference 2000 Census tract map for Seattle](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2000CensusSeattleTractMap.pdf)
+
+[Reference 2010 census tract map for Seattle](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2010CensusTractsandZipCodeBoundaries.pdf)
 

### 2018-06-04 helpers/load_data.R

@@ -17,7 +17,7 @@ load_data <- function(){
                     cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
       mutate(listingDate = as.Date(listingDate),
              listingQtr = as.yearqtr(listingDate)) %>%
-      filter(listingQtr >= "2017 Q2") %>%
+      filter(listingQtr >= "2017 Q1") %>%
       group_by(GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
       summarize(nListings = n(),
                 n1B = sum(cleanBeds == 1),

### 2018-06-04 helpers/n_heatmap.R

@@ -11,7 +11,7 @@ sea_cl <- cl %>%
   filter(GISJOIN %in% sea_shp@data$GISJOIN) %>% #filter to KC only (db has metro area)
   mutate(listingDate = as.Date(listingDate),
          listingQtr = as.yearqtr(listingDate)) %>%
-  filter(listingQtr >= "2017 Q2", !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft),
+  filter(listingQtr >= "2017 Q1", !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft),
          cleanBeds == 1) %>% 
   filter(matchType != "Google Maps Lat/Long") %>% #no Google lat/long only geocodes
   distinct(matchAddress, matchAddress2, cleanBeds, cleanRent, cleanSqft, .keep_all = T)

### 2018-06-04 helpers/rent_lineplot.R

@@ -25,10 +25,10 @@ sumCL <- cl %>%
   distinct(matchAddress, matchAddress2, catBeds, cleanSqft, cleanRent, .keep_all = T) %>%
   mutate(listingQtr = as.yearqtr(as.Date(listingDate))) %>%
   group_by(listingQtr) %>%
+  filter(listingQtr >= "2017 Q1") %>%
   summarize(Rent = median(cleanRent, na.rm=T),
             nHU = n()) %>%
   rename(moYr = listingQtr) %>%
-  filter(nHU >= 200) %>%
   mutate(dataSrc = "Craigslist") %>%
   select(moYr, Rent, dataSrc) %>%
   mutate(moYr = as.Date(moYr))

### 2018-06-04 helpers/rent_scatter_across.R

@@ -21,7 +21,7 @@ scatter_cl <- scatter_sea %>%
                 cleanBeds, cleanRent, cleanSqft, lat, lng) %>% #SELECT these columns
   mutate(listingDate = as.Date(listingDate),
          listingQtr = as.yearqtr(listingDate)) %>%
-  filter(cleanBeds %in% c(1), listingQtr >= "2017 Q2")
+  filter(cleanBeds %in% c(1), listingQtr >= "2017 Q1")
 
 coordinates(scatter_cl) <- cbind(scatter_cl$lng, scatter_cl$lat)
 

### 2018-06-05 dashboard/index.rmd
progress through 6/4

@@
 temporal and spatial).
-All of the candidate models' posterior distributions are estimated with [integrated nested Laplace approximations](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) (INLA) using the default, weakly-informative priors for all model hyperparameters.
+All of the candidate models' posterior distributions are estimated with [integrated nested Laplace approximations (INLA)](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) using the default, weakly-informative priors for all model hyperparameters.
 Throughout the following analyses, the training data are `r min(sea_df$listingQtr)` up to the prior quarter (i.e.
@@
   mutate(var = paste0(as.yearqtr(Sys.Date()), "-to-Date")) %>%
-  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, spt_Med, NAMELSAD10, n1B)
+  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, bym_Med, NAMELSAD10, n1B)
 
@@
   geom_polygon(color = "grey85", lwd = .05) +
-  scale_fill_viridis_c(limits = c(750, 2500), labels = scales::dollar) +
+  scale_fill_viridis_c(limits = c(750, 2500), labels = scales::dollar, option = "A") +
   coord_quickmap() +
@@
   scale_x_continuous(labels = scales::dollar) +
-  scale_color_viridis_d(end = .8) +
+  scale_color_viridis_d(end = .8, option = "A") +
   xlab("\nMedian 1B Rent") +
@@
   geom_point() +
-  scale_color_gradient2_tableau() +
+  scale_color_viridis_c(option = "A") +
   scale_y_continuous(labels = scales::dollar) +
@@
   geom_polygon(color = "grey85", lwd = .05) +
-  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
+  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80", option = "A") +
   coord_quickmap() +

### 2018-06-05 manuscript/index.Rmd
progress through 6/4

@@
 independent of each other, and non-interactive.
-This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $95 for the training data and $140 for the test data."
+This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $100 for the training data and $140 for the test data."
 ---
@@
 library(latex2exp)
+library(plotly)
 
@@
 
-[American Community Survey (ACS)](https://www.census.gov/programs-surveys/acs/) data have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more.
-This publication lag for ACS data prohibits analyzing any market dynamics beyond long-term trends, particularly for estimates of smaller areas of aggregation like neighborhoods where ACS uses 5-year rolling-average methods as well.
+Census data have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more.
+This publication lag for these data prohibits analyzing market dynamics beyond long-term trends, particularly for smaller areas of aggregation like neighborhoods where the estimate may be an average of survey responses from a 5 year period.
 While more up-to-date options for neighborhood rent data exist through privately-held companies, such data can be somewhat of a black box in their creation and may be biased towards the costs of corporate-held properties.
@@
 Summaries of the Craigslist data show agreement with existing sources in their distribution but do suggest higher median levels for most neighborhoods, likely given the listings' recent collection relative to other data.
-The smoothing models use the panel of quarterly neighborhood estimates' spatial and temporal structure to understand the underlying generative process, finding that a combination of local smoothing and autoregression (order 1) predict future levels with an average error of about $140 or 8%.
+The models use a quarterly panel of observed neighborhood median one-bedroom rents and produce temporally and spatially smoothed predictions about neighborhood rent levels.
+The best model specification in terms of forecasting Q2 2018 based on Q1 2017 and Q1 2018 has a combination of local spatial smoothing and autoregression (order 1), with an average forecast error of about $140 or 8% for what.
 In a more general sense, the methods and frameworks used in this study also offer a template for efficiently collecting web data, conducting reproducible research, and publishing interactive results to GitHub.
@@
 
-Existing public data are essentially limited to the American Community Survey (ACS) and [American Housing Survey (AHS)](https://www.census.gov/programs-surveys/ahs.html) following the cessation of long-form decennial censuses in 2000.
+Existing public data are essentially limited to the [American Community Survey (ACS)](https://www.census.gov/programs-surveys/acs/) and [American Housing Survey (AHS)](https://www.census.gov/programs-surveys/ahs.html) following the cessation of long-form decennial censuses in 2000.
 These data theoretically have two sources of error relative to the ``true'' current median rent for a market and  neighborhood approximations like census tracts.
@@
 
+<br>
+
 ```{r map_point_density, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-![**Figure 1.** Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](../output/map_point_density.png)
+![**Figure 1.** _Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right)._](../output/map_point_density.png)
 
@@
 
-Figure 2 shows a heatmap for the time-series of one-bedroom listings per day.
+Figure 2 is a heatmap for the time-series of one-bedroom listings per day.
 The grey cells indicate days where our scraper was offline.
@@
 everything is posted at the beginning of the weekend in order to maximize weekend housing searchers' likelihoods of seeing it).
-Perhaps unsurprisingly, there were dips in listing activity on prominent holiday's like New Year's Day, Easter and Memorial Day.
+Perhaps unsurprisingly, there were dips in listing activity on prominent holidays like Christmas, New Year's Day, and Memorial Day.
+
+<br>
 
@@
 
-![**Figure 2.** Heatmap of N Seattle Craigslist one-bedroom listings scraped per day time-series.](../output/n_heatmap.png)
+![**Figure 2.** _Heatmap of N Seattle Craigslist one-bedroom listings scraped per day time-series._](../output/n_heatmap.png)
 
@@
 
+<br>
+
 ```{r rent_lineplot, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-![**Figure 3.** Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018.
+![**Figure 3.** _Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018.
 Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual (March and September), and ACS are annual (i.e.
-respective 1-Year Estimates for each year period).](../output/rent_lineplot.png)
+respective 1-Year Estimates for each year period)._](../output/rent_lineplot.png)
+
+<br>
 
 Only the Craigslist and Dupre+Scott can offer insight into more recent median levels for the Seattle rental housing market.
-Both data suggest continued increases in the marketwide median rent into Fall 2017, though the Craigslist slope is about twice in its size.
-Some of the spike observed between the second and third data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate upward in favor of regularly posted commerical listings.
+Both data suggest continued increases in the marketwide median rent for the first half of 2017, though the Craigslist slope is about twice in its size.
+Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for 2017 Q2 upward in favor of regularly posted commerical listings.
 Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability.
@@
 The focus of the data description now turns towards estimating median rents for specific neighborhoods.
-While zip codes, unofficial City neighborhoods and even arbitrary grid geographies can all be used to aggregate (or in a graphical sense, rasterize) micro-data about rental housing,  .
-Figure 4
+While zip codes, unofficial City neighborhoods and even arbitrary grid geographies can all be used to aggregate micro-data about rental housing, we compare census tract estimates because they roughly approximate neighborhood-size units and can be linked with a variety of other data in Seattle and more generally.
+The following links provide reference maps for Seattle Census tracts ([2000](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2000CensusSeattleTractMap.pdf) and [2010](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2010CensusTractsandZipCodeBoundaries.pdf))
 
-[Reference 2000 Census tract map for Seattle](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2000CensusSeattleTractMap.pdf)
+In the following graphic, the left most pane shows median one-bedroom asking rent estimates using data collected from Craigslist in 2017-2018.
+The middle pane shows Dupre+Scott tract median 1B contract rent estimates for one-bedrooms in 20+ unit apartment structures according to their September 2017 (Fall) survey of local landlords and property managers.
+These data were only available in 2000 Census geography, so the neighborhood units differ slightly (124 tracts versus 134 tracts) in this map compared to the other two.
+Though less than ideal, this limitation does not drastically alter the comparison between data.
+Finally, on the right are ACS 2012-2016 5-Year Estimates (the most recent ACS small-area data product available).
+In all map panes, grey cells denote missing values.
 
-[Reference 2010 census tract map for Seattle](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2010CensusTractsandZipCodeBoundaries.pdf)
+<br>
 
@@
 
-![**Figure 4.** Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates.
-Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled.](../output/rent_map_across_1B.png)
+![**Figure 4.** _Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates.
+Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled._](../output/rent_map_across_1B.png)
+
+<br>
+
+Both the Craigslist and Dupre+Scott data largely agree on the spatial distribution of one-bedroom rents.
+The similarity is greatest in the downtown and higher-density neighborhoods where the stock of corporate-held rental housing units theoretically accounts for a most of a tract's rental housing units (i.e.
+Dupre+Scott closer to a census if they sample all 20+ unit properties).
+Nonetheless, the systematic difference is such that Craigslist levels are typically at least a little higher.
+A detailed comparison of the Craigslist and Dupre+Scott estimates follows in the next figure where we directly compare 2000-definition census tract estimates from Craigslist and Dupre+Scott.
+
+Though the spatial distribution of the ACS estimates resembles that of the other two data sources, the values are nonethless systematically lower (in most cases by hundreds of dollars).
+The ACS measures therefore give insight into the relative ranking of neighborhoods in terms of rental costs, but are limited for studying neighborhood dynamics outside of long-run trend.
 
@@
 <hr>
+<br>
+
+Figure 5 shows a scatterplot/bubbpleplot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 75 tracts where comparable estimates exist.
+The purple dashed line indicates the loess smoother fit between the measures.
+Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and good agreement between the measures (mean absolute difference = $186).
+The Craigslist measure reflects asking rents and the Dupre+Scott measure reflects contract rents for a neighborhood (i.e.
+the asking rent less an incentives a landlord offers).
+Given the currently low vacancy rates in most properties, the only neighborhoods where there would theoretically be more salience to the asking vs contract rent distinction (i.e.
+a wider gap in these indicators) are at the high end of the distribution where demand for luxury apartments has slackened some.
+At the high-end of the neighborhood rent distribution, however, Craigslist and Dupre+Scott are in arguably the strongest agreement.
+
 <br>
@@
 
-![**Figure 5.** Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott.
-Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.](../output/rent_scatter_across_1B.png)
+![**Figure 5.** _Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott.
+Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.
+Purple dashed line indicates loess fit._](../output/rent_scatter_across_1B.png)
+
+<br>
+
+An alternative explanation fpr the difference between measures comes in considering the units these properties are sampling from in a given neighborhood.
+The greater difference between the measures among lower-rent neighborhoods (with Craigslist estimating these as more expensive than Dupre+Scott) provides some evidence for this theory.
+These tracts are found towards the North and South edges of the city, and generally are more residential neighborhoods where 20+ unit apartment complexes are not the entire stock of rental unit properties.
+In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf).
+
+To the extent that these smaller property units may tend to be newer, larger in housing unit square footage or offer added amenities, the true value of rent for the neighborhood will be higher than what it is estimated at using 20+ unit properties alone.
+Similarly, the more that smaller properties account for the overall population of rental housing units, the more likely that higher values indicated in Craigslist are not erroneous but instead reflective of a broader set of residential units.
 
@@
 
-<br>
+The following models learn from the Craigslist data accumulated up till the prior quarter (i.e.
+`r as.yearqtr(Sys.Date())-1`) to accomplish a few different goals.
+The most prominent goal is to use the panel data's temporal and spatial structure to understand variation in rent levels over time and across space.
+A secondary goal is to forecast the Seattle neighborhood rent surface for the current quarter.
+Finally, the following models allow for relatively straightforward imputation of missing values based on time period and the rent values of proximate tracts (though there are only two tracts where we have never observed a one-bedroom listing).
+
+I assume that the log of the median one-bedroom rent for Seattle neighborhoods is normally distributed and estimate the posterior distribution of parameters $\theta_{it}$ using [integrated nested Laplace approximations (INLA)](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) with the default, weakly-informative priors for all random-effects.
+
+Formally, models assume:
+
+$log(Y_{it})|\theta_{it} \sim \mbox{Normal}(\theta_{it})$ with:
+
+Model                    | Abbr.   | Formula
+------------------------ | ------- | ----------------------------------------------
+Fixed Intercept          | int     | $\theta_{it} = \beta_0 + Qtr$
+Non-Spatial              | ns      | $\theta_{it} = \beta_0 + \epsilon_i + Qtr$
+Non-Spatial AR(1)        | nsar1   | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i$
+BYM Spatial AR(1)        | bym     | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i + S_i.$
+BYM Spatiotemporal AR(1) | spt     | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i + S_i. + \epsilon_{it}$
+
+The Fixed Intercept model (int) uses fixed effects for quarters only.
+The Non-Spatial (ns) model adds random noise (i.i.d.) tract random-effects.
+The Non-Spatial AR(1) model (nsar1) uses AR(1) process to model the temporal dynamics (formally, $\epsilon_1 \sim \mbox{Normal}(0, (\tau(1-\rho^2))^{-1}))$ where $\epsilon_t = \rho x_{t-1} + \epsilon_t$).
+This model includes a random noise term for each quarter too.
+
+The BYM Spatial with AR(1) is the [Besags-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466) intrinsic conditional autoregressive spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure.
+The random-effect $S_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $S_i|S_j, j \in ne(i) ~ \mbox{Normal}(\bar{S_i}, \frac{\sigma^{2}_s}{m_j})$.
+Lastly, the BYM Spatiotemporal AR(1) (spt) adds an exchangeable random-effect for each period to allow for area-specific deviations in time trends.
+
+The space and time smoothing models used in this section partially pool the data (e.g.
+the same observation at different points in time, a tract's neighboring units) to measure the degree of variation across different quarters and at different points in space.
+These models are common in epidemiology and disease-mapping applications where researchers need to estimate a quantity of interest like a relative risk but have sparse data for some areas (which may differ from truth due to sampling variability).
+Modelling a panel's error structure with random-effects for space and time penalizes bumpiness (local deviations) in predicted values and theoretically improves out-of-sample accuracy (because large local deviations can stem from low $n$ in a given area).
+The spatial models in particular allow for missing data to be imputed based off of a local average of neighbors rather than the global or period-specific average.
 
@@
 
-- Need to build tables from output
-- Need to do write up for tables
-- Need to add graphic for predicted values
+Visit the interactive results dashboard [here](http://hesscl.com/smooth-sea/dashboard/index.html).
 
@@
 
-summary(m.int)
+#summary(m.int)
 sea_df$int_Med <- m.int$summary.linear.predictor[, "0.5quant"]
@@
 
-summary(m.ns)
+#summary(m.ns)
 sea_df$ns_Med <- m.ns$summary.linear.predictor[, "0.5quant"]
@@
 
-summary(m.nsar1) #NB: does not improve fit to use random effect time specification, using fixed effects
+#summary(m.nsar1) #NB: does not improve fit to use random effect time specification, using fixed effects
 sea_df$nsar1_Med <- m.nsar1$summary.linear.predictor[, "0.5quant"]
@@
 
-summary(m.bym)
+#summary(m.bym)
 sea_df$bym_Med <- m.bym$summary.linear.predictor[, "0.5quant"]
@@
 
-summary(m.spt)
+#summary(m.spt)
 sea_df$spt_Med <- m.spt$summary.linear.predictor[, "0.5quant"]
@@
 
+<br>
+
+```{r bym_lineplot, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+bym_df <- sea_df %>%
+  group_by(listingQtr) %>%
+  mutate(value = bym_Med,
+         qtrDev = value-mean(value, na.rm=T),
+         model = "E. Spatial AR(1)") %>%
+  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)
+
+ggplot(bym_df %>% filter(model == "E. Spatial AR(1)"), 
+         aes(x = listingQtr, y = value, group = GISJOIN, color = qtrDev, label = GISJOIN)) +
+  geom_line(lwd = .25, color = "black", alpha = .5) +
+  geom_point() +
+  scale_y_continuous(labels = scales::dollar) +
+  scale_color_viridis_c(option = "A", labels = scales::dollar) +
+  xlab("") +
+  ylab("Median 1B Rent\n") +
+  labs(color="Quarterly\nDeviation") +
+  theme_minimal() +
+  ggsave(filename = "../output/bym_lineplot.png",
+         width = 6, height = 4, dpi = 500)
+```
+
+![**Figure 6.** _Lineplot of Spatial AR(1) model predicted values._](../output/bym_lineplot.png)
+
+<br>
+
+```{r avp_map_series, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+obs_map_series <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = actualRent)) +
+  facet_grid(~ listingQtr) +
+  geom_polygon(color = "grey85", lwd = .05) +
+  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80",
+                       limits = c(800, 2800), option = "A") +
+  coord_quickmap() +
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        panel.background = element_blank(),
+        legend.position = "top") +
+  guides(fill = "none") +
+  labs(subtitle = "Observed", fill = "1B Rent")
+
+bym_map_series <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = bym_Med)) +
+  facet_grid(~ listingQtr) +
+  geom_polygon(color = "grey85", lwd = .05) +
+  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80",
+                       limits = c(800, 2800), option = "A") +
+  coord_quickmap() +
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        panel.background = element_blank(),
+        legend.position = "bottom",
+        legend.key.height = unit(.1, "inch")) +
+  guides(fill = guide_colourbar(barwidth = 15,
+       label.position = "bottom")) +
+  labs(subtitle = "Spatial AR(1) Prediction", fill = "1B Rent")
+
+obs_map_series + bym_map_series +
+  plot_layout(ncol = 1) +
+  ggsave(filename = "../output/avp_map_series.png",
+         width = 6, height = 5, dpi = 500)
+```
+
+![**Figure 7.** _Chloropleth maps of actual versus predicted values (BYM Spatial AR(1) model) for the Q1 2017 to Q2 2018 period._](../output/avp_map_series.png)
+
+<br>
+
 `r knitr::kable(fit_stats %>% select(train_test, ends_with("rmse")))`

### 2018-06-05 helpers/map_point_density.R

@@ -30,16 +30,19 @@ blank <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", geom = "blank")
 with_points <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", darken = .50, legend = "bottom") +
   stat_density_2d(aes(fill = ..level..), geom = "polygon",
                   alpha = .75, color = NA, size = .05) +
-  scale_fill_viridis_c("1B Listings") +
+  scale_fill_viridis_c("1B Listings", option = "A") +
   theme(axis.title = element_blank(),
         axis.text = element_blank(),
         axis.ticks = element_blank(),
         legend.text = element_text(angle = 45, vjust = .8),
+        legend.key.size = unit(.1, "inch"),
         panel.background = element_blank(),
-        strip.background = element_blank())
+        strip.background = element_blank()) +
+  guides(fill = guide_colourbar(barwidth = 7,
+                              label.position = "bottom"))
 
 blank + with_points +
   ggsave(filename = "../output/map_point_density.png",
-         width = 5, height = 5, units = "in", dpi = 300)
+         width = 6, height = 5.5, units = "in", dpi = 500)
 
 

### 2018-06-05 helpers/n_heatmap.R

@@ -53,21 +53,23 @@ cal<-plyr::ddply(cal,plyr::.(yearmon),transform,monthweek=1+week-min(week))
 
 #make the heatmap
 ggplot(cal, aes(x = monthweek, y = weekdayf, fill = n)) +
-  scale_fill_viridis_c(na.value = "grey80", direction = 1) +
+  scale_fill_viridis_c(na.value = "grey80", direction = 1, option = "A") +
   scale_x_continuous(breaks = c(1, 2, 3, 4)) +
   scale_size(range=c(5,20))+
   geom_tile(color = "white") +
   facet_grid(year~monthf) +
   theme_minimal() +
   theme(axis.text = element_text(size = 6),
-        legend.text = element_text(size = 8),
-        legend.title = element_text(size = 10),
+        legend.key.size = unit(.1, "inch"),
+        legend.position = "bottom",
         axis.title = element_text(size = 8),
         strip.text = element_text(size = 8),
         panel.grid = element_blank()) +
+  guides(fill = guide_colourbar(barwidth = 11,
+                                label.position = "bottom")) +
   xlab("\nWeek of Month") +
   ylab("Day of Week\n") +
-  labs(fill = "N Listings") +
+  labs(fill = "1B Listings") +
   ggsave(filename = "../output/n_heatmap.png",
-         width = 6, height = 4, dpi = 300)
+         width = 6, height = 3, dpi = 500)
 

### 2018-06-05 helpers/rent_lineplot.R

@@ -104,7 +104,7 @@ sumRent$dataSrc <- factor(sumRent$dataSrc, levels = levels(sumRent$dataSrc)[c(2,
 ggplot(sumRent, aes(x = moYr, y = Rent, group = dataSrc, color = dataSrc, shape = dataSrc)) +
   geom_line(lwd = 1.0, alpha = .5) +
   geom_point(size = 1.5) +
-  scale_color_viridis_d(begin = .1, end = .9) +
+  scale_color_viridis_d(begin = .2, end = .8, option = "A") +
   scale_y_continuous(labels = scales::dollar) +
   scale_x_date(limits = c(min(sumRent$moYr), max(sumRent$moYr))) +
   theme_minimal() +
@@ -113,4 +113,4 @@ ggplot(sumRent, aes(x = moYr, y = Rent, group = dataSrc, color = dataSrc, shape 
   labs(color = "Data Source",
        shape = "Data Source") +
   ggsave(file = "../output/rent_lineplot.png",
-         device = "png", width = 6, height = 4)
+         device = "png", width = 6, height = 4, dpi = 500)

### 2018-06-05 helpers/rent_map_across.R

@@ -138,7 +138,7 @@ ggplot(tract_f %>% filter(catBeds == "1 Bedroom"), aes(x = long, y = lat, group 
   geom_polygon(data = tract2010_bg, aes(x = long, y = lat, group = group), 
                color = "grey90", fill = "grey70", lwd = .15) +
   geom_polygon(color = "grey90", lwd = .15) +
-  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey70") +
+  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey70", option = "A") +
   coord_quickmap() +
   theme(axis.title = element_blank(),
         axis.text = element_blank(),
@@ -147,7 +147,8 @@ ggplot(tract_f %>% filter(catBeds == "1 Bedroom"), aes(x = long, y = lat, group 
         strip.background = element_blank(),
         strip.text = element_text(size = 14),
         legend.title = element_text(size = 14),
-        legend.text = element_text(size = 12)) +
+        legend.text = element_text(size = 12),
+        legend.key.width = unit(.1, "inch")) +
   labs(fill = "Asking Rent") +
   ggsave(filename = "../output/rent_map_across_1B.png",
-         width = 8, height = 4, dpi = 300)
+         width = 8, height = 4, dpi = 500)

### 2018-06-05 helpers/rent_scatter_across.R

@@ -14,7 +14,7 @@ scatter_cl <- scatter_sea %>%
   filter(seattle==1) %>%
   collect %>% #bring db query into memory
   filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
-         GISJOIN %in% sea_shp@data$GISJOIN, matchType != "Google Maps Lat/Long") %>% #only listings with valid Bed/Rent, seattle tracts
+         GISJOIN %in% tract2000@data$GISJOIN, matchType != "Google Maps Lat/Long") %>% #only listings with valid Bed/Rent, seattle tracts
   distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
            matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
   dplyr::select(listingYear, listingDate, GISJOIN, seattle, matchAddress, matchType, 
@@ -61,12 +61,15 @@ cl2017 <- scatter_cl %>%
   #filter(nHU >= 15) %>%
   select(GISJOIN, clRent, nHU)
 
-across2017 <- left_join(cl2017, ds2017)
+across2017 <- left_join(tract2000@data, cl2017)
+across2017 <- left_join(across2017, ds2017)
 pal <- brewer.pal(3, "Purples")
 
 ggplot(across2017, aes(x = dsRent, y = clRent, size = nHU)) + 
   geom_point() +
   geom_abline(slope = 1, intercept = 0) +
+  geom_smooth(show.legend = FALSE, se=FALSE, linetype = 5, 
+              color = viridis(3, begin = .25, option = "A")[1], lwd = 1.25) +
   theme_minimal() +
   coord_cartesian() +
   scale_x_continuous(limits = c(800, 2500), labels = scales::dollar) +
@@ -76,4 +79,4 @@ ggplot(across2017, aes(x = dsRent, y = clRent, size = nHU)) +
   ylab("Craigslist 2017-2018 Median 1B Rent\n") +
   labs(size = "N Craigslist\nlistings") +
   ggsave(file = "../output/rent_scatter_across_1B.png",
-         dpi = 300, width = 6, height = 4)
+         width = 6, height = 4, dpi = 300)

### 2018-06-05 manuscript/index.Rmd
almost...

@@
 title: "Estimating Neighborhood Rents with Scraped Data"
-author: "Chris Hess"
+author: "Chris Hess and Sarah Chasins"
 date: "June 2, 2018"
@@
 
-Demand for rental housing units in Seattle, WA surged in the past decade, due to relative scarcity in available housing amidst a period of historic population growth.
-Substantial appreciation for almost 2 years straight raised Seattle home prices to levels where [most households are no longer able to buy a home](https://www.seattletimes.com/business/real-estate/seattle-area-homebuyers-needed-11000-pay-raise-in-past-year-to-keep-up-with-rising-prices/?utm_source=The+Seattle+Times&utm_campaign=de86e2d581-Morning_Brief_11_28_2017&utm_medium=email&utm_term=0_5beb38b61e-de86e2d581-122931741) short of "driving till they qualify".
-More and more households must therefore rent to live in Seattle, and [almost half](http://www.jchs.harvard.edu/ARH_2017_cost_burdens_by_metro) of these renter households pay 30% of their incomes on housing costs, if not more.
-Given the universal need for housing, the individual dynamics related to rising housing costs have a role in shaping neighborhood processes of "demographic inversion" like urban gentrification and [suburbanizing poverty](https://www.brookings.edu/testimonies/the-changing-geography-of-us-poverty/) too.
+Demand for rental housing units in Seattle, WA surged in the past decade, due to relative scarcity in available housing amidst a period of [historic population growth](https://www.seattletimes.com/seattle-news/data/114000-more-people-seattle-now-this-decades-fastest-growing-big-city-in-all-of-united-states/){target="_blank"}.
+Substantial appreciation for almost 2 years straight raised Seattle home prices to levels where [most households are no longer able to buy a home](https://www.seattletimes.com/business/real-estate/seattle-area-homebuyers-needed-11000-pay-raise-in-past-year-to-keep-up-with-rising-prices/?utm_source=The+Seattle+Times&utm_campaign=de86e2d581-Morning_Brief_11_28_2017&utm_medium=email&utm_term=0_5beb38b61e-de86e2d581-122931741){target="_blank"} short of "driving till they qualify".
+More and more households must therefore rent to live in Seattle, and [almost half](http://www.jchs.harvard.edu/ARH_2017_cost_burdens_by_metro){target="_blank"} of these renter households pay 30% of their incomes on housing costs, if not more.
+Given the universal need for housing, the individual dynamics related to rising housing costs have a role in shaping neighborhood processes of "demographic inversion" like urban gentrification and [suburbanizing poverty](https://www.brookings.edu/testimonies/the-changing-geography-of-us-poverty/){target="_blank"} too.
 
 Census data have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more.
-This publication lag for these data prohibits analyzing market dynamics beyond long-term trends, particularly for smaller areas of aggregation like neighborhoods where the estimate may be an average of survey responses from a 5 year period.
+This publication lag prohibits analyzing market dynamics beyond long-term trends, particularly for smaller areas of aggregation like neighborhoods where the estimate may be an average of survey responses from a 5 year period.
 While more up-to-date options for neighborhood rent data exist through privately-held companies, such data can be somewhat of a black box in their creation and may be biased towards the costs of corporate-held properties.
@@
 Summaries of the Craigslist data show agreement with existing sources in their distribution but do suggest higher median levels for most neighborhoods, likely given the listings' recent collection relative to other data.
-The models use a quarterly panel of observed neighborhood median one-bedroom rents and produce temporally and spatially smoothed predictions about neighborhood rent levels.
-The best model specification in terms of forecasting Q2 2018 based on Q1 2017 and Q1 2018 has a combination of local spatial smoothing and autoregression (order 1), with an average forecast error of about $140 or 8% for what.
+The models use a quarterly panel of observed neighborhood median one-bedroom rents and produce temporally- and spatially-smoothed predictions about neighborhood rent levels.
+The best model specification in terms of forecasting Q2 2018 based on Q1 2017 through Q1 2018 has a combination of local spatial smoothing and autoregression (order 1), with an average forecast error of about $140 or 8% relative to the median neighborhood.
 In a more general sense, the methods and frameworks used in this study also offer a template for efficiently collecting web data, conducting reproducible research, and publishing interactive results to GitHub.
@@
 
-Existing public data are essentially limited to the [American Community Survey (ACS)](https://www.census.gov/programs-surveys/acs/) and [American Housing Survey (AHS)](https://www.census.gov/programs-surveys/ahs.html) following the cessation of long-form decennial censuses in 2000.
-These data theoretically have two sources of error relative to the ``true'' current median rent for a market and  neighborhood approximations like census tracts.
-Census sources of rental information cannot provide insight into any change in rental cost associated with newer developments becoming available (at least not until considerable time has already passed) given the temporal delay in data publication (i.e.
-ACS estimates are usually at least 2 years old, decennial many more).
+Existing public data are essentially limited to the [American Community Survey (ACS)](https://www.census.gov/programs-surveys/acs/) and [American Housing Survey (AHS)](https://www.census.gov/programs-surveys/ahs.html){target="_blank"} following the cessation of long-form decennial censuses in 2000.
+These data theoretically have two sources of error relative to the "true"" current median rent for a market and neighborhood approximations like census tracts.
+Census sources of rental information cannot provide insight into any change in rental cost associated with newer developments becoming available (at least not until considerable time has already passed) given the temporal delay in data publication.
 This feature of Census data is salient given newer housing units' generally higher rents and greater-than-average vacancy, both factors which lead to underestimating the present levels of rents for available housing units when either the ACS or decennial census are used as a current indicator.
@@
 
-[Helena](http://helena-lang.org) is a high-level programming language for end users to automate web scraping as a method of data collection.
+[Helena](http://helena-lang.org){target="_blank"} is a high-level programming language aimed at helping end users automate web scraping as a method of data collection.
 Users write programs in an extension for Google Chrome by demonstrating the first row of data they want to collect using clicks, form interactions and highlighting values.
@@
 It is important to note that web scraping does violate Craigslist's Terms of Service, and that we are using all scraped data exclusively for academic research.
-For a more thorough discussion of this issue, consult page 3 of [Boeing and Waddell (2016)](https://arxiv.org/ftp/arxiv/papers/1605/1605.05397.pdf).
+For a more thorough discussion of this issue, consult page 3 of [Boeing and Waddell (2016)](https://arxiv.org/ftp/arxiv/papers/1605/1605.05397.pdf){target="_blank"}.
 
@@
 In an attempt to compare summaries about like rental units as much as possible, visualizations focus on one-bedroom units in all but one instance.
-One-bedroom units are tied with the combined share of two and three bedroom units for the [most common unit bedroom size](https://factfinder.census.gov/bkmk/table/1.0/en/ACS/16_1YR/S2504/1600000US5363000) in Seattle according to 2016 ACS data.
+One-bedroom units are tied with the combined share of two and three bedroom units for the [most common unit bedroom size](https://factfinder.census.gov/bkmk/table/1.0/en/ACS/16_1YR/S2504/1600000US5363000){target="_blank"} in Seattle according to 2016 ACS data.
 For Figure 2, data limitations require that the comparison is all housing units since ACS only recently (i.e.
@@
 While zip codes, unofficial City neighborhoods and even arbitrary grid geographies can all be used to aggregate micro-data about rental housing, we compare census tract estimates because they roughly approximate neighborhood-size units and can be linked with a variety of other data in Seattle and more generally.
-The following links provide reference maps for Seattle Census tracts ([2000](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2000CensusSeattleTractMap.pdf) and [2010](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2010CensusTractsandZipCodeBoundaries.pdf))
+The following links provide reference maps for Seattle Census tracts ([2000](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2000CensusSeattleTractMap.pdf){target="_blank"} and [2010](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2010CensusTractsandZipCodeBoundaries.pdf){target="_blank"})
 
@@
 These tracts are found towards the North and South edges of the city, and generally are more residential neighborhoods where 20+ unit apartment complexes are not the entire stock of rental unit properties.
-In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf).
+In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf){target="_blank"}.
 
@@
 
-The following models learn from the Craigslist data accumulated up till the prior quarter (i.e.
+The following models learn from Craigslist data accumulated up till the prior quarter (i.e.
 `r as.yearqtr(Sys.Date())-1`) to accomplish a few different goals.
@@
 A secondary goal is to forecast the Seattle neighborhood rent surface for the current quarter.
-Finally, the following models allow for relatively straightforward imputation of missing values based on time period and the rent values of proximate tracts (though there are only two tracts where we have never observed a one-bedroom listing).
+Finally, the following models allow for relatively straightforward imputation of missing values based on time period and the rent values of proximate tracts.
 
-I assume that the log of the median one-bedroom rent for Seattle neighborhoods is normally distributed and estimate the posterior distribution of parameters $\theta_{it}$ using [integrated nested Laplace approximations (INLA)](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) with the default, weakly-informative priors for all random-effects.
+I assume that the log of the median one-bedroom rent for Seattle neighborhoods is normally distributed and estimate the posterior distribution of parameters $\theta_{it}$ using [integrated nested Laplace approximations (INLA)](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf){target="_blank"} with the default, weakly-informative priors for all random-effects.
+I choose to log the dependent variable because all values are positive and this transformation improves model fit.
 
@@
 ------------------------ | ------- | ----------------------------------------------
-Fixed Intercept          | int     | $\theta_{it} = \beta_0 + Qtr$
-Non-Spatial              | ns      | $\theta_{it} = \beta_0 + \epsilon_i + Qtr$
-Non-Spatial AR(1)        | nsar1   | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i$
-BYM Spatial AR(1)        | bym     | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i + S_i.$
-BYM Spatiotemporal AR(1) | spt     | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i + S_i. + \epsilon_{it}$
+Fixed Intercept          | `int`     | $\theta_{it} = \beta_0 + Qtr$
+Non-Spatial              | `ns`      | $\theta_{it} = \beta_0 + \epsilon_i + Qtr$
+Non-Spatial AR(1)        | `nsar1`  | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i$
+BYM Spatial AR(1)        | `bym `    | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i + S_i.$
+BYM Spatiotemporal AR(1) | `spt`     | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i + S_i. + \epsilon_{it}$
 
-The Fixed Intercept model (int) uses fixed effects for quarters only.
-The Non-Spatial (ns) model adds random noise (i.i.d.) tract random-effects.
-The Non-Spatial AR(1) model (nsar1) uses AR(1) process to model the temporal dynamics (formally, $\epsilon_1 \sim \mbox{Normal}(0, (\tau(1-\rho^2))^{-1}))$ where $\epsilon_t = \rho x_{t-1} + \epsilon_t$).
+The Fixed Intercept model (`int`) uses fixed effects for quarters only.
+The Non-Spatial (`ns`) model adds random noise (i.i.d.) tract random-effects.
+The Non-Spatial AR(1) model (`nsar1`) uses AR(1) process to model the temporal dynamics (formally, $\epsilon_1 \sim \mbox{Normal}(0, (\tau(1-\rho^2))^{-1}))$ where $\epsilon_t = \rho x_{t-1} + \epsilon_t$).
 This model includes a random noise term for each quarter too.
 
-The BYM Spatial with AR(1) is the [Besags-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466) intrinsic conditional autoregressive spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure.
+The BYM Spatial with AR(1) is the [Besags-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466){target="_blank"} intrinsic conditional autoregressive spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure (referred to as `bym` using shorthand).
 The random-effect $S_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $S_i|S_j, j \in ne(i) ~ \mbox{Normal}(\bar{S_i}, \frac{\sigma^{2}_s}{m_j})$.
-Lastly, the BYM Spatiotemporal AR(1) (spt) adds an exchangeable random-effect for each period to allow for area-specific deviations in time trends.
+Lastly, the BYM Spatiotemporal AR(1) (`spt`) adds an exchangeable random-effect for each period to allow for area-specific deviations in time trends.
 
@@
 
-Visit the interactive results dashboard [here](http://hesscl.com/smooth-sea/dashboard/index.html).
+Visit the interactive results dashboard [here](http://hesscl.com/smooth-sea/dashboard/index.html){target="_blank"}.
 
@@
 
+The following tables show in-sample and out-of-sample (i.e.
+2018 Q2) accuracy for the 5 model specifications in terms of root mean squared-error (RMSE) and mean absolute error (MAE).
+Note how though the Non-Spatial models (i.e.
+`ns` and `nsar1`) achieve slightly better in-sample fits, both Spatial models are more accurate at predicting the current month's surface of rent values.
+It is important to note how the goal in the present case is to understand the data generating process, that is, the systematic variation in the data less the random component.
+So while we want to minimize model error as a way to judge if our models improve on each other, we also want to smooth over sampling variability in the Craigslist data that induces local deviations compared to the true median rent.
+The [information criteria for the models](http://hesscl.com/smooth-sea/dashboard/index.html#model-fit){target="_blank"} generally show that the spatial specifications are more likely models of the neighborhood rent data generating process.
+
 <br>
+
+`r knitr::kable(fit_stats %>% select(train_test, ends_with("rmse")))`
+
+`r knitr::kable(fit_stats %>% select(train_test, ends_with("mae")))`
+
+<br>
+
+Figure 6 depicts the predicted temporal structure based on the top fitting model, the Spatial AR(1) specification.
+The quarterly fluctuation is about $50-$100 depending on the tract in question.
+Modeling neighborhood median rents on log scale allows for non-constant variance across the range of y, such that the periodic deviations are largest at the top of the distribution and more subdued among low-rent neighborhoods.
 
@@
   theme_minimal() +
+  theme(legend.key.width = unit(.1, "inch")) +
   ggsave(filename = "../output/bym_lineplot.png",
-         width = 6, height = 4, dpi = 500)
+         width = 6, height = 3, dpi = 500)
 ```
 
-![**Figure 6.** _Lineplot of Spatial AR(1) model predicted values._](../output/bym_lineplot.png)
+![**Figure 6.** _Lineplot of Spatial AR(1) model predicted values.
+Colors denote a tract's deviation from the average neighborhood 1B rent level on a quarter-by-quarter basis._](../output/bym_lineplot.png)
 
 <br>
+
+Figure 7 provides a spatial time-series for the actual and predicted values using the Spatial AR(1) `bym` model.
+Since there is no space time interaction, the spatial distribution is constant across the quarter periods.
+The quarterly fluctuation is visible in both the observed and predicted data, though the spatial distribution of rent levels in the model predictions smooths over particularly high and low local deviations.
+The model imputes reasonable values for missing data over the entire panel, giving an estimate of what we would expect a one-bedroom to cost based on the period, neighbors and any earlier observations of the tract (if they exist).
+Overall, one clear takeaway is that the spatial variation in rent cost swamps that of the short-term dynamics observed to date.
+In this sense, estimating the spatial distribution completely and accurately is a crucial goal of any indicator of the market---whether a simple median or a complex model like the current one.
 
@@
 
-`r knitr::kable(fit_stats %>% select(train_test, ends_with("rmse")))`
+## Discussion
 
-`r knitr::kable(fit_stats %>% select(train_test, ends_with("mae")))`
-
-<br>
-
-## Discussion
+This study highlights the importance of considering scraped web data for understanding the dynamics of neighborhood rental housing costs.
+Throughout this work we demonstrate how Craigslist rental housing listings scraped may be used to research the estimated rent for the Seattle market and its neighborhoods.
+Models
 
@@
 ## References
+
+[Boeing, G.
+and Waddell, P., 2017.
+New insights into rental housing markets across the united states: web scraping and analyzing craigslist rental listings.
+Journal of Planning Education and Research, 37(4), pp.457-476.](http://journals.sagepub.com/doi/abs/10.1177/0739456X16664789){target="_blank"}
+
+[Besag, J., York, J.
+and Mollie, A., 1991.
+Bayesian image restoration with two applications in spatial statistics (with discussion) Annals of the Institute of Statistical Mathematics.
+43, pp.1-59.](https://link.springer.com/article/10.1007/BF00116466){target="_blank"}
+
+[Blangiardo, M., Cameletti, M., Baio, G.
+and Rue, H., 2013.
+Spatial and spatio-temporal models with R-INLA.
+Spatial and spatio-temporal epidemiology, 4, pp.33-49.](https://www.sciencedirect.com/science/article/pii/S1877584512000846){target="_blank"}

### 2018-06-05 helpers/n_heatmap.R

@@ -71,5 +71,5 @@ ggplot(cal, aes(x = monthweek, y = weekdayf, fill = n)) +
   ylab("Day of Week\n") +
   labs(fill = "1B Listings") +
   ggsave(filename = "../output/n_heatmap.png",
-         width = 6, height = 3, dpi = 500)
+         width = 7, height = 3, dpi = 500)
 

### 2018-06-05 manuscript/index.Rmd
submission version

@@
 The key steps of any data process can be distilled into three tasks: a.) _extract_ a table of interest from a specified origin source, b.) _transform_ it to support the desired analyses and c.) _load_ the transformed data to a storage destination for future use.
+In the present case, a program extracts scraped listings observed since the data updated by sending a download request to the scraper's accumulated data in the form of a URL with a suffix indicating the number of hours that passed since the last processing.
+The processing routine then transforms the data by cleaning up some of the scraped fields and appending spatial information through a multi-step geocoding process.
 
-In the present case, a program extracts scraped listings observed since the data updated by sending a download request to the scraper's accumulated data in the form of a URL with a suffix indicating the number of hours that passed since the last processing.
-The processing routine then transforms the data by cleaning up some of the scraped fields and appending spatial information through a multi-step geocoding proces.
 If address information was successfully scraped from the listing metadata, the script makes a call to Google Maps Geocoding API using this listing feature and a constraint to search within Washington state.
@@
 Throughout this work we demonstrate how Craigslist rental housing listings scraped may be used to research the estimated rent for the Seattle market and its neighborhoods.
-Models
+Using Bayesian spatial smoothing models further exploit the scraped data's spatial and temporal richness, giving researchers an option for studying rental housing market dynamics on a more granular scale where time is not only considered in terms of trend, but also random shock and seasonality.
+
+Finally, this paper has endeavored (perhaps a little too much...) to provide an example for using scraped rental housing listings for research, but also with to model for how Rmarkdown, Git and GitHub are transformative technologies for our disciplines.
+There are no free lunches with reproducibility: these open-source technologies demand that we learn how to incorporate them into our work.
+To the extent that this requires thinking about analysis beyond a userbase of 1, there can be some growing pains.
+However, there are some positive spillovers beyond improving the _science_ of social science---including new platforms and media for sharing research findings and ideas.
 

### 2018-06-05 manuscript/index.Rmd
caught a couple date typos

@@
 author: "Chris Hess and Sarah Chasins"
-date: "June 2, 2018"
+date: "June 5, 2018"
 output: 
@@
 All models' posterior distribution are estimated with integrated nested Laplace approximations (INLA).
-Training data for models cover quarterly tract estimates for 2017 Q2 through 2018 Q1, while the test data are 2018 Q2 median values as observed through June 3rd.
+Training data for models cover quarterly tract estimates for 2017 Q2 through 2018 Q1, while the test data are 2018 Q2 median values as observed through June 5rd.
 The top model assumes that spatial and temporal dynamics are separable, i.e.

### 2018-06-05 manuscript/index.Rmd
final version for panel data

@@
 All models' posterior distribution are estimated with integrated nested Laplace approximations (INLA).
-Training data for models cover quarterly tract estimates for 2017 Q2 through 2018 Q1, while the test data are 2018 Q2 median values as observed through June 5rd.
+Training data for models cover quarterly tract estimates for 2017 Q2 through 2018 Q1, while the test data are 2018 Q2 median values as observed through June 4th.
 The top model assumes that spatial and temporal dynamics are separable, i.e.

### 2018-06-05 manuscript/index.Rmd
catching a few typos in discussion

@@
   ggsave(filename = "../output/bym_lineplot.png",
-         width = 6, height = 3, dpi = 500)
+         width = 6, height = 3, dpi =300)
 ```
@@
   ggsave(filename = "../output/avp_map_series.png",
-         width = 6, height = 5, dpi = 500)
+         width = 6, height = 5, dpi =300)
 ```
@@
 
-Finally, this paper has endeavored (perhaps a little too much...) to provide an example for using scraped rental housing listings for research, but also with to model for how Rmarkdown, Git and GitHub are transformative technologies for our disciplines.
+Finally, this paper has endeavored (perhaps a little too much...) to provide an example of using scraped rental housing listings for research, but also to model how Rmarkdown, Git and GitHub are transformative  for our disciplines.
 There are no free lunches with reproducibility: these open-source technologies demand that we learn how to incorporate them into our work.
 To the extent that this requires thinking about analysis beyond a userbase of 1, there can be some growing pains.
-However, there are some positive spillovers beyond improving the _science_ of social science---including new platforms and media for sharing research findings and ideas.
+However, there are some positive spillovers beyond improving the _science_ of social science---including new platforms and media for sharing research findings.
 

### 2018-06-05 helpers/map_point_density.R

@@ -43,6 +43,6 @@ with_points <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", darken = 
 
 blank + with_points +
   ggsave(filename = "../output/map_point_density.png",
-         width = 6, height = 5.5, units = "in", dpi = 500)
+         width = 6, height = 5.5, units = "in", dpi = 300)
 
 

### 2018-06-05 helpers/n_heatmap.R

@@ -71,5 +71,5 @@ ggplot(cal, aes(x = monthweek, y = weekdayf, fill = n)) +
   ylab("Day of Week\n") +
   labs(fill = "1B Listings") +
   ggsave(filename = "../output/n_heatmap.png",
-         width = 7, height = 3, dpi = 500)
+         width = 7, height = 3, dpi = 300)
 

### 2018-06-05 helpers/rent_lineplot.R

@@ -113,4 +113,4 @@ ggplot(sumRent, aes(x = moYr, y = Rent, group = dataSrc, color = dataSrc, shape 
   labs(color = "Data Source",
        shape = "Data Source") +
   ggsave(file = "../output/rent_lineplot.png",
-         device = "png", width = 6, height = 4, dpi = 500)
+         device = "png", width = 6, height = 4, dpi = 300)

### 2018-06-05 helpers/rent_map_across.R

@@ -151,4 +151,4 @@ ggplot(tract_f %>% filter(catBeds == "1 Bedroom"), aes(x = long, y = lat, group 
         legend.key.width = unit(.1, "inch")) +
   labs(fill = "Asking Rent") +
   ggsave(filename = "../output/rent_map_across_1B.png",
-         width = 8, height = 4, dpi = 500)
+         width = 8, height = 4, dpi = 300)

### 2018-06-15 manuscript/index.Rmd
catching typos, working on the helper scripts some

@@
 
-Census data have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more.
+Census data have common currency among scholars and practitioners interested in metropolitan housing markets, but nevertheless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more.
 This publication lag prohibits analyzing market dynamics beyond long-term trends, particularly for smaller areas of aggregation like neighborhoods where the estimate may be an average of survey responses from a 5 year period.
@@
 Only the Craigslist and Dupre+Scott can offer insight into more recent median levels for the Seattle rental housing market.
-Both data suggest continued increases in the marketwide median rent for the first half of 2017, though the Craigslist slope is about twice in its size.
-Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for 2017 Q2 upward in favor of regularly posted commerical listings.
+Both data suggest continued increases in the market-wide median rent for the first half of 2017, though the Craigslist slope is about twice in its size.
+Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for 2017 Q2 upward in favor of regularly posted commercial listings.
 Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability.
@@
 
-Though the spatial distribution of the ACS estimates resembles that of the other two data sources, the values are nonethless systematically lower (in most cases by hundreds of dollars).
+Though the spatial distribution of the ACS estimates resembles that of the other two data sources, the values are nonetheless systematically lower (in most cases by hundreds of dollars).
 The ACS measures therefore give insight into the relative ranking of neighborhoods in terms of rental costs, but are limited for studying neighborhood dynamics outside of long-run trend.
@@
 
-Figure 5 shows a scatterplot/bubbpleplot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 75 tracts where comparable estimates exist.
+Figure 5 shows a scatter plot/bubble plot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 75 tracts where comparable estimates exist.
 The purple dashed line indicates the loess smoother fit between the measures.
@@
 
-An alternative explanation fpr the difference between measures comes in considering the units these properties are sampling from in a given neighborhood.
+An alternative explanation for the difference between measures comes in considering the units these properties are sampling from in a given neighborhood.
 The greater difference between the measures among lower-rent neighborhoods (with Craigslist estimating these as more expensive than Dupre+Scott) provides some evidence for this theory.
@@
 
-The BYM Spatial with AR(1) is the [Besags-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466){target="_blank"} intrinsic conditional autoregressive spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure (referred to as `bym` using shorthand).
+The BYM Spatial with AR(1) is the [Besag-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466){target="_blank"} intrinsic conditional autoregressive spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure (referred to as `bym` using shorthand).
 The random-effect $S_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $S_i|S_j, j \in ne(i) ~ \mbox{Normal}(\bar{S_i}, \frac{\sigma^{2}_s}{m_j})$.
@@
 There are no free lunches with reproducibility: these open-source technologies demand that we learn how to incorporate them into our work.
-To the extent that this requires thinking about analysis beyond a userbase of 1, there can be some growing pains.
+To the extent that this requires thinking about analysis beyond a user base of 1, there can be some growing pains.
 However, there are some positive spillovers beyond improving the _science_ of social science---including new platforms and media for sharing research findings.

### 2018-06-15 helpers/load_data.R

@@ -8,7 +8,7 @@ load_data <- function(){
     #compute tract aggregates for CL listing count
     tract <- cl %>%
       collect %>% #bring db query into memory
-      filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
+      filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), !is.na(matchAddress), !is.na(matchAddress2),
              GISJOIN %in% sea_shp@data$GISJOIN, #only listings with valid Bed/Rent, seattle tracts
              matchType != "Google Maps Lat/Long") %>% #need to have address, not approximate
       distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
@@ -17,7 +17,7 @@ load_data <- function(){
                     cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
       mutate(listingDate = as.Date(listingDate),
              listingQtr = as.yearqtr(listingDate)) %>%
-      filter(listingQtr >= "2017 Q1") %>%
+      filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3") %>%
       group_by(GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
       summarize(nListings = n(),
                 n1B = sum(cleanBeds == 1),
@@ -32,8 +32,8 @@ load_data <- function(){
     write_csv(tract, "../input/tractCl.csv")
     return(tract)
   } else{
-    #if github config, read in extract from cl db
-    tract <- read_csv("../input/tractCl.csv")
+    #if github config
+    print("Need input data")
     return(tract)
   }
 }

### 2018-06-15 helpers/map_point_density.R

@@ -14,15 +14,17 @@ map_raw <- tbl(DB, "raw")
 map_cl <- map_clean %>% 
   filter(seattle==1) %>%
   collect %>% #bring db query into memory
-  filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
-         GISJOIN %in% sea_shp@data$GISJOIN, matchType != "Google Maps Lat/Long") %>% #only listings with valid Bed/Rent, seattle tracts
+  filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), !is.na(matchAddress), !is.na(matchAddress2),
+         GISJOIN %in% sea_shp@data$GISJOIN, #only listings with valid Bed/Rent, seattle tracts
+         matchType != "Google Maps Lat/Long") %>% #need to have address, not approximate
   distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
            matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
   dplyr::select(listingDate, GISJOIN, seattle, matchAddress, matchType, 
                 cleanBeds, cleanRent, cleanSqft, lat, lng) %>% #SELECT these columns
   mutate(listingDate = as.Date(listingDate),
          listingQtr = as.yearqtr(listingDate)) %>%
-  filter(cleanBeds %in% c(1), listingQtr >= "2017 Q2")
+  filter(cleanBeds %in% c(1), listingQtr >= "2017 Q1", listingQtr < "2018 Q3")
+
 dbDisconnect(DB)
 
 blank <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", geom = "blank") 

### 2018-06-15 helpers/n_heatmap.R

@@ -5,15 +5,16 @@ cl <- tbl(DB, "clean") #clean listing table
 
 #compute tract aggregates for CL listing count
 sea_cl <- cl %>%
-  filter(!is.na(GISJOIN), seattle == 1) %>% #only listings with valid Bed/Rent
-  dplyr::select(listingMoYr, listingDate, GISJOIN, seattle, matchAddress, matchAddress2, matchType, cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
+  dplyr::select(listingMoYr, listingDate, GISJOIN, seattle, matchAddress, matchAddress2, 
+                matchType, cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
   collect %>% #bring db query into memory
-  filter(GISJOIN %in% sea_shp@data$GISJOIN) %>% #filter to KC only (db has metro area)
+  filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), !is.na(matchAddress), !is.na(matchAddress2),
+         GISJOIN %in% sea_shp@data$GISJOIN, #only listings with valid Bed/Rent, seattle tracts
+         matchType != "Google Maps Lat/Long") %>% #need to have address, not approximate
   mutate(listingDate = as.Date(listingDate),
          listingQtr = as.yearqtr(listingDate)) %>%
-  filter(listingQtr >= "2017 Q1", !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft),
+  filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3",
          cleanBeds == 1) %>% 
-  filter(matchType != "Google Maps Lat/Long") %>% #no Google lat/long only geocodes
   distinct(matchAddress, matchAddress2, cleanBeds, cleanRent, cleanSqft, .keep_all = T)
 dbDisconnect(DB)
 

### 2018-06-15 helpers/rent_lineplot.R

@@ -20,12 +20,13 @@ sumCL <- cl %>%
   mutate(catSqft = fct_reorder(catSqft, cleanSqft)) %>%
   arrange(listingDate) %>%
   select(-ends_with("10")) %>%
-  filter(!is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), !is.na(matchAddress), !is.na(matchAddress2),
-         matchType != "Google Maps Lat/Long") %>%
+  filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), !is.na(matchAddress), !is.na(matchAddress2),
+         GISJOIN %in% sea_shp@data$GISJOIN, #only listings with valid Bed/Rent, seattle tracts
+         matchType != "Google Maps Lat/Long") %>% #need to have address, not approximate
   distinct(matchAddress, matchAddress2, catBeds, cleanSqft, cleanRent, .keep_all = T) %>%
   mutate(listingQtr = as.yearqtr(as.Date(listingDate))) %>%
+  filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3") %>%
   group_by(listingQtr) %>%
-  filter(listingQtr >= "2017 Q1") %>%
   summarize(Rent = median(cleanRent, na.rm=T),
             nHU = n()) %>%
   rename(moYr = listingQtr) %>%

### 2018-06-15 helpers/rent_map_across.R

@@ -20,10 +20,10 @@ tract2010_bg <- fortify(tract2010)
 cl2017 <- rent_map_sea %>% 
   filter(seattle==1) %>% 
   collect %>% 
-  filter(listingYear >= 2017, 
-         matchType != "Google Maps Lat/Long",
-         !is.na(matchAddress), !is.na(matchAddress2), !is.na(catBeds), !is.na(cleanSqft), !is.na(cleanRent),
-         catBeds %in% c("0", "1", "2"), !is.na(lat) & !is.na(lng)) %>%
+  mutate(listingQtr = as.yearqtr(listingDate)) %>%
+  filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3",
+         !is.na(GISJOIN), !is.na(matchAddress), !is.na(matchAddress2), !is.na(catBeds), !is.na(cleanSqft), !is.na(cleanRent),
+         catBeds %in% c("0", "1", "2")) %>%
   distinct(matchAddress, matchAddress2, catBeds, cleanSqft, cleanRent, .keep_all = T) %>%
   group_by(GISJOIN, catBeds) %>%
   summarize(rent = median(cleanRent, na.rm = T),

### 2018-06-15 helpers/rent_scatter_across.R

@@ -13,7 +13,7 @@ scatter_ds <- read_csv("R:/Project/seattle_rental_market/data/d+s/geocoded_D+S.c
 scatter_cl <- scatter_sea %>% 
   filter(seattle==1) %>%
   collect %>% #bring db query into memory
-  filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
+  filter(!is.na(GISJOIN), !is.na(matchAddress), !is.na(matchAddress2), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
          GISJOIN %in% tract2000@data$GISJOIN, matchType != "Google Maps Lat/Long") %>% #only listings with valid Bed/Rent, seattle tracts
   distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
            matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
@@ -21,7 +21,7 @@ scatter_cl <- scatter_sea %>%
                 cleanBeds, cleanRent, cleanSqft, lat, lng) %>% #SELECT these columns
   mutate(listingDate = as.Date(listingDate),
          listingQtr = as.yearqtr(listingDate)) %>%
-  filter(cleanBeds %in% c(1), listingQtr >= "2017 Q1")
+  filter(cleanBeds %in% c(1), listingQtr >= "2017 Q1", listingQtr < "2018 Q3")
 
 coordinates(scatter_cl) <- cbind(scatter_cl$lng, scatter_cl$lat)
 

### 2018-06-15 manuscript/index.Rmd
updating some language

@@
 This study demonstrates why scraped rental housing listings are a valuable source of data in subject areas such as housing studies and neighborhood demography.
-I draw on a set of descriptive analyses and models of quarterly rent levels among census tracts in Seattle, WA for this research objective.
+We draw on a set of descriptive analyses and models of quarterly rent levels among census tracts in Seattle, WA for this research objective.
 Summaries of the Craigslist data show agreement with existing sources in their distribution but do suggest higher median levels for most neighborhoods, likely given the listings' recent collection relative to other data.
@@
 
-I start by briefly reviewing some background relevant to studying neighborhood rental housing.
+We start by briefly reviewing some background relevant to studying neighborhood rental housing.
 I then outline a process for automated web scraping of a rental listing source like Craigslist using a program called Helena, and compare aggregates from the collected rental listings with other data to assess rent agreement in measures.
-After describing the data and process, I present a set of models predicting neighborhood rent levels to understand different how different temporal and spatial structures fit the data.
-I conclude by discussing the promise of web data for subfields like housing studies and the value of R Markdown and GitHub for social science research.
+After describing the data and process, We present a set of models predicting neighborhood rent levels to understand different how different temporal and spatial structures fit the data.
+We conclude by discussing the promise of web data for subfields like housing studies and the value of R Markdown and GitHub for social science research.
 
@@
 
-I assume that the log of the median one-bedroom rent for Seattle neighborhoods is normally distributed and estimate the posterior distribution of parameters $\theta_{it}$ using [integrated nested Laplace approximations (INLA)](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf){target="_blank"} with the default, weakly-informative priors for all random-effects.
-I choose to log the dependent variable because all values are positive and this transformation improves model fit.
+We assume that the log of the median one-bedroom rent for Seattle neighborhoods is normally distributed and compute the posterior distributions of each model's hyperparameters and parameters $\theta_{it}$ using [integrated nested Laplace approximations (INLA)](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf){target="_blank"} with the default, weakly-informative priors for all random-effects.
+We choose to log the dependent variable because all values are positive and this transformation improves model fit.
 

### 2018-06-15 helpers/load_data.R

@@ -29,7 +29,6 @@ load_data <- function(){
       ungroup %>%
       arrange(GISJOIN, listingQtr)
     dbDisconnect(DB)
-    write_csv(tract, "../input/tractCl.csv")
     return(tract)
   } else{
     #if github config

### 2018-06-15 manuscript/index.Rmd
fixing notation for models

@@
 Non-Spatial              | `ns`      | $\theta_{it} = \beta_0 + \epsilon_i + Qtr$
-Non-Spatial AR(1)        | `nsar1`  | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i$
-BYM Spatial AR(1)        | `bym `    | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i + S_i.$
-BYM Spatiotemporal AR(1) | `spt`     | $\theta_{it} = \beta_0 + \epsilon_t + \epsilon_{t-1} + \epsilon_i + S_i. + \epsilon_{it}$
+Non-Spatial AR(1)        | `nsar1`  | $\theta_{it} = \beta_0 + x_t + \epsilon_t + \epsilon_i$
+BYM Spatial AR(1)        | `bym `    | $\theta_{it} = \beta_0 + x_t + \epsilon_t  + \epsilon_i + S_i.$
+BYM Spatiotemporal AR(1) | `spt`     | $\theta_{it} = \beta_0 + x_t + \epsilon_t  + \epsilon_i + S_i + \epsilon_{it}$
 
@@
 The Non-Spatial (`ns`) model adds random noise (i.i.d.) tract random-effects.
-The Non-Spatial AR(1) model (`nsar1`) uses AR(1) process to model the temporal dynamics (formally, $\epsilon_1 \sim \mbox{Normal}(0, (\tau(1-\rho^2))^{-1}))$ where $\epsilon_t = \rho x_{t-1} + \epsilon_t$).
+The Non-Spatial AR(1) model (`nsar1`) uses AR(1) process to model the temporal dynamics (formally, $x_1 \sim \mbox{Normal}(0, (\tau(1-\rho^2))^{-1})$ where $x_t = \rho x_{t-1} + \epsilon_t$).
 This model includes a random noise term for each quarter too.

### 2018-06-16 manuscript/index.Rmd
fixing notation to match Blangiardo et al and Wakefield

@@
 ------------------------ | ------- | ----------------------------------------------
-Fixed Intercept          | `int`     | $\theta_{it} = \beta_0 + Qtr$
-Non-Spatial              | `ns`      | $\theta_{it} = \beta_0 + \epsilon_i + Qtr$
-Non-Spatial AR(1)        | `nsar1`  | $\theta_{it} = \beta_0 + x_t + \epsilon_t + \epsilon_i$
-BYM Spatial AR(1)        | `bym `    | $\theta_{it} = \beta_0 + x_t + \epsilon_t  + \epsilon_i + S_i.$
-BYM Spatiotemporal AR(1) | `spt`     | $\theta_{it} = \beta_0 + x_t + \epsilon_t  + \epsilon_i + S_i + \epsilon_{it}$
+Fixed Intercept          | `int`   | $\theta_{it} = \beta_0 + \gamma_t$
+Non-Spatial              | `ns`    | $\theta_{it} = \beta_0 + \nu_i + \gamma_t$
+Non-Spatial AR(1)        | `nsar1` | $\theta_{it} = \beta_0 + x_t + \phi_t + \nu_i$
+BYM Spatial AR(1)        | `bym `  | $\theta_{it} = \beta_0 + x_t + \phi_t + v_i + \nu_i $
+BYM Spatiotemporal AR(1) | `spt`   | $\theta_{it} = \beta_0 + x_t + \phi_t + v_i + \nu_i + \delta_{it}$
 
-The Fixed Intercept model (`int`) uses fixed effects for quarters only.
-The Non-Spatial (`ns`) model adds random noise (i.i.d.) tract random-effects.
-The Non-Spatial AR(1) model (`nsar1`) uses AR(1) process to model the temporal dynamics (formally, $x_1 \sim \mbox{Normal}(0, (\tau(1-\rho^2))^{-1})$ where $x_t = \rho x_{t-1} + \epsilon_t$).
-This model includes a random noise term for each quarter too.
+The Fixed Intercept model (`int`) has a global intercept ($\beta_0$) and fixed effects ($\gamma_t$) for each quarter of each year.
+The Non-Spatial (`ns`) model adds random noise (i.i.d.) tract random-effects ($\nu_i \sim Normal(0, \tau_\nu)$) to the Fixed Intercept specification.
+The Non-Spatial AR(1) model (`nsar1`) then incorporates term an AR(1) random-effect ($x_1 \sim \mbox{Normal}(0, (\tau(1-\rho^2))^{-1})$ where $x_t = \rho x_{t-1} + \epsilon_t$) to dynamically model the temporal variation.
+This model includes a random noise term ($\phi_t \sim Normal(0, \tau_\phi)$) for each quarter too.
 
 The BYM Spatial with AR(1) is the [Besag-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466){target="_blank"} intrinsic conditional autoregressive spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure (referred to as `bym` using shorthand).
-The random-effect $S_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $S_i|S_j, j \in ne(i) ~ \mbox{Normal}(\bar{S_i}, \frac{\sigma^{2}_s}{m_j})$.
-Lastly, the BYM Spatiotemporal AR(1) (`spt`) adds an exchangeable random-effect for each period to allow for area-specific deviations in time trends.
+The random-effect $S_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $v_i|v_j, j \in ne(i) ~ \mbox{Normal}(\bar{v_i}, \frac{\sigma^{2}_v}{m_j})$.
+Lastly, the BYM Spatiotemporal AR(1) (`spt`) adds an exchangeable random-effect ($\delta_i \sim Normal(0, \tau_\delta)$) for each period to allow for area-specific deviations in time trends.
 

### 2018-06-16 manuscript/index.Rmd
one more tweak to notation

@@
 Non-Spatial AR(1)        | `nsar1` | $\theta_{it} = \beta_0 + x_t + \phi_t + \nu_i$
-BYM Spatial AR(1)        | `bym `  | $\theta_{it} = \beta_0 + x_t + \phi_t + v_i + \nu_i $
+BYM Spatial AR(1)        | `bym `  | $\theta_{it} = \beta_0 + x_t + \phi_t + v_i + \nu_i$
 BYM Spatiotemporal AR(1) | `spt`   | $\theta_{it} = \beta_0 + x_t + \phi_t + v_i + \nu_i + \delta_{it}$

### 2018-06-27 manuscript/manuscript.Rmd
adding manuscript.rmd for working on PDF version

--- 2018-06-16 manuscript/index.Rmd
@@
 
-The BYM Spatial with AR(1) is the [Besag-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466){target="_blank"} intrinsic conditional autoregressive spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure (referred to as `bym` using shorthand).
-The random-effect $S_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $v_i|v_j, j \in ne(i) ~ \mbox{Normal}(\bar{v_i}, \frac{\sigma^{2}_v}{m_j})$.
+The BYM Spatial with AR(1) is the [Besag-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466){target="_blank"} intrinsic conditional autoregressive (ICAR) spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure (referred to as `bym` using shorthand).
+The random-effect $v_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $v_i|v_j, j \in ne(i) ~ \mbox{Normal}(\bar{v_i}, \frac{\sigma^{2}_v}{m_j})$.
 Lastly, the BYM Spatiotemporal AR(1) (`spt`) adds an exchangeable random-effect ($\delta_i \sim Normal(0, \tau_\delta)$) for each period to allow for area-specific deviations in time trends.

### 2018-06-28 manuscript/manuscript.Rmd
updating to include PDF output version

@@
 date: "June 5, 2018"
-output: 
-  html_document:
-    toc: true
-    toc_float: true
-    includes:
-      in_header: "header.html"
-      after_body: "../footer.html"
+output:  pdf_document
 abstract: "Online classified ads are a key source of information about local rental housing markets.
@@
 The second component of the paper investivates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models.
-All models' posterior distribution are estimated with integrated nested Laplace approximations (INLA).
+Posterior distributions of all model hyperparameters and parameters are estimated with integrated nested Laplace approximations (INLA).
 Training data for models cover quarterly tract estimates for 2017 Q2 through 2018 Q1, while the test data are 2018 Q2 median values as observed through June 4th.
@@
 ---
-
-<style>
-.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
-    background-color: #4b2e83;
-}
-</style>
 
@@
 
-Demand for rental housing units in Seattle, WA surged in the past decade, due to relative scarcity in available housing amidst a period of [historic population growth](https://www.seattletimes.com/seattle-news/data/114000-more-people-seattle-now-this-decades-fastest-growing-big-city-in-all-of-united-states/){target="_blank"}.
-Substantial appreciation for almost 2 years straight raised Seattle home prices to levels where [most households are no longer able to buy a home](https://www.seattletimes.com/business/real-estate/seattle-area-homebuyers-needed-11000-pay-raise-in-past-year-to-keep-up-with-rising-prices/?utm_source=The+Seattle+Times&utm_campaign=de86e2d581-Morning_Brief_11_28_2017&utm_medium=email&utm_term=0_5beb38b61e-de86e2d581-122931741){target="_blank"} short of "driving till they qualify".
-More and more households must therefore rent to live in Seattle, and [almost half](http://www.jchs.harvard.edu/ARH_2017_cost_burdens_by_metro){target="_blank"} of these renter households pay 30% of their incomes on housing costs, if not more.
-Given the universal need for housing, the individual dynamics related to rising housing costs have a role in shaping neighborhood processes of "demographic inversion" like urban gentrification and [suburbanizing poverty](https://www.brookings.edu/testimonies/the-changing-geography-of-us-poverty/){target="_blank"} too.
+Demand for rental housing units in Seattle, WA surged in the past decade, due to relative scarcity in available housing amidst a period of [historic population growth](https://www.seattletimes.com/seattle-news/data/114000-more-people-seattle-now-this-decades-fastest-growing-big-city-in-all-of-united-states/).
+Substantial appreciation for almost 2 years straight raised Seattle home prices to levels where [most households are no longer able to buy a home](https://www.seattletimes.com/business/real-estate/seattle-area-homebuyers-needed-11000-pay-raise-in-past-year-to-keep-up-with-rising-prices/?utm_source=The+Seattle+Times&utm_campaign=de86e2d581-Morning_Brief_11_28_2017&utm_medium=email&utm_term=0_5beb38b61e-de86e2d581-122931741) short of "driving till they qualify".
+More and more households must therefore rent to live in Seattle, and [almost half](http://www.jchs.harvard.edu/ARH_2017_cost_burdens_by_metro) of these renter households pay 30% of their incomes on housing costs, if not more.
+Given the universal need for housing, the individual dynamics related to rising housing costs have a role in shaping neighborhood processes of "demographic inversion" like urban gentrification and [suburbanizing poverty](https://www.brookings.edu/testimonies/the-changing-geography-of-us-poverty/) too.
 
@@
 
-Existing public data are essentially limited to the [American Community Survey (ACS)](https://www.census.gov/programs-surveys/acs/) and [American Housing Survey (AHS)](https://www.census.gov/programs-surveys/ahs.html){target="_blank"} following the cessation of long-form decennial censuses in 2000.
+Existing public data are essentially limited to the [American Community Survey (ACS)](https://www.census.gov/programs-surveys/acs/) and [American Housing Survey (AHS)](https://www.census.gov/programs-surveys/ahs.html) following the cessation of long-form decennial censuses in 2000.
 These data theoretically have two sources of error relative to the "true"" current median rent for a market and neighborhood approximations like census tracts.
@@
 
-[Helena](http://helena-lang.org){target="_blank"} is a high-level programming language aimed at helping end users automate web scraping as a method of data collection.
+[Helena](http://helena-lang.org) is a high-level programming language aimed at helping end users automate web scraping as a method of data collection.
 Users write programs in an extension for Google Chrome by demonstrating the first row of data they want to collect using clicks, form interactions and highlighting values.
@@
 It is important to note that web scraping does violate Craigslist's Terms of Service, and that we are using all scraped data exclusively for academic research.
-For a more thorough discussion of this issue, consult page 3 of [Boeing and Waddell (2016)](https://arxiv.org/ftp/arxiv/papers/1605/1605.05397.pdf){target="_blank"}.
+For a more thorough discussion of this issue, consult page 3 of [Boeing and Waddell (2016)](https://arxiv.org/ftp/arxiv/papers/1605/1605.05397.pdf).
 
@@
 In an attempt to compare summaries about like rental units as much as possible, visualizations focus on one-bedroom units in all but one instance.
-One-bedroom units are tied with the combined share of two and three bedroom units for the [most common unit bedroom size](https://factfinder.census.gov/bkmk/table/1.0/en/ACS/16_1YR/S2504/1600000US5363000){target="_blank"} in Seattle according to 2016 ACS data.
+One-bedroom units are tied with the combined share of two and three bedroom units for the [most common unit bedroom size](https://factfinder.census.gov/bkmk/table/1.0/en/ACS/16_1YR/S2504/1600000US5363000) in Seattle according to 2016 ACS data.
 For Figure 2, data limitations require that the comparison is all housing units since ACS only recently (i.e.
@@
 
-![**Figure 1.** _Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right)._](../output/map_point_density.png)
+![Map of Seattle (left), Seattle Craigslist one-bedroom listing point and 2D-density map (right).](../output/map_point_density.png)
 
@@
 
-![**Figure 2.** _Heatmap of N Seattle Craigslist one-bedroom listings scraped per day time-series._](../output/n_heatmap.png)
+![Heatmap of N Seattle Craigslist one-bedroom listings scraped per day time-series.](../output/n_heatmap.png)
 
@@
 
-![**Figure 3.** _Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018.
+![Time-series of median rent (all bedroom sizes) for Seattle, 1997-2018.
 Historical values adjusted with CPI; Craigslist estimates are quarterly, Dupre+Scott are semi-annual (March and September), and ACS are annual (i.e.
-respective 1-Year Estimates for each year period)._](../output/rent_lineplot.png)
+respective 1-Year Estimates for each year period).](../output/rent_lineplot.png)
 
@@
 While zip codes, unofficial City neighborhoods and even arbitrary grid geographies can all be used to aggregate micro-data about rental housing, we compare census tract estimates because they roughly approximate neighborhood-size units and can be linked with a variety of other data in Seattle and more generally.
-The following links provide reference maps for Seattle Census tracts ([2000](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2000CensusSeattleTractMap.pdf){target="_blank"} and [2010](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2010CensusTractsandZipCodeBoundaries.pdf){target="_blank"})
+The following links provide reference maps for Seattle Census tracts ([2000](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2000CensusSeattleTractMap.pdf) and [2010](https://www.seattle.gov/Documents/Departments/OPCD/Demographics/GeographicFilesandMaps/2010CensusTractsandZipCodeBoundaries.pdf))
 
@@
 
-![**Figure 4.** _Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates.
-Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled._](../output/rent_map_across_1B.png)
+![Chloropleth map of neighborhood 1-bedroom rents across Craigslist, Dupre+Scott and ACS 2012-2016 5-Year Estimates.
+Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties, while ACS releases median gross (asking) rents for all one-bedroom renter-occupied housing units sampled.](../output/rent_map_across_1B.png)
 
@@
 
-![**Figure 5.** _Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott.
+![Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott.
 Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.
-Purple dashed line indicates loess fit._](../output/rent_scatter_across_1B.png)
+Purple dashed line indicates loess fit.](../output/rent_scatter_across_1B.png)
 
@@
 These tracts are found towards the North and South edges of the city, and generally are more residential neighborhoods where 20+ unit apartment complexes are not the entire stock of rental unit properties.
-In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf){target="_blank"}.
+In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf).
 
@@
 
-We assume that the log of the median one-bedroom rent for Seattle neighborhoods is normally distributed and compute the posterior distributions of each model's hyperparameters and parameters $\theta_{it}$ using [integrated nested Laplace approximations (INLA)](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf){target="_blank"} with the default, weakly-informative priors for all random-effects.
+We assume that the log of the median one-bedroom rent for Seattle neighborhoods is normally distributed and compute the posterior distributions of each model's hyperparameters and parameters $\theta_{it}$ using [integrated nested Laplace approximations (INLA)](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) with the default, weakly-informative priors for all random-effects.
 We choose to log the dependent variable because all values are positive and this transformation improves model fit.
@@
 
-The BYM Spatial with AR(1) is the [Besag-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466){target="_blank"} intrinsic conditional autoregressive (ICAR) spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure (referred to as `bym` using shorthand).
+The BYM Spatial with AR(1) includes a random-effect drawn from the [Besag-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466) intrinsic conditional autoregressive (ICAR) distribution, along with AR(1) and exchangeable hyperparameters for the panel's temporal structure (referred to as `bym` using shorthand).
 The random-effect $v_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $v_i|v_j, j \in ne(i) ~ \mbox{Normal}(\bar{v_i}, \frac{\sigma^{2}_v}{m_j})$.
@@
 
-Visit the interactive results dashboard [here](http://hesscl.com/smooth-sea/dashboard/index.html){target="_blank"}.
+Visit the interactive results dashboard [here](http://hesscl.com/smooth-sea/dashboard/index.html).
 
@@
 So while we want to minimize model error as a way to judge if our models improve on each other, we also want to smooth over sampling variability in the Craigslist data that induces local deviations compared to the true median rent.
-The [information criteria for the models](http://hesscl.com/smooth-sea/dashboard/index.html#model-fit){target="_blank"} generally show that the spatial specifications are more likely models of the neighborhood rent data generating process.
+The [information criteria for the models](http://hesscl.com/smooth-sea/dashboard/index.html#model-fit) generally show that the spatial specifications are more likely models of the neighborhood rent data generating process.
 
@@
 
-![**Figure 6.** _Lineplot of Spatial AR(1) model predicted values.
-Colors denote a tract's deviation from the average neighborhood 1B rent level on a quarter-by-quarter basis._](../output/bym_lineplot.png)
+![Lineplot of Spatial AR(1) model predicted values. Colors denote a tract's deviation from the average neighborhood 1B rent level on a quarter-by-quarter basis.](../output/bym_lineplot.png)
 
@@
 
-![**Figure 7.** _Chloropleth maps of actual versus predicted values (BYM Spatial AR(1) model) for the Q1 2017 to Q2 2018 period._](../output/avp_map_series.png)
+![Chloropleth maps of actual versus predicted values (BYM Spatial AR(1) model) for the Q1 2017 to Q2 2018 period.](../output/avp_map_series.png)
 
@@
 
-This study highlights the importance of considering scraped web data for understanding the dynamics of neighborhood rental housing costs.
-Throughout this work we demonstrate how Craigslist rental housing listings scraped may be used to research the estimated rent for the Seattle market and its neighborhoods.
-Using Bayesian spatial smoothing models further exploit the scraped data's spatial and temporal richness, giving researchers an option for studying rental housing market dynamics on a more granular scale where time is not only considered in terms of trend, but also random shock and seasonality.
+This study highlights the importance of considering scraped web data for research about the dynamics of neighborhood rental housing costs.
+Throughout this work we demonstrate how Craigslist rental housing listings may be used to research the estimated rent for the Seattle market and its neighborhoods.
+Using Bayesian spatial smoothing models further exploits the scraped data's spatial and temporal richness, giving researchers an option for studying rental housing market dynamics on a more granular scale where time is not only considered in terms of trend, but also random shock and seasonality.
 
-Finally, this paper has endeavored (perhaps a little too much...) to provide an example of using scraped rental housing listings for research, but also to model how Rmarkdown, Git and GitHub are transformative  for our disciplines.
-There are no free lunches with reproducibility: these open-source technologies demand that we learn how to incorporate them into our work.
-To the extent that this requires thinking about analysis beyond a user base of 1, there can be some growing pains.
-However, there are some positive spillovers beyond improving the _science_ of social science---including new platforms and media for sharing research findings.
+Finally, this paper has endeavored to provide a model of how Rmarkdown, Git and GitHub can be transformative for social science disciplines.
+There are no free lunches with reproducibility: these open-source technologies demand that researchers learn how to incorporate them into our work.
+While there can be growing pains to this process, there are some positive spillovers beyond improving the _science_ of social science---including new platforms and media for sharing research findings.
 
@@
 New insights into rental housing markets across the united states: web scraping and analyzing craigslist rental listings.
-Journal of Planning Education and Research, 37(4), pp.457-476.](http://journals.sagepub.com/doi/abs/10.1177/0739456X16664789){target="_blank"}
+Journal of Planning Education and Research, 37(4), pp.457-476.](http://journals.sagepub.com/doi/abs/10.1177/0739456X16664789)
 
@@
 and Mollie, A., 1991.
-Bayesian image restoration with two applications in spatial statistics (with discussion) Annals of the Institute of Statistical Mathematics.
-43, pp.1-59.](https://link.springer.com/article/10.1007/BF00116466){target="_blank"}
+Bayesian image restoration with two applications in spatial statistics (with discussion).
+Annals of the Institute of Statistical Mathematics.
+43, pp.1-59.](https://link.springer.com/article/10.1007/BF00116466)
 
@@
 Spatial and spatio-temporal models with R-INLA.
-Spatial and spatio-temporal epidemiology, 4, pp.33-49.](https://www.sciencedirect.com/science/article/pii/S1877584512000846){target="_blank"}
+Spatial and spatio-temporal epidemiology, 4, pp.33-49.](https://www.sciencedirect.com/science/article/pii/S1877584512000846)

### 2018-07-03 dashboard/index.rmd
committing work through 7/3: the data are now full, i.e. Q1 2017 through end of Q2 2018. The manuscript .Rmds are effectively synced (index.rmd = manuscript.rmd with a different header). Also, cleaned up the helpers so all apply the same data load procedure

@@
   mutate(actualRent = med1B, #first preserve the observed / observed-to-date
-         med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B)) %>% #set to NA for INLA to forecast current qtr
+         med1B = ifelse(listingQtr == "2018 Q2", NA, med1B)) %>% #set to NA for INLA to forecast current qtr
   mutate(p1B = n1B/nListings) %>% #compute share of 1B listings
@@
 fit_stats <- sea_df %>%
-  mutate(train_test = ifelse(listingQtr == as.yearqtr(Sys.Date()), "Test", "Training")) %>%
+  mutate(train_test = ifelse(listingQtr == "2018 Q2", "Test", "Training")) %>%
   mutate(int_err = actualRent - int_Med,
@@
 All of the candidate models' posterior distributions are estimated with [integrated nested Laplace approximations (INLA)](https://folk.ntnu.no/hrue/r-inla.org/papers/inla-rss.pdf) using the default, weakly-informative priors for all model hyperparameters.
-Throughout the following analyses, the training data are `r min(sea_df$listingQtr)` up to the prior quarter (i.e.
-`r as.yearqtr(Sys.Date())-.25`).
-The test period is a forecast for the current period and includes comparison to the appropriate median rent estimates for data observed so far in this period.
+Throughout the following analyses, the training data are 2017 Q1 up through 2018 Q1.
+The test period is a forecast for 2018 Q2 and includes comparison to the observed median rent estimates for data collected in this period.
 
@@
   mutate(rent = actualRent) %>%
-  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
-  mutate(var = paste0(as.yearqtr(Sys.Date()), "-to-Date")) %>%
+  filter(listingQtr == "2018 Q2") %>%
+  mutate(var = "2018 Q2 Observed") %>%
   dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, bym_Med, NAMELSAD10, n1B)
@@
   mutate(rent = bym_Med) %>%
-  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
-  mutate(var = paste0(as.yearqtr(Sys.Date()), " Forecast")) %>%
+  filter(listingQtr == "2018 Q2") %>%
+  mutate(var = "2018 Q2 Smoothed") %>%
   dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, bym_Med, NAMELSAD10, n1B)
@@
 comb <- bind_rows(obs, pred) %>%
-  mutate(tooltip = paste0(NAMELSAD10, "<br>", "Forecast: $", round(bym_Med), "<br>", "Qtr-to-Date: $", round(actualRent), "<br>", "N Qtr-to-Date: ", n1B))
+  mutate(tooltip = paste0(NAMELSAD10, "<br>", "Smoothed: $", round(bym_Med), "<br>", "Observed: $", round(actualRent), "<br>", "N Observed: ", n1B))
 comb <- SharedData$new(comb, ~GISJOIN)

### 2018-07-03 manuscript/index.Rmd
committing work through 7/3: the data are now full, i.e. Q1 2017 through end of Q2 2018. The manuscript .Rmds are effectively synced (index.rmd = manuscript.rmd with a different header). Also, cleaned up the helpers so all apply the same data load procedure

@@
 author: "Chris Hess and Sarah Chasins"
-date: "June 5, 2018"
 output: 
@@
 The second component of the paper investivates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models.
-All models' posterior distribution are estimated with integrated nested Laplace approximations (INLA).
-Training data for models cover quarterly tract estimates for 2017 Q2 through 2018 Q1, while the test data are 2018 Q2 median values as observed through June 4th.
+Posterior distributions of all model hyperparameters and parameters are estimated with integrated nested Laplace approximations (INLA).
+Training data for models cover quarterly tract estimates for Q2 2017 through Q1 2018, while the test data are Q2 2018 median values as observed.
 The top model assumes that spatial and temporal dynamics are separable, i.e.
@@
 Both data suggest continued increases in the market-wide median rent for the first half of 2017, though the Craigslist slope is about twice in its size.
-Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for 2017 Q2 upward in favor of regularly posted commercial listings.
+Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for Q2 2017 upward in favor of regularly posted commercial listings.
 Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability.
@@
 However, by the final point of Dupre+Scott's series (Fall 2017, corresponding to September), the difference compared to the Craigslist median was about $75.
-Mirroring other accounts about a decline in rents during Winter 2018, the Craigslist data showed a fairly precipitous decline between
+Mirroring other accounts about a decline in rents during Winter 2018, the Craigslist data showed a fairly precipitous decline between Q4 2017 and Q1 2018.
 
@@
 The purple dashed line indicates the loess smoother fit between the measures.
-Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and good agreement between the measures (mean absolute difference = $186).
+Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and relative agreement between the measures (mean absolute difference = $186).
 The Craigslist measure reflects asking rents and the Dupre+Scott measure reflects contract rents for a neighborhood (i.e.
@@
 These tracts are found towards the North and South edges of the city, and generally are more residential neighborhoods where 20+ unit apartment complexes are not the entire stock of rental unit properties.
-In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf){target="_blank"}.
+In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf).
 
-To the extent that these smaller property units may tend to be newer, larger in housing unit square footage or offer added amenities, the true value of rent for the neighborhood will be higher than what it is estimated at using 20+ unit properties alone.
+To the extent that units in smaller properties (in terms of the number of units) may tend to be newer, larger in housing unit square footage or offer added amenities, the true value of rent for the neighborhood will be higher than what it is estimated at using 20+ unit properties alone.
 Similarly, the more that smaller properties account for the overall population of rental housing units, the more likely that higher values indicated in Craigslist are not erroneous but instead reflective of a broader set of residential units.
@@
 
-The following models learn from Craigslist data accumulated up till the prior quarter (i.e.
-`r as.yearqtr(Sys.Date())-1`) to accomplish a few different goals.
+The following models learn from Craigslist data accumulated from Q1 2017 Q1 through Q1 2018 to accomplish a few different goals.
 The most prominent goal is to use the panel data's temporal and spatial structure to understand variation in rent levels over time and across space.
@@
 
-The BYM Spatial with AR(1) is the [Besag-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466){target="_blank"} intrinsic conditional autoregressive spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure (referred to as `bym` using shorthand).
-The random-effect $S_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $v_i|v_j, j \in ne(i) ~ \mbox{Normal}(\bar{v_i}, \frac{\sigma^{2}_v}{m_j})$.
+The BYM Spatial with AR(1) includes a random-effect drawn from the [Besag-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466) intrinsic conditional autoregressive (ICAR) distribution, along with AR(1) and exchangeable hyperparameters for the panel's temporal structure (referred to as `bym` using shorthand).
+The random-effect $v_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $v_i|v_j, j \in ne(i) ~ \mbox{Normal}(\bar{v_i}, \frac{\sigma^{2}_v}{m_j})$.
 Lastly, the BYM Spatiotemporal AR(1) (`spt`) adds an exchangeable random-effect ($\delta_i \sim Normal(0, \tau_\delta)$) for each period to allow for area-specific deviations in time trends.
@@
 
-#spatial random effect model (BYM 1991) with AR(1)
+#spatial random effect model (BYM 1991) with quarterly seasonal fixed-effect
 form.bym <- log(med1B) ~ 1 + 
@@
 The following tables show in-sample and out-of-sample (i.e.
-2018 Q2) accuracy for the 5 model specifications in terms of root mean squared-error (RMSE) and mean absolute error (MAE).
+Q2 2018) accuracy for the 5 model specifications in terms of root mean squared-error (RMSE) and mean absolute error (MAE).
 Note how though the Non-Spatial models (i.e.

### 2018-07-03 helpers/rent_map_across.R

@@ -21,7 +21,7 @@ cl2017 <- rent_map_sea %>%
   filter(seattle==1) %>% 
   collect %>% 
   mutate(listingQtr = as.yearqtr(listingDate)) %>%
-  filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3",
+  filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3", matchType != "Google Maps Lat/Long",
          !is.na(GISJOIN), !is.na(matchAddress), !is.na(matchAddress2), !is.na(catBeds), !is.na(cleanSqft), !is.na(cleanRent),
          catBeds %in% c("0", "1", "2")) %>%
   distinct(matchAddress, matchAddress2, catBeds, cleanSqft, cleanRent, .keep_all = T) %>%

### 2018-07-03 helpers/rent_scatter_across.R

@@ -66,7 +66,7 @@ across2017 <- left_join(across2017, ds2017)
 pal <- brewer.pal(3, "Purples")
 
 ggplot(across2017, aes(x = dsRent, y = clRent, size = nHU)) + 
-  geom_point() +
+  geom_point(alpha = .6) +
   geom_abline(slope = 1, intercept = 0) +
   geom_smooth(show.legend = FALSE, se=FALSE, linetype = 5, 
               color = viridis(3, begin = .25, option = "A")[1], lwd = 1.25) +

### 2018-07-04 manuscript/index.Rmd
implemented use of load_data() helper throughout, all db queries now follow same structure via fn

@@
 
+Only the Craigslist and Dupre+Scott can offer insight into more recent median levels for the Seattle rental housing market.
+Both data suggest continued increases in the market-wide median rent for the first half of 2017, though the Craigslist slope is about twice in its size.
+Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for Q2 2017 upward in favor of regularly posted commercial listings.
+Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability.
+This volatility in Craigslist asking rents could also be brought on by recently-developed buildings in "lease-up" adding a glut of expensive listings to the quarters' samples.
+
+However, by the final point of Dupre+Scott's series (Fall 2017, corresponding to September), the difference compared to the Craigslist median was about $75.
+Mirroring other accounts about a decline in rents during Winter 2018, the Craigslist data showed a fairly precipitous decline between Q4 2017 and Q1 2018.
+
 <br>
@@
 respective 1-Year Estimates for each year period).](../output/rent_lineplot.png)
-
-<br>
-
-Only the Craigslist and Dupre+Scott can offer insight into more recent median levels for the Seattle rental housing market.
-Both data suggest continued increases in the market-wide median rent for the first half of 2017, though the Craigslist slope is about twice in its size.
-Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for Q2 2017 upward in favor of regularly posted commercial listings.
-Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability.
-This volatility in Craigslist asking rents could also be brought on by recently-developed buildings in "lease-up" adding a glut of expensive listings to the quarters' samples.
-
-However, by the final point of Dupre+Scott's series (Fall 2017, corresponding to September), the difference compared to the Craigslist median was about $75.
-Mirroring other accounts about a decline in rents during Winter 2018, the Craigslist data showed a fairly precipitous decline between Q4 2017 and Q1 2018.
 
@@
 
+Both the Craigslist and Dupre+Scott data largely agree on the spatial distribution of one-bedroom rents.
+The similarity is greatest in the downtown and higher-density neighborhoods where the stock of corporate-held rental housing units theoretically accounts for a most of a tract's rental housing units (i.e.
+Dupre+Scott closer to a census if they sample all 20+ unit properties).
+Nonetheless, the systematic difference is such that Craigslist levels are typically at least a little higher.
+A detailed comparison of the Craigslist and Dupre+Scott estimates follows in the next figure where we directly compare 2000-definition census tract estimates from Craigslist and Dupre+Scott.
+
+Though the spatial distribution of the ACS estimates resembles that of the other two data sources, the values are nonetheless systematically lower (in most cases by hundreds of dollars).
+The ACS measures therefore give insight into the relative ranking of neighborhoods in terms of rental costs, but are limited for studying neighborhood dynamics outside of long-run trend.
+
 <br>
@@
 <br>
-
-Both the Craigslist and Dupre+Scott data largely agree on the spatial distribution of one-bedroom rents.
-The similarity is greatest in the downtown and higher-density neighborhoods where the stock of corporate-held rental housing units theoretically accounts for a most of a tract's rental housing units (i.e.
-Dupre+Scott closer to a census if they sample all 20+ unit properties).
-Nonetheless, the systematic difference is such that Craigslist levels are typically at least a little higher.
-A detailed comparison of the Craigslist and Dupre+Scott estimates follows in the next figure where we directly compare 2000-definition census tract estimates from Craigslist and Dupre+Scott.
-
-Though the spatial distribution of the ACS estimates resembles that of the other two data sources, the values are nonetheless systematically lower (in most cases by hundreds of dollars).
-The ACS measures therefore give insight into the relative ranking of neighborhoods in terms of rental costs, but are limited for studying neighborhood dynamics outside of long-run trend.
-
-<br>
 <hr>
@@
 
-Figure 5 shows a scatter plot/bubble plot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 75 tracts where comparable estimates exist.
+Figure 5 shows a bubbleplot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 80 tracts where comparable estimates exist (i.e.
+about 2/3 of the 124 tracts in Seattle using 2000 definitions).
 The purple dashed line indicates the loess smoother fit between the measures.
@@
 At the high-end of the neighborhood rent distribution, however, Craigslist and Dupre+Scott are in arguably the strongest agreement.
+
+An alternative explanation for the difference between measures comes in considering the units these properties are sampling from in a given neighborhood.
+The greater difference between the measures among lower-rent neighborhoods (with Craigslist estimating these as more expensive than Dupre+Scott) provides some evidence for this theory.
+These tracts are found towards the North and South edges of the city, and generally are more residential neighborhoods where 20+ unit apartment complexes are not the entire stock of rental unit properties.
+In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf).
+
+To the extent that units in smaller properties (in terms of the number of units) may tend to be newer, larger in housing unit square footage or offer added amenities, the true value of rent for the neighborhood will be higher than what it is estimated at using 20+ unit properties alone.
+Similarly, the more that smaller properties account for the overall population of rental housing units, the more likely that higher values indicated in Craigslist are not erroneous but instead reflective of a broader set of residential units.
 
@@
 
-![Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott.
+![Bubbleplot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott.
 Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.
@@
 <br>
-
-An alternative explanation for the difference between measures comes in considering the units these properties are sampling from in a given neighborhood.
-The greater difference between the measures among lower-rent neighborhoods (with Craigslist estimating these as more expensive than Dupre+Scott) provides some evidence for this theory.
-These tracts are found towards the North and South edges of the city, and generally are more residential neighborhoods where 20+ unit apartment complexes are not the entire stock of rental unit properties.
-In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf).
-
-To the extent that units in smaller properties (in terms of the number of units) may tend to be newer, larger in housing unit square footage or offer added amenities, the true value of rent for the neighborhood will be higher than what it is estimated at using 20+ unit properties alone.
-Similarly, the more that smaller properties account for the overall population of rental housing units, the more likely that higher values indicated in Craigslist are not erroneous but instead reflective of a broader set of residential units.
-
+<hr>
 <br>

### 2018-07-04 manuscript/manuscript.Rmd
implemented use of load_data() helper throughout, all db queries now follow same structure via fn

@@
 author: "Chris Hess and Sarah Chasins"
-date: "June 5, 2018"
 output:  pdf_document
@@
 Posterior distributions of all model hyperparameters and parameters are estimated with integrated nested Laplace approximations (INLA).
-Training data for models cover quarterly tract estimates for 2017 Q2 through 2018 Q1, while the test data are 2018 Q2 median values as observed through June 4th.
+Training data for models cover quarterly tract estimates for Q2 2017 through Q1 2018, while the test data are Q2 2018 median values as observed.
 The top model assumes that spatial and temporal dynamics are separable, i.e.
@@
 respective 1-Year Estimates for each year period).](../output/rent_lineplot.png)
-
-<br>
-
-Only the Craigslist and Dupre+Scott can offer insight into more recent median levels for the Seattle rental housing market.
-Both data suggest continued increases in the market-wide median rent for the first half of 2017, though the Craigslist slope is about twice in its size.
-Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for 2017 Q2 upward in favor of regularly posted commercial listings.
-Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability.
-This volatility in Craigslist asking rents could also be brought on by recently-developed buildings in "lease-up" adding a glut of expensive listings to the quarters' samples.
-
-However, by the final point of Dupre+Scott's series (Fall 2017, corresponding to September), the difference compared to the Craigslist median was about $75.
-Mirroring other accounts about a decline in rents during Winter 2018, the Craigslist data showed a fairly precipitous decline between
 
@@
 
-Figure 5 shows a scatter plot/bubble plot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 75 tracts where comparable estimates exist.
+Figure 5 shows a scatter plot/bubble plot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 80 tracts where comparable estimates exist (i.e.
+about 2/3 of the 124 2000-definition tracts in Seattle).
 The purple dashed line indicates the loess smoother fit between the measures.
-Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and good agreement between the measures (mean absolute difference = $186).
+Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and relative agreement between the measures (mean absolute difference = $186).
 The Craigslist measure reflects asking rents and the Dupre+Scott measure reflects contract rents for a neighborhood (i.e.
@@
 <br>
-
-An alternative explanation for the difference between measures comes in considering the units these properties are sampling from in a given neighborhood.
-The greater difference between the measures among lower-rent neighborhoods (with Craigslist estimating these as more expensive than Dupre+Scott) provides some evidence for this theory.
-These tracts are found towards the North and South edges of the city, and generally are more residential neighborhoods where 20+ unit apartment complexes are not the entire stock of rental unit properties.
-In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf).
-
-To the extent that these smaller property units may tend to be newer, larger in housing unit square footage or offer added amenities, the true value of rent for the neighborhood will be higher than what it is estimated at using 20+ unit properties alone.
-Similarly, the more that smaller properties account for the overall population of rental housing units, the more likely that higher values indicated in Craigslist are not erroneous but instead reflective of a broader set of residential units.
-
+<hr>
 <br>

### 2018-07-04 helpers/load_data.R

@@ -1,38 +1,47 @@
-load_data <- function(){
+load_data <- function(listings = FALSE){
   #determine config (dev or github)
   if(file.exists("../../data/cl/craigslistDB.sqlite")){
     #connect to database if file.exists
     DB <- dbConnect(SQLite(), dbname="../../data/cl/craigslistDB.sqlite")
     cl <- tbl(DB, "clean") #create dbi for clean listing table
     
-    #compute tract aggregates for CL listing count
-    tract <- cl %>%
+    listing <- cl %>%
       collect %>% #bring db query into memory
       filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), !is.na(matchAddress), !is.na(matchAddress2),
              GISJOIN %in% sea_shp@data$GISJOIN, #only listings with valid Bed/Rent, seattle tracts
-             matchType != "Google Maps Lat/Long") %>% #need to have address, not approximate
+             matchType != "Google Maps Lat/Long", #need to have address, not approximate
+             cleanBeds %in% c(0, 1, 2, 3)) %>% 
       distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
                matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
       dplyr::select(listingDate, GISJOIN, seattle, matchAddress, matchType, 
-                    cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
+                    cleanBeds, catBeds, cleanRent, cleanSqft, lat, lng) %>% #SELECT these columns
       mutate(listingDate = as.Date(listingDate),
              listingQtr = as.yearqtr(listingDate)) %>%
-      filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3") %>%
-      group_by(GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
-      summarize(nListings = n(),
-                n1B = sum(cleanBeds == 1),
-                medRent = median(cleanRent),
-                med0B = median(cleanRent[cleanBeds==0]),
-                med1B = median(cleanRent[cleanBeds==1]),
-                med2B = median(cleanRent[cleanBeds==2]),
-                med3B = median(cleanRent[cleanBeds==3])) %>% #create tract aggregates
-      ungroup %>%
-      arrange(GISJOIN, listingQtr)
-    dbDisconnect(DB)
-    return(tract)
[…]

### 2018-07-04 helpers/map_point_density.R

@@ -5,30 +5,17 @@
 library(ggmap)
 library(patchwork)
 
-#data
-DB <- dbConnect(SQLite(), dbname="R:/Project/seattle_rental_market/data/cl/craigslistDB.sqlite")
-map_clean <- tbl(DB, "clean") #clean seattle listing table
-map_raw <- tbl(DB, "raw")
-
 #load table
-map_cl <- map_clean %>% 
-  filter(seattle==1) %>%
-  collect %>% #bring db query into memory
-  filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), !is.na(matchAddress), !is.na(matchAddress2),
-         GISJOIN %in% sea_shp@data$GISJOIN, #only listings with valid Bed/Rent, seattle tracts
-         matchType != "Google Maps Lat/Long") %>% #need to have address, not approximate
-  distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
-           matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
-  dplyr::select(listingDate, GISJOIN, seattle, matchAddress, matchType, 
-                cleanBeds, cleanRent, cleanSqft, lat, lng) %>% #SELECT these columns
-  mutate(listingDate = as.Date(listingDate),
-         listingQtr = as.yearqtr(listingDate)) %>%
-  filter(cleanBeds %in% c(1), listingQtr >= "2017 Q1", listingQtr < "2018 Q3")
+map_cl <- load_data(listings = TRUE)
 
-dbDisconnect(DB)
+#filter to 1BD listings
+map_cl <- map_cl %>%
+  filter(cleanBeds == 1)
 
+#create left panel
 blank <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", geom = "blank") 
 
+#create right panel
 with_points <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", darken = .50, legend = "bottom") +
   stat_density_2d(aes(fill = ..level..), geom = "polygon",
                   alpha = .75, color = NA, size = .05) +
@@ -43,8 +30,7 @@ with_points <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", darken = 
   guides(fill = guide_colourbar(barwidth = 7,
                               label.position = "bottom"))
 
+#append panels together and save
 blank + with_points +
   ggsave(filename = "../output/map_point_density.png",
          width = 6, height = 5.5, units = "in", dpi = 300)
-
-

### 2018-07-04 helpers/n_heatmap.R

@@ -1,23 +1,11 @@
-
-#connect to db
-DB <- dbConnect(SQLite(), dbname="R:/Project/seattle_rental_market/data/cl/craigslistDB.sqlite")
-cl <- tbl(DB, "clean") #clean listing table
-
 #compute tract aggregates for CL listing count
-sea_cl <- cl %>%
-  dplyr::select(listingMoYr, listingDate, GISJOIN, seattle, matchAddress, matchAddress2, 
-                matchType, cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
-  collect %>% #bring db query into memory
-  filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), !is.na(matchAddress), !is.na(matchAddress2),
-         GISJOIN %in% sea_shp@data$GISJOIN, #only listings with valid Bed/Rent, seattle tracts
-         matchType != "Google Maps Lat/Long") %>% #need to have address, not approximate
-  mutate(listingDate = as.Date(listingDate),
-         listingQtr = as.yearqtr(listingDate)) %>%
-  filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3",
-         cleanBeds == 1) %>% 
-  distinct(matchAddress, matchAddress2, cleanBeds, cleanRent, cleanSqft, .keep_all = T)
-dbDisconnect(DB)
+sea_cl <- load_data(listings = TRUE)
 
+#filter to 1BDs
+sea_cl <- sea_cl %>% 
+  filter(cleanBeds == 1)
+
+#function to compute the daily ts for n with missingness allowed
 computeCal <- function(x){
   y <- x %>%
     group_by(listingDate) %>%
@@ -38,7 +26,7 @@ computeCal <- function(x){
   n <- cbind.data.frame(n, fullDates)
 }
 
-
+#compute the daily ts of n listings, rework fields to what is needed for gg
 cal <- computeCal(sea_cl) %>%
   mutate(weekdayf = weekdays(fullDates) %>% as.factor,
          monthf = month(fullDates, label = T),
@@ -47,10 +35,11 @@ cal <- computeCal(sea_cl) %>%
          week = as.numeric(format(fullDates,"%W")))
 sum(is.na(cal$n))/nrow(cal)
 
+#order the days of week to match graphic format
 cal$weekdayf <- factor(cal$weekdayf, levels = rev(levels(cal$weekdayf)[c(2,6,7,5,1,3,4)]))
 
[…]

### 2018-07-04 helpers/rent_lineplot.R

@@ -1,31 +1,18 @@
 
 ### Helper script to produce across data lineplot of median rent for Seattle
 
-#data
-DB <- dbConnect(SQLite(), dbname="R:/Project/seattle_rental_market/data/cl/craigslistDB.sqlite")
-sea <- tbl(DB, "clean") #clean seattle listing table
-raw <- tbl(DB, "raw")
-
-#for code that has not been updated to db calls
-cl <- sea %>% collect %>% filter(seattle==1)
+#query listings
+cl <- load_data(listings = TRUE)
 
 #load Seattle market summaries from D+S
 ds <- read_csv(file = "R:/Project/seattle_rental_market/data/d+s/city-level D+S.csv")
 
-
 #### A. Prepare a table for ggplot ---------------------------------------------
 
 sumCL <- cl %>%
-  #mutate(cleanRent = ifelse(cleanRent > 5000, NA, cleanRent)) %>%
-  mutate(catSqft = fct_reorder(catSqft, cleanSqft)) %>%
+  mutate(cleanRent = ifelse(year(listingDate) == 2017, cleanRent * 1.02, cleanRent)) %>%
   arrange(listingDate) %>%
   select(-ends_with("10")) %>%
-  filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), !is.na(matchAddress), !is.na(matchAddress2),
-         GISJOIN %in% sea_shp@data$GISJOIN, #only listings with valid Bed/Rent, seattle tracts
-         matchType != "Google Maps Lat/Long") %>% #need to have address, not approximate
-  distinct(matchAddress, matchAddress2, catBeds, cleanSqft, cleanRent, .keep_all = T) %>%
-  mutate(listingQtr = as.yearqtr(as.Date(listingDate))) %>%
-  filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3") %>%
   group_by(listingQtr) %>%
   summarize(Rent = median(cleanRent, na.rm=T),
             nHU = n()) %>%

### 2018-07-04 helpers/rent_map_across.R

@@ -4,9 +4,7 @@
 rent_map_acs <- read_csv(file = "R:/Project/seattle_rental_market/data/census/acs/nhgis0051_ds226_20165_2016_tract.csv")
 
 #connect to craigslist data
-DB <- dbConnect(SQLite(), dbname="R:/Project/seattle_rental_market/data/cl/craigslistDB.sqlite")
-rent_map_sea <- tbl(DB, "clean") #clean seattle listing table
-rent_map_raw <- tbl(DB, "raw")
+rent_map_sea <- load_data(listings = TRUE)
 
 #load d+s data
 rent_map_ds <- read_csv("R:/Project/seattle_rental_market/data/d+s/geocoded_D+S.csv")
@@ -16,27 +14,20 @@ tract2010_bg <- fortify(tract2010)
 
 #### A. Prep CL table ---------------------------------------------------------
 
-#2017-2018 listings in seattle with non-missing lat/lng, drop 2010 geog fields
+#2017-2018 listings in seattle with non-missing lat/lng
 cl2017 <- rent_map_sea %>% 
-  filter(seattle==1) %>% 
-  collect %>% 
-  mutate(listingQtr = as.yearqtr(listingDate)) %>%
-  filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3", matchType != "Google Maps Lat/Long",
-         !is.na(GISJOIN), !is.na(matchAddress), !is.na(matchAddress2), !is.na(catBeds), !is.na(cleanSqft), !is.na(cleanRent),
-         catBeds %in% c("0", "1", "2")) %>%
-  distinct(matchAddress, matchAddress2, catBeds, cleanSqft, cleanRent, .keep_all = T) %>%
+  filter(cleanBeds %in% c(0, 1, 2)) %>% #this is inconsequential, just to preserve studio/2BD comparisons for future
   group_by(GISJOIN, catBeds) %>%
   summarize(rent = median(cleanRent, na.rm = T),
             nHU = n(),
             data = "Craigslist 2017-2018") %>%
-  select(GISJOIN, data, catBeds, rent, nHU)
-dbDisconnect(DB)
+  dplyr::select(GISJOIN, data, catBeds, rent, nHU)
 
[…]

### 2018-07-04 helpers/rent_scatter_across.R

@@ -1,9 +1,7 @@
 #### Helper script to compute scatter plot of CL X D+S
 
 #data
-DB <- dbConnect(SQLite(), dbname="R:/Project/seattle_rental_market/data/cl/craigslistDB.sqlite")
-scatter_sea <- tbl(DB, "clean") #clean seattle listing table
-scatter_raw <- tbl(DB, "raw")
+scatter_sea <- load_data(listings = TRUE)
 
 #read in D+S tract summaries
 scatter_ds <- read_csv("R:/Project/seattle_rental_market/data/d+s/geocoded_D+S.csv")
@@ -11,17 +9,8 @@ scatter_ds <- read_csv("R:/Project/seattle_rental_market/data/d+s/geocoded_D+S.c
 #### A. Prepare a table for ggplot ---------------------------------------------
 
 scatter_cl <- scatter_sea %>% 
-  filter(seattle==1) %>%
-  collect %>% #bring db query into memory
-  filter(!is.na(GISJOIN), !is.na(matchAddress), !is.na(matchAddress2), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
-         GISJOIN %in% tract2000@data$GISJOIN, matchType != "Google Maps Lat/Long") %>% #only listings with valid Bed/Rent, seattle tracts
-  distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
-           matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
-  dplyr::select(listingYear, listingDate, GISJOIN, seattle, matchAddress, matchType, 
-                cleanBeds, cleanRent, cleanSqft, lat, lng) %>% #SELECT these columns
-  mutate(listingDate = as.Date(listingDate),
-         listingQtr = as.yearqtr(listingDate)) %>%
-  filter(cleanBeds %in% c(1), listingQtr >= "2017 Q1", listingQtr < "2018 Q3")
+  filter(cleanBeds == 1) %>%
+  select(-GISJOIN)
 
 coordinates(scatter_cl) <- cbind(scatter_cl$lng, scatter_cl$lat)
 
@@ -58,7 +47,6 @@ cl2017 <- scatter_cl %>%
   group_by(GISJOIN) %>%
   summarize(clRent = median(cleanRent[cleanBeds==1], na.rm =T),
             nHU = n()) %>%
-  #filter(nHU >= 15) %>%
[…]

### 2018-07-27 manuscript/index.Rmd
adding section break tags after models, results
Small typo-type edits.

@@
 This research first assesses the agreement of census tract rent estimates based on scraped rental advertisements with existing data sources.
-The second component of the paper investivates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models.
+The second component of the paper investigates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models.
 Posterior distributions of all model hyperparameters and parameters are estimated with integrated nested Laplace approximations (INLA).
@@
 Figure 2 is a heatmap for the time-series of one-bedroom listings per day.
-The grey cells indicate days where our scraper was offline.
+The gray cells indicate days where our scraper was offline.
 Outages were more common in the earliest phases of work, but now are relatively rare.
@@
 Finally, on the right are ACS 2012-2016 5-Year Estimates (the most recent ACS small-area data product available).
-In all map panes, grey cells denote missing values.
+In all map panes, gray cells denote missing values.
 
@@
 about 2/3 of the 124 tracts in Seattle using 2000 definitions).
-The purple dashed line indicates the loess smoother fit between the measures.
+The purple dashed line indicates the LOESS smoother fit between the measures.
 Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and relative agreement between the measures (mean absolute difference = $186).
@@
 Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties.
-Purple dashed line indicates loess fit.](../output/rent_scatter_across_1B.png)
+Purple dashed line indicates LOESS fit.](../output/rent_scatter_across_1B.png)
 
@@
 These models are common in epidemiology and disease-mapping applications where researchers need to estimate a quantity of interest like a relative risk but have sparse data for some areas (which may differ from truth due to sampling variability).
-Modelling a panel's error structure with random-effects for space and time penalizes bumpiness (local deviations) in predicted values and theoretically improves out-of-sample accuracy (because large local deviations can stem from low $n$ in a given area).
+Modeling a panel's error structure with random-effects for space and time penalizes bumpiness (local deviations) in predicted values and theoretically improves out-of-sample accuracy (because large local deviations can stem from low $n$ in a given area).
 The spatial models in particular allow for missing data to be imputed based off of a local average of neighbors rather than the global or period-specific average.
+
+<br>
+<hr>
+<br>
 

### 2018-07-27 manuscript/manuscript.Rmd
adding section break tags after models, results
Small typo-type edits.

@@
 author: "Chris Hess and Sarah Chasins"
-output:  pdf_document
-abstract: "Online classified ads are a key source of information about local rental housing markets.
-This research first assesses the agreement of census tract rent estimates based on scraped rental advertisements with existing data sources.
-The second component of the paper investivates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models.
-Posterior distributions of all model hyperparameters and parameters are estimated with integrated nested Laplace approximations (INLA).
-Training data for models cover quarterly tract estimates for Q2 2017 through Q1 2018, while the test data are Q2 2018 median values as observed.
-The top model assumes that spatial and temporal dynamics are separable, i.e.
-independent of each other, and non-interactive.
-This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $100 for the training data and $140 for the test data."
+output:
+  html_document:
+    df_print: paged
+abstract: Online classified ads are a key source of information about local rental
+  housing markets. This research first assesses the agreement of census tract rent
+  estimates based on scraped rental advertisements with existing data sources. The
+  second component of the paper investivates temporal and spatial dynamics in one-bedroom
+  rents among Seattle neighborhoods using spatial models. Posterior distributions
+  of all model hyperparameters and parameters are estimated with integrated nested
+  Laplace approximations (INLA). Training data for models cover quarterly tract estimates
+  for Q2 2017 through Q1 2018, while the test data are Q2 2018 median values as observed.
+  The top model assumes that spatial and temporal dynamics are separable, i.e. independent
+  of each other, and non-interactive. This spatial random-effects model with AR(1)
+  temporal structure achieves a mean absolute error (MAE) of about $100 for the training
+  data and $140 for the test data.
 ---
@@
 
+<br>
+<hr>
+<br>
+
 ## Results

### 2018-07-27 manuscript/manuscript.Rmd
In the manuscript file now.

@@
   estimates based on scraped rental advertisements with existing data sources. The
-  second component of the paper investivates temporal and spatial dynamics in one-bedroom
+  second component of the paper investigates temporal and spatial dynamics in one-bedroom
   rents among Seattle neighborhoods using spatial models. Posterior distributions
@@
 about 2/3 of the 124 2000-definition tracts in Seattle).
-The purple dashed line indicates the loess smoother fit between the measures.
+The purple dashed line indicates the LOESS smoother fit between the measures.
 Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and relative agreement between the measures (mean absolute difference = $186).
@@
 These models are common in epidemiology and disease-mapping applications where researchers need to estimate a quantity of interest like a relative risk but have sparse data for some areas (which may differ from truth due to sampling variability).
-Modelling a panel's error structure with random-effects for space and time penalizes bumpiness (local deviations) in predicted values and theoretically improves out-of-sample accuracy (because large local deviations can stem from low $n$ in a given area).
+Modeling a panel's error structure with random-effects for space and time penalizes bumpiness (local deviations) in predicted values and theoretically improves out-of-sample accuracy (because large local deviations can stem from low $n$ in a given area).
 The spatial models in particular allow for missing data to be imputed based off of a local average of neighbors rather than the global or period-specific average.

### 2018-07-27 manuscript/manuscript.Rmd
Editing the Helena parts.

@@
 We start by briefly reviewing some background relevant to studying neighborhood rental housing.
-I then outline a process for automated web scraping of a rental listing source like Craigslist using a program called Helena, and compare aggregates from the collected rental listings with other data to assess rent agreement in measures.
+I then outline a process for automated web scraping of a rental listing source like Craigslist using a Programming by Demonstration (PBD) tool called Helena and compare aggregates from the collected rental listings with other data to assess rent agreement in measures.
 After describing the data and process, We present a set of models predicting neighborhood rent levels to understand different how different temporal and spatial structures fit the data.
@@
 [Helena](http://helena-lang.org) is a high-level programming language aimed at helping end users automate web scraping as a method of data collection.
-Users write programs in an extension for Google Chrome by demonstrating the first row of data they want to collect using clicks, form interactions and highlighting values.
-The extension translates an initial demonstration into a starting script written in Helena, which a user can further modify or set to run on a regular schedule.
-Though Helena programs collect data roughly at the speed they are recorded at, users can run multiple programs in parallel to increase the volume of data collected.
-Data accumulate on a remote server (soon-to-be AWS) where users can later query either the full table of scraped data or a subset scraped within a specified number of hours.
+Users write programs in an extension for Google Chrome.
+To write a new program, they record themselves collecting the first row of the data they want to collect; they use standard browser interactions -- clicking, typing, and interacting with webpage widgets, and they highlight content they want to scrape.
+The Helena Chrome extension translates this input demonstration into a scraping program written in the Helena language, which a user can run, edit, or schedule for later runs.
+Although Helena programs collect data at roughly the speed of the user demonstration, users can run Helena programs in parallel to increase the rate of data collection.
+Collected data is stored on a remote server, from which users can later download either the full table of scraped data or a subset scraped within a specified window of time.
 
-Helena learns where features of are through the user interface (UI) rather than the underlying document object model (DOM).
-Though the HTML code of a site may regularly change over the course of regular data scraping, websites less frequently overhaul what their users actually interact with.
-Helena also uses skip blocks to avoid re-scraping the an already-collected item.
-These two aspects of Helena provide its programs robustness to common headaches encountered with web scraping by default.
-Overall, Helena's programming-by-demonstration method reduces time needed building a scraper from scratch, giving social scientists the capacity to quickly scrape a single table of data or build ongoing data processes.
+Helena learns how to find data on a webpage using features of the user-facing webpage interface rather than relying on the stability of the underlying document object model (DOM) structure.
+Thus, even if the HTML code of a site changes frequently over the course of longitudinal data scraping, as long as the website limits overhauls of the features users see, Helena can often adapt to webpage redesigns.
+Helena also uses a language construct called the skip block to avoid re-scraping already-collected items.
+These features make Helena programs robust to issues that commonly plague traditional web scraping workflows.
+Overall, Helena's demonstration-based method reduces the time needed to build a scraper from scratch, giving social scientists the capacity to quickly scrape a single dataset or to build ongoing data collection processes.
 
-For the UW team's data collection, a Helena program runs 24-hours a day and scrapes listings from the Seattle Craigslist.
-It is important to note that web scraping does violate Craigslist's Terms of Service, and that we are using all scraped data exclusively for academic research.
+For the UW team's data collection, a Helena program runs 24 hours a day to scrape rental listings from the Seattle Craigslist.
+It is important to note that web scraping does violate Craigslist's Terms of Service and that we are using all scraped data exclusively for academic research.
 For a more thorough discussion of this issue, consult page 3 of [Boeing and Waddell (2016)](https://arxiv.org/ftp/arxiv/papers/1605/1605.05397.pdf).

### 2018-08-01 manuscript/index.Rmd
updating analysis to drop rent field from unique listing definition

@@
   scale_y_continuous(labels = scales::dollar) +
-  scale_color_viridis_c(option = "A", labels = scales::dollar) +
+  scale_color_viridis_c(option = "A", labels = scales::dollar, direction = -1) +
   xlab("") +
@@
   scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80",
-                       limits = c(800, 2800), option = "A") +
+                       limits = c(800, 2800), option = "A", direction = -1) +
   coord_quickmap() +

### 2018-08-01 manuscript/manuscript.Rmd
updating analysis to drop rent field from unique listing definition

@@
 output:
-  html_document:
+  pdf_document:
     df_print: paged

### 2018-08-01 helpers/load_data.R

@@ -11,13 +11,15 @@ load_data <- function(listings = FALSE){
              GISJOIN %in% sea_shp@data$GISJOIN, #only listings with valid Bed/Rent, seattle tracts
              matchType != "Google Maps Lat/Long", #need to have address, not approximate
              cleanBeds %in% c(0, 1, 2, 3)) %>% 
-      distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
-               matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
-      dplyr::select(listingDate, GISJOIN, seattle, matchAddress, matchType, 
-                    cleanBeds, catBeds, cleanRent, cleanSqft, lat, lng) %>% #SELECT these columns
-      mutate(listingDate = as.Date(listingDate),
+      mutate(lat = round(lat, 3),
+             lng = round(lng, 3),
+             listingDate = as.Date(listingDate),
              listingQtr = as.yearqtr(listingDate)) %>%
-      filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3")
+      arrange(desc(listingDate)) %>%
+      distinct(cleanBeds, cleanSqft, lat, lng, GISJOIN, .keep_all = T) %>% #dedupe to unique latXlngXtractXbedXsqft combos
+      dplyr::select(listingDate, listingQtr, GISJOIN, seattle, matchAddress, matchAddress2, matchType, 
+                    cleanBeds, catBeds, cleanRent, cleanSqft, lat, lng) %>% #SELECT these columns
+      filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3") 
     
     #return table of listings if arg listings == TRUE
     if(listings){

### 2018-08-01 helpers/map_point_density.R

@@ -19,7 +19,7 @@ blank <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", geom = "blank")
 with_points <- qmplot(lng, lat, data = map_cl, maptype = "toner-lite", darken = .50, legend = "bottom") +
   stat_density_2d(aes(fill = ..level..), geom = "polygon",
                   alpha = .75, color = NA, size = .05) +
-  scale_fill_viridis_c("1B Listings", option = "A") +
+  scale_fill_viridis_c("1B Listings", option = "A", direction = -1) +
   theme(axis.title = element_blank(),
         axis.text = element_blank(),
         axis.ticks = element_blank(),

### 2018-08-01 helpers/n_heatmap.R

@@ -43,7 +43,7 @@ cal <- plyr::ddply(cal, plyr::.(yearmon), transform, monthweek = 1+week-min(week
 
 #make the heatmap
 ggplot(cal, aes(x = monthweek, y = weekdayf, fill = n)) +
-  scale_fill_viridis_c(na.value = "grey80", direction = 1, option = "A") +
+  scale_fill_viridis_c(na.value = "grey80", direction = -1, option = "A") +
   scale_x_continuous(breaks = c(1, 2, 3, 4)) +
   scale_size(range=c(5,20))+
   geom_tile(color = "white") +

### 2018-08-01 helpers/rent_lineplot.R

@@ -92,7 +92,7 @@ sumRent$dataSrc <- factor(sumRent$dataSrc, levels = levels(sumRent$dataSrc)[c(2,
 ggplot(sumRent, aes(x = moYr, y = Rent, group = dataSrc, color = dataSrc, shape = dataSrc)) +
   geom_line(lwd = 1.0, alpha = .5) +
   geom_point(size = 1.5) +
-  scale_color_viridis_d(begin = .2, end = .8, option = "A") +
+  scale_color_viridis_d(begin = .2, end = .8, option = "A", direction = -1) +
   scale_y_continuous(labels = scales::dollar) +
   scale_x_date(limits = c(min(sumRent$moYr), max(sumRent$moYr))) +
   theme_minimal() +

### 2018-08-01 helpers/rent_map_across.R

@@ -129,7 +129,7 @@ ggplot(tract_f %>% filter(catBeds == "1 Bedroom"), aes(x = long, y = lat, group 
   geom_polygon(data = tract2010_bg, aes(x = long, y = lat, group = group), 
                color = "grey90", fill = "grey70", lwd = .15) +
   geom_polygon(color = "grey90", lwd = .15) +
-  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey70", option = "A") +
+  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey70", option = "A", direction = -1) +
   coord_quickmap() +
   theme(axis.title = element_blank(),
         axis.text = element_blank(),

### 2018-08-01 helpers/load_data.R
reversing table order to chronological prior to deduplication---this is due to focus on time for this memo rather than cross-section only for SAMFR

@@ -15,11 +15,11 @@ load_data <- function(listings = FALSE){
              lng = round(lng, 3),
              listingDate = as.Date(listingDate),
              listingQtr = as.yearqtr(listingDate)) %>%
-      arrange(desc(listingDate)) %>%
+      arrange(listingDate) %>%
+      filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3") %>%
       distinct(cleanBeds, cleanSqft, lat, lng, GISJOIN, .keep_all = T) %>% #dedupe to unique latXlngXtractXbedXsqft combos
       dplyr::select(listingDate, listingQtr, GISJOIN, seattle, matchAddress, matchAddress2, matchType, 
-                    cleanBeds, catBeds, cleanRent, cleanSqft, lat, lng) %>% #SELECT these columns
-      filter(listingQtr >= "2017 Q1", listingQtr < "2018 Q3") 
+                    cleanBeds, catBeds, cleanRent, cleanSqft, lat, lng) #SELECT these columns
     
     #return table of listings if arg listings == TRUE
     if(listings){
